### 2018-02-27 paper/gloaguenAppliedNMF.tex
add paper and experiment

\documentclass[review,5p,twocolumn,sort&compress,times]{elsarticle}

\usepackage{lineno}
% \usepackage{hyperref}

%%%%%% SYMBOLES %%%%%
\usepackage{tipa}	% pour avoir l'accent concave
\usepackage{lmodern}	% pour les guillemets
\usepackage{nth}	% pour ^th près des chiffres

%%%%%% EQUATION %%%%%%
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{fancybox}
\usepackage{xfrac}	% fraction de type "1/4"
\usepackage[fleqn]{cases}	% système équation
\usepackage[overload]{empheq}
\usepackage{bm}		% pour mettre en gras .
\usepackage{units} 	% x/y barre latérale pour les fractions
%\usepackage{split}

%%%%%% FIGURE %%%%%%
\usepackage{graphicx}	% insérer des graphiques
\usepackage{float}	% utiliser H dans les figures
\usepackage{subcaption}

%%%%%% TABLEAUX %%%%%%
\usepackage{array,multirow,makecell}
\usepackage[table,xcdraw]{xcolor} % pour avoir des lignes colorées dans les tableau
\usepackage{hhline}	% pour les lignes horizontales
\usepackage{tabularx} % permet itemize dans les cellules
\usepackage{booktabs}

\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

%%%%%%%%%%%%%%%%%%%%%
\usepackage{url}	% gérer les adresses www.

\modulolinenumbers[5]

\begin{document}

\begin{frontmatter}

\title{Road traffic sound level estimation from realistic urban sound mixtures by Non-negative Matrix Factorization}

\author[ifsttar]{Jean-R\'emy Gloaguen\corref{cor1}}
\ead{jean-remy.gloaguen@ifsttar.fr}
\author[ifsttar]{Arnaud Can}
\author[ls2n]{Mathieu Lagrange}
\author[ls2n]{Jean-Fran\c cois Petiot}

\cortext[cor1]{Corresponding author}
\address[ifsttar]{Ifsttar Centre de Nantes, UMRAE, All\'ee des Ponts et Chauss\'es, 44344 Bouguenais, France}
\address[ls2n]{LS2N, 1 rue de No\"e, 44331 Nantes, France}

\begin{abstract}

Acoustic sensor networks are increasingly deployed in cities, and appear more and more as a possible tool to enrich modeled road traffic noise maps through data assimilation techniques. This, or more simply the validation of the modeled maps through measures, requires first being able to isolate from the measured sound mixtures the road traffic sound level. This task is anything but trivial because of the multiple sound sources that overlap within urban sound mixtures. In this paper, a Non-negative Matrix Factorization (NMF) framework is developed to estimate road traffic noise levels within urban sound scenes. A corpus of sound scenes is artificially built imitating real annotated recordings. The realism of the scenes is validated through a perceptual test, forming a protocol that both reproduces the sensor network outputs, and in which the actual occurrence and sound level of each source is known. Three variants of NMF are tested, namely supervised, semi-supervised, and threshold initialized NMF. While the semi-supervised approach is the most appropriated for park environments, threshold initialized NMF, which is developed especially for this research purpose, appears to be the best generic approach, allowing road traffic noise level estimation with mean average errors of 1.3 dB over the tested corpus of sound scenes.

\end{abstract}

\begin{keyword}
non-negative matrix factorization \sep urban sound environment \sep road traffic sound level
\end{keyword}

\end{frontmatter}

\linenumbers

\section{Introduction}

Noise mapping has been recommended as a tool to tackle noise pollution, in response to the growing demand from urban dwellers for a better environment. The enactment of the European Directive 2002/EC/49 makes such maps mandatory to cities over 100 000 inhabitants. Those maps play an important informative role, establishing the distribution of the sound levels all over the cities as well as the estimation of the number of city dwellers exposed to high sound level ($>$ 55 dB(A)) \cite{nugent2014noise}. Road traffic concentrates particular attention as it is the main urban source of noise annoyance. Road traffic noise maps are built from data collection that consist of traffic data collected on the main roads (flow rates, mean speeds and heavy vehicle ratio) and urban geographic data (building heights and location, topology, ground surfaces, etc.). Follows sound emission and sound propagation computations, resulting in the production of the two indicators equivalent A-weighted sound levels, $L_{DEN}$ (\textit{Day-Evening-Night}) and $L_N$ (\textit{Night}) \cite{kephalopoulos2012common}. This procedure also enables drawing up action plans to reduce the noise exposure. Despite their unanimously recognized interest, noise maps suffer some limitations. The computer efficiency required to produce noise maps at the city scale calls simplifications of the numerical tools and the simulation models that both generate uncertainties \cite{van_leeuwen_noise_2015}. Data collection is itself a vector of uncertainty. Moreover, the produced aggregated indicators mask the sound levels evolution due to the traffic variations throughout the day.

Therefore, noise measurements are increasingly used in addition to simulation to describe urban noise environments. Several measurement set-ups have been proposed in the last years, including  mobile measurements with high quality microphones \cite{manvell2004sadmam, can2014measurement}, participative sensing through dedicated smartphone applications \cite{picaut:halshs-01565214, maisonneuve2009noisetube}, or the development of fixed-sensor networks. In this latter case, the sensor networks can be based either on high-quality sensors as in \cite{mietlicki2012innovative, maijala2018environmental}, or low-cost sensors as in the DYNAMAP project \cite{dynamap_2016} or the CENSE project \cite{picaut2017characterization}. The costs and benefits of each protocol are discussed. Mobile and participatory measures increase spatial coverage at low cost, but lack temporal representativeness. Fixed networks are very reliable for measuring sound levels temporal variations, but allow only a small spatial coverage of the network. In addition, the low-cost sensors enable a wider deployment, but at the cost of increased uncertainties, the most extreme example being smartphone applications.

\begin{figure}[t]
\centering
\includegraphics[width=.9\linewidth]{figures/bloc_diagram_source_separation.pdf}
\caption{Block diagram of the blind source separation model}
\label{fig:source_separation}
\end{figure}

All these measurement protocol enables a priori combining measures and modeling to improve the accuracy of the produced noise maps. Traffic noise maps and measurements were compared on restrictive areas in \cite{lefebvre2017traffic} and \cite{mioduszewski2011noise}. Wei et al. modify the acoustical parameters of the simulation thanks to noise measurements. Mallet et al. call for data assimilation techniques between models and measurements to reduce the uncertainty of the produced noise maps. However, these works make the implicit assumption that the noise measurements consist mainly of road traffic. In the aim to improve road traffic noise maps, the use of measurements has first to deal with the challenge to estimate correctly the road traffic sound level.
Even if road traffic is predominant on many urban areas, urban sound environments are composed of many different overlapping sound sources (passing cars, voices, footsteps, car horn, whistling birds \dots), what makes the task of estimating correctly the traffic sound level within an urban sound mixture not trivial.

Many works have dealt with the detection \cite{luitel2016sound} or the recognition \cite{defreville_automatic_2006} of sound events in environmental sound scenes. In these cases, a usually two-step schemes is followed were audio samples are described with a set of features (Mel Frequency Cepstral Coefficient, MPEG-7 descriptors \dots) and classified them with the help of a classifiers (Gaussian Mixtures Models, Artificial Neural Network \dots) \cite{chu2008environmental, cowling_comparison_2003}. The classifiers is learnt from a learning database and are next applied on a test database to validate the algorithms. 
Dedicated to the traffic, in \cite{socoro_anomalous_2017}, an Anomalous Event Detection, based on MFCC features, is proposed with the specific aim to improve the traffic sound estimation. It is based on the detection of sound events in order to discard them.

An other approach, followed in this paper, is to consider the Blind Source Separation paradigm, see Figure \ref{fig:source_separation}, which consists in the extraction of a specific signal inside a set of mixed signals. From the different existing methods (CASA, ICA), Non-negative Matrix Factorization (NMF) \cite{lee_learning_1999}, seems the most relevant method for monophonic sensor networks. Dealing easily with the overlapping between the sound sources, this method approximes the magnitude spectrogram of a single signal by the product of two non negative matrices. A lot of applications can be found for musical \cite{smaragdis_non-negative_2003, benetos2006musical} and speech \cite{wilson_speech_2008, mysore2011non} contents. Dedicated to sound separation, Immani and Kasa\"i used NMF in a two steps sound separation with the help of time variant gain features. A first study \cite{} has been conducted, in which diverse NMF formula, namely the supervised, the semi-supervised, and the threshold initialized NMF, have been applied on a large set of simulated sound scenes that which mixed traffic component with specific urban sounds at calibrated sound levels. The study proved the interest of NMF for urban sound environments and compared the benefits of each approach. However, it has now to face to real urban sound scenes.

In this paper, the NMF framework is applied on a corpus of simulated sound scenes, generated based on annotated urban recordings, and whose realism is validated by a perceptual test. The NMF framework and its different implemented versions are described in section \ref{part:nmf}. Next, the corpus of urban sound scenes is presented in section \ref{part:urban_scene}, from the sound database built-up to its validation through a perceptual test. Finally, in section \ref{part:expProtocol} and \ref{part:results}, the experimental protocol and the results are exposed.

\section{Non-negative Matrix Factorization}\label{part:nmf}

Non-negative Matrix Factorization (NMF) is a linear approximation method proposed by Paatero and Tapper \cite{paatero1994positive} and popularized by Lee and Seung \cite{lee_learning_1999}. It consists in approximating a non negative matrix $\mathbf{V}$ $\in \mathbf{R}^+_{F \times N}$ by the product of two non negative matrices: $\mathbf{W}$, called \textit{dictionary} (or basis), and $\mathbf{H}$, called the matrix \textit{activation} with the dimensions $F \times K$ and $K \times N$ respectively.

\begin{equation}\label{eq:nmf}
\mathbf{V} \approx \mathbf{WH}.
\end{equation}

The choice of the dimensions is often made such as $F\times K + K \times N < F \times N$ so that NMF can be a low rank approximation. This condition however is not mandatory. When an audio file is considering, $\mathbf{V}$ is usually considered as the magnitude spectrogram obtained by a Short-Time Frequency Transform, $\mathbf{W}$ includes audio spectra and $\mathbf{H}$ is equivalent to the temporal evolution of each spectrum, see Figure \ref{fig:exampleNMF}. Because of the non-negativity constraint, only additive combinations between the elements of $\mathbf{W}$ are considered. The dictionary is then composed of elementary elements providing a part based representation.

\begin{figure}[t]
\includegraphics[width=.9\linewidth]{./figures/schema_introduction_nmf.pdf} 
\caption{NMF for an audio sample with 3 elements ($K$ = 3): passing car (a), car horn (b) and whistling bird (c)}
\label{fig:exampleNMF}
\end{figure}

The approximation of $\mathbf{V}$ by $\mathbf{WH}$ product is defined by a cost function to minimize,

\begin{equation}\label{eq:min-D-WH}
\underset{\mathbf{H} \geq 0, \mathbf{W} \geq 0}{\min} D\left(\mathbf{V} \Vert \mathbf{WH}\right), 
\end{equation}

where $D(\bullet \Vert \bullet)$ is a divergence calculation such as:
\begin{equation}
D\left(\textbf{V} \vert\vert \mathbf{WH} \right) = \sum_{f = 1}^{F} \sum_{n = 1}^{N} d_{\beta}
\left(\textbf{V}_{fn} \vert \left[ \textbf{WH} \right]_{fn} \right).
\end{equation}

$d_{\beta}$ is mainly chosen as a $\beta$-divergence \cite{fevotte_algorithms_2011}, a sub-classes belonging to the Bregman divergences \cite{hennequin_beta-divergence_2011} which include 3 specific divergence and distance calculations: the Euclidean distance (eq. \ref{eq:def_distEUC}), the Kullback-Leibler divergence (eq. \ref{eq:def_divKL}) and the Itakura-Sa\"{i}to divergence (eq. \ref{eq:def_divIS}):

\begin{subequations}\label{eq:divBetaGenerale}
\begin{numcases}{d_{\beta}(x\vert y) =}
    \frac{1}{2}(x-y)^2, & $\beta = 2$, \label{eq:def_distEUC}\\
    x\log \dfrac{x}{y} - x + y, & $\beta = 1$, \label{eq:def_divKL} \\
    \frac{x}{y} - \log \frac{x}{y} - 1 & $\beta = 0$. \label{eq:def_divIS}
\end{numcases}
\end{subequations}

The minimization problem (\ref{eq:min-D-WH}) is solved iteratively by updating the form of matrices $\mathbf{W}$ and $\mathbf{H}$. Different algorithms such as Alternating Least Square Method \cite{cichocki_regularized_2007} or Projected Gradient \cite{lin_projected_2007} have been proposed. The most commonly algorithm used, and the chosen method here, is Multiplicative Update \cite{lee_algorithms_2000} as it ensures non-negative results and the convergence of the results \cite{fevotte_algorithms_2011}.

\subsection{Supervised NMF}
The most easiest case of NMF is the one where the sound sources can be known \textit{a priori} and $\mathbf{W}$ can be built directly from audio samples. It leads to \textit{supervised} NMF (SUP-NMF). $\mathbf{H}$ is then the only matrix to determine and is updated at every iteration (eq. \ref{eq:updateH}) \cite{fevotte_algorithms_2011}.

\begin{equation} \label{eq:updateH}
\textbf{H}^{(i+1)} \leftarrow \textbf{H}^{(i)}\otimes\left(\frac{\textbf{W}^T \left[\left(\textbf{WH}^{(i)} \right)^{(\beta-2)}\otimes\textbf{V} \right]}{\textbf{W}^T \left[\textbf{WH}^{(i)} \right]^{(\beta-1)}}\right)^{\gamma(\beta)}
\end{equation}

with $\gamma(\beta) = \frac{1}{2-\beta},$ for $\beta < 1$, $ \gamma(\beta) = 1$, for $\beta \in \left[1,2\right]$ and $\gamma(\beta) = \frac{1}{\beta-1}$ for $\beta > 2$. The product $A\otimes B$ and $A/B$ symbolized the Hadamard product and ratio.

Here, in an urban context, the sound sources are known and their audio samples can be obtained to learn $\mathbf{W}$, see section \ref{part:dictionary_building}. As the position of each element is indexed, the traffic source separation from the other sound sources is made by extracting, from the dictionary and the activation matrix, the related elements:

\begin{equation}\label{eq:separationExtraction}
\mathbf{\tilde{V}}_{traffic} = \left[ \mathbf{WH} \right]_{traffic}.
\end{equation}

\subsection{Semi-supervised NMF}
The main issue with the supervised approach is the limit imposed by $\mathbf{W}$: to be successful, all the acoustical sources must be considered in it which is not possible in an urban environment. As the main source that interests us can be known \textit{a priori}, semi-supervised NMF (SEM-NMF) \cite{lee_semi-supervised_2010} is considered to better take into account the interfering sound sources. This approach proposes to decompose $\mathbf{W}_{F \times (K+J)}$ with two distinctive matrices: $\mathbf{W} = \left[ \mathbf{W_s}~\mathbf{W_r} \right]$ where $\mathbf{W_s}_{F \times K}$ is a fixed part of $\mathbf{W}$ composed of fixed audio spectra and $ \mathbf{W_r}_{F \times J}$, a mobile part which is updated, see eq. \ref{eq:W_r_SS}. Thus it is possible to include elements not present in $\mathbf{W_s}$. The dimension of $\mathbf{W_r}$ is set up as $J << K$ in order to best considered the sound source present in $\mathbf{W_s}$. $\mathbf{H}$ is then also decomposed in two matrices, $\mathbf{H}_{(K+J) \times N} = \genfrac[]{0pt}{0}{\mathbf{H_s}}{\mathbf{H_r}}$. The eq. \ref{eq:nmf} becomes

\begin{equation}
\mathbf{V} = \mathbf{W_s H_s}+\mathbf{W_r H_r}.
\end{equation}

$\mathbf{H_r}$ and $\mathbf{H_s}$ are updated separately, see eq. \ref{eq:H_r_SS} \ref{eq:H_s_SS}.

{\scriptsize
\begin{subequations}\label{eq:WH-SSupdate}
\begin{align}
\mathbf{W_r}^{(i+1)} &\leftarrow \mathbf{W_r}^{(i)}\otimes\left(\frac{\left[\left(\mathbf{W_r H_r}^{(i)} \right)^{(\beta-2)}\otimes\mathbf{V} \right]\mathbf{H_r}^T}{\left(\mathbf{W_r H_r}^{(i)} \right)^{(\beta-1)}\mathbf{H_r}^T}\right)^{\gamma(\beta)}.\label{eq:W_r_SS}\\
\mathbf{H_r}^{(i+1)} &\leftarrow \mathbf{H_r}^{(i)}\otimes\left(\frac{\mathbf{W_r}^T \left[\left(\mathbf{W_r H_r}^{(i)} \right)^{(\beta-2)}\otimes\mathbf{V} \right]}{\mathbf{W_r}^T \left(\mathbf{W_r H_r}^{(i)} \right)^{(\beta-1)}}\right)^{\gamma(\beta)},\label{eq:H_r_SS}\\
\mathbf{H_s}^{(i+1)} &\leftarrow \mathbf{H_s}^{(i)}\otimes\left(\frac{\mathbf{W_s}^T \left[\left(\mathbf{W_s H_s}^{(i)} \right)^{(\beta-2)}\otimes\mathbf{V} \right]}{\mathbf{W_s}^T \left(\mathbf{W_s H_s}^{(i)} \right)^{(\beta-1)}}\right)^{\gamma(\beta)},\label{eq:H_s_SS}
\end{align}
\end{subequations}}

In this study, $\mathbf{W_s}$ is composed of traffic audio spectra to include in $\mathbf{W_r}$ all the other sources that can be present in the urban sound scenes. The traffic signal estimation is next defined by the fixed part,

\begin{equation}\label{eq:separationExtraction_SS}
\mathbf{\tilde{V}}_{traffic} = \left[ \mathbf{W_s H_s} \right].
\end{equation}

This approach as the advantage, with the add of the mobile part $\mathbf{W_r}$, to bring more flexibility and then to be more adaptive to the different urban sound environments. Applications of Sem-NMF can be found for musical \cite{weninger2012supervised, kitamura_music_2014} and speech content \cite{joder2012real, mysore2011non}.

\subsection{Thresholded Initialized NMF}\label{part:NMF_TI}

We propose a third approach based on the unsupervised NMF framework: Threshold Initialized NMF (TI-NMF). Usually, in unsupervised NMF, the dictionary is initiated randomly when there is no \textit{prior} knowledge on the sound sources present. Here, as the target sound source is known and the spectra are available, an initial dictionary, $\mathbf{W_0}$, is designed and then updated alternatively with $\mathbf{H}$,

\begin{equation}\label{eq:updateW_unsup}
\textbf{W}^{(i+1)} \leftarrow \mathbf{W}^{(i)}\otimes \left(\frac{\left[\left(\mathbf{W}^{(i)}\mathbf{H} \right)^{(\beta-2)}\otimes \mathbf{V} \right]\mathbf{H}^T}{\left[\mathbf{W}^{(i)}\mathbf{H} \right]^{(\beta-1)}\mathbf{H}^T}\right)^{\gamma(\beta)}. 
\end{equation}

With this operation, $\mathbf{W_0}$ is oriented to the focused sound source (the road traffic) but also can be adapted to the content of the scene thanks to the updates. After $N$ iterations, each element $k$ of the final dictionary, $\mathbf{W'}$, is compared with its initial value in $\mathbf{W_0}$, in order to identify which element is stayed closed to the traffic component. A cosine similarity $D_{\theta}\left(\mathbf{W_0} \Vert \mathbf{W'} \right)$ is computed for each element $k$ as it is an invariant scale and a bounded method,

\begin{equation}
D_{\theta}\left(\mathbf{w_0} \Vert \mathbf{w'} \right) = \frac{\mathbf{w_0}.\mathbf{w'}}{\Vert \mathbf{w_0}  \Vert . \Vert \mathbf{w'} \Vert}.
\end{equation}

where $\mathbf{w}$ is a $k$ element of $\mathbf{W}$ of $F \times 1$ dimension. When $D_{\theta}\left(\mathbf{w_0} \Vert \mathbf{w'} \right)$=1, the element $k$ from $\mathbf{W'}$ is exactly similar than in $\mathbf{W_0}$. If $D_{\theta}\left(\mathbf{w_0} \Vert \mathbf{w'} \right)$=0, the element is fully different. Next, the similarities are sorted in descending order. The extraction of traffic elements in $\mathbf{W'}$ is carried out by two thresholding methods. For both, the traffic elements are estimated by weighting $\mathbf{W'}$ according to $D_{\theta}\left(\mathbf{w_0} \Vert \mathbf{w'} \right)$ and a threshold value such as :

\begin{equation}
\mathbf{w}_{traffic} = \alpha_k \mathbf{w'}.
\end{equation}

The first method is the hard thresholding \cite{donoho1995noising}. It selects the traffic elements from $\mathbf{W'}$ in a binary way by comparing their similarities to a fixed threshold $t_h$

\begin{subequations}\label{eq:WH-TIupdateHard}
\begin{numcases}{\alpha_k =}
1  & iff \quad $D_{\theta}\left(\mathbf{w}_{0} \Vert \mathbf{w'} \right) > t_h$, \\
 0 & else.
\end{numcases}
\end{subequations}

A second approach is based on firm thresholding \cite{gao1997waveshrink} which considers two thresholds, $t_{f,1}$ and $t_{f,2}$ with $t_{f,1}$ $>$ $t_{f,2}$. The weight $\alpha_k$ is then defined as

\begin{subequations}\label{eq:WH-TIupdateFirm}
\begin{numcases}{\alpha_k =}
	1 & {\scriptsize iff $D_{\theta}\left(\mathbf{w}_{0} \Vert \mathbf{w'} \right) > t_{f,1}$} \\
    0 & {\scriptsize iff $D_{\theta}\left(\mathbf{w}_{0} \Vert \mathbf{w'} \right) \leq t_{f,2}$} \\
	\mathcal{N} \left(D_{\theta}(\mathbf{w_0}\Vert \mathbf{w'}) \right)& else.
\end{numcases}
\end{subequations}

% \dfrac{d_{\beta}(\mathbf{w_0}\Vert \mathbf{w'}) - \min(d_{\beta}(\mathbf{w_0}\Vert \mathbf{w'}))}{\max(d_{\beta}(\mathbf{w_0}\Vert \mathbf{w'}))-\min(d_{\beta}(\mathbf{w_0}\Vert \mathbf{w'}))}
where $\mathcal{N}\left(D_{\theta}(\mathbf{w_0}\Vert \mathbf{w'}) \right)$ is a normalization between 0 and 1 of the similarity. 
The most similar elements to $\mathbf{W}_0$ are then keeping as they are whereas the elements whose the similarities are located between the two thresholds are then weighted. The allure of $D_{\theta}\left(\mathbf{W_0} \Vert \mathbf{W'} \right)$ and $\alpha$ are displayed in Figure \ref{fig:example_threshold}.\\

\begin{figure}[t]
    \centering
       \includegraphics[width=0.9\linewidth]{./figures/thresh.pdf}
    \caption{Evolution of the $\alpha$ weight according to $D_{\theta}\left(\mathbf{W_0} \Vert \mathbf{W'} \right)$ for hard thresholding with $t_h$ = 0.6 (in dashed line) and firm thresholding with $t_{f,1}$ = 0.67 and $t_{f,2}$ = 0.40 (in full line)}
    \label{fig:example_threshold}
\end{figure}

These methods are applied on simulated sound scenes in order to compare the estimated sound levels with the exact solutions. In \cite{}, the sound corpus was composed of a mix of traffic samples with specific sound classes (\textit{alert, animals, climate, human, mechanics, transportation}. Here, in order to implement this method in embedded sensors, a new more realistic sound corpus is generated based on urban recordings mixing all kind of sound sources.

\section{Design of realistic urban sound scenes}\label{part:urban_scene}

The urban sound scenes are taken from 76 recordings from 2 to 5 min, achieved in the \nth{13} district of Paris (France) at 19 different locations \footnote{Recordings were made as part of the Grafic project funded by Ademe}, which cover four various sound environments (Figure \ref{fig:map_grafic}). A complete description of the experimental protocol can be found in \cite{aumond_modelling_2017}. Two of the 76 recordings are rejected for the analysis because the audio files were corrupted, resulting in 74 valid audio files assumed as representative of the variety of sound environments. The recordings are listened and categorized within four different sound environments, as proposed in \cite{can_describing_2015}: park (P, 8 audio files), quiet street (Q, 35 audio files), noisy street (N, 23 audio files) and very noisy street (vN, 8 audio files). Then, each audio file is annotated, noting the start and end time of each sound event along with its sound class (described here below). The aim of the annotation phase is next to transcribe the recordings, in order to obtain simulated sound scenes with the same distribution of sound events as the recordings and therefore as close as possible to the realistic scenes.

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{./figures/trajet_19pts.png}
\caption{Walked path with the 19 stop points  \cite{aumond_modelling_2017}}.
\label{fig:map_grafic}
\end{figure}

\subsection{Transcription of the recordings}\label{part:simScene}
The sound scenes are generated with the \textit{SimScene} software\footnote{Open-source project available at: \url{https://bitbucket.org/mlagrange/simscene}}, \cite{rossignol_simscene:_2015}, which is a simulation software generating monaural sound mixtures in wav format with a 44.1 kHz sampling rate from an isolated sound database. 
This software have already been used in a wide range of experiments for sound detection algorithm assessment \cite{lafay_new_2014, benetos2016detection}. The control of high level parameters can be handle by the user as the sound class presence, the time between each sample of one sound class, the ratio between the sound level of an event class with the background (i.e the \textit{event background ratio} shorten \textit{ebr})\dots Each parameter is completed with a standard deviation to bring random behavior. It allows too the design of a sound mixture from an annotation text file. 
As output, \textit{SimScene} generates an audio of the global sound mixture and an audio for each sound class present in the scene, which makes it possible to know their exact contributions in the scene, see Figure \ref{fig:example_simScene}.

\begin{figure}[t]
    \centering
       \includegraphics[width=.8\linewidth]{./figures/exampleSimScene.pdf}
    \caption{Spectrogram of a simple scene created with the \textit{SimScene} software with a sound background (road traffic in green) and 3 sound events (car horn in purple, passing car in red and whistling bird in blue)} 
    \label{fig:example_simScene}
\end{figure}

To transcribe the recordings in simulated scenes, a high quality sound database (wav format, 44.1 kHz sampling rate, high \textit{Signal Noise Ratio}) has been built-up from audio samples found online (\textit{freesound.org}) or with the help of an already existing sound database \cite{salamon2014dataset}. The sound database is composed of two categories of sound : the \textit{event} category, which includes 245 brief sound samples considered as salient, with a 1 to 20 seconds duration and classified among 21 sound classes (\textit{ringing bell, whistling bird, car horn, passing car, hammer, barking dog, siren, footstep, metallic noise, voice \dots}) and the \textit{background} (or \textit{texture}) category gathering 154 long duration sounds ($\approx$ 1m30), whose acoustic properties do not vary in time. This category includes among others\textit{whistling bird, crowd noise, rain, children playing in schoolyard, constant traffic noise} \dots sound classes. Each sound class is present in multiple samples (\textit{carHorn01.wav, carHorn02.wav \dots}) to bring diversity.
As the road traffic is the main component in urban environment and is the sound source of interest, recordings of car passages has been made on the Ifsttar's runway. The recordings has been made for 4 cars (Renault Senic, Renault Megane and Renault Clio, Dacia Sandero),  for different speeds and gear ratios. In all, 103 car passages have been recorded. The audio samples of the first two cars (Renault Senic and Renault Megane) are included in the \textit{SimScene}'s sound database (50 audio files totally). The last 53 audio samples are dedicated to the dictionary design as part of NMF, see section \ref{part:dictionary_building}. A full description of the recording can be found in \cite{gloaguen_creation_2017}.

With this built-up database, the \textit{SimScene} software and the annotations of the recordings, 74 simulated sound scenes are generated, which have the same temporal structure than the recordings. The \textit{ebr} parameter is adjusted manually on each sound scene to be faithful compared to the recorded scenes. To validate its realism, the corpus of transcribed urban sound scenes is submitted to a perceptual test.

\subsection{Perceptual test}
The perceptual test is conducted with a panel of 50 listeners that are asked to assess the level of realism on a 7-point scale (1 is \textit{not realistic at all}, 7 is \textit{very realistic}) of a mix of transcribed and recorded scenes. The total number of sound scenes tested is set at 40. The first half includes 20 30-seconds audio files, including 5 scenes that belong to the sound environment \textit{park}, 6 from \textit{quiet street}, 4 from \textit{noisy street} and 5 from \textit{very noisy street} chosen randomly among the recorded scenes. The second half is composed of the same 30-second transcribed scenes. In order to limit the duration of the test and to preserve the concentration of the listeners, each of them listens a subset of 20 sound scenes, which mix recorded and transcribed samples. Furthermore, all the scenes are normalized to the same sound level, chosen at 65 dB, to prevent the listeners from changing the sound level of their speakers.

The experimental design is elaborated following a partially Balanced Incomplete Block Design (PBIBD) \cite{john1977optimal} that determines the listening order of each participant based on fixed parameters (number of listeners, total number of audio, number of audio assessed by each listeners). The listening order is then built in a such way than each listener listens the same amount of recorded scenes and transcribed scenes to avoide statistical biais. The experimental design and the listening order per participant are performed with the package \textit{sensoMineR} on the \textit{R} software \cite{le_sensominer:_2008}.\\

The test was administered online on the 8 February 2017 and the number of participant has been reached 12 days later. During the test, the participants had the possibility to listen to each scene as many times as wanted before assessing, without being able to change their judgment afterwards. The participants could also leave a comment on each audio to explain the rating. Based on the information provided, the panel of 50 listeners was made of 31 males and 18 females (one not documented) with an average age of 36 ($\pm$ 12) years old. $62\%$ of the participants declared having no experience in the listening of urban sound mixtures. Figure \ref{fig:boxPlot_test} summarizes the score distributions of all the listeners for the recorded (rec.) and the transcribed (trans.) scenes.

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{figures/testPerceptif_boxplotType_EN.pdf} 
\caption{Box-and whiskers plot of the rating of realism according to the type (recorded /transcribed)}
\label{fig:boxPlot_test}
\end{figure}

The distributions of the notations according to the type are extremely similar. The mean score for the transcribed scenes is even superior to the recorded ones ($m_{trans.}$ = 5.1 ($\pm 1.6$), $m_{rec}$ = 4.9 ($\pm 1.6$)). These results confirm that all the recorded and the transcribed scenes are perceived in a similar way by the panel. More details on the results can be found in  \cite{gloaguen_creation_2017}.

As the transcribed sound corpus is sufficiently close to the recordings, these sound mixtures can be used to assess the performances of NMF to estimate the traffic sound level.

\section{Experimental protocol}\label{part:expProtocol}

The experiment protocol consists in estimate, on the 74 scenes available classified on 4 sound environments, the equivalent sound level of the traffic of the entire scenes, $\tilde{L}_{p, traffic}$ (dB) and to compare them to their exact values given by the simulation process, $L_{p,traffic}$, see Figure \ref{fig:bloc_diagram_estimator}.

\begin{figure}[t]
\centering
\includegraphics[width=.9\linewidth]{figures/bloc_diagram_estimator.pdf}
\caption{Bloc diagram of the experimental protocol}
\label{fig:bloc_diagram_estimator}
\end{figure}

\begin{figure*}[t]
\centering
\includegraphics[width=0.8\linewidth]{figures/extractionDictionary2.pdf}
\caption{Dictionary building on a 3 second example of a passing car with $w_t = 1$ second. The dictionary are cut in 3 frames of $w_t$ duration. On each the rms value is calculated to obtained 3 spectra destined to $\mathbf{W}$.}
\label{fig:example_dictionary}
\end{figure*}

As the road traffic is mainly composed of a low frequency content, a first estimator is considered through a frequency low-pass filter (LP filter). It consists in filtering the sound scenes at different cut-off frequencies $f_c \in \lbrace$500, 1, 2k, 5k, 10k, 20k$\rbrace$ Hz. The remaining energy located in the pass-band is assimilated to road traffic,

\begin{equation}
\mathbf{\tilde{V}}_{traffic} = \mathbf{V}_{f_c}.
\end{equation}

The second estimator is based on the three NMF formula presented in part \ref{part:nmf} (see Figure \ref{fig:bloc_diagram_nmf}). Between the dictionary building to the metric calculation, multiple experimental factors take part in these cases where each of them having different modalities.

\begin{figure}[t]
\centering
\includegraphics[width=.9\linewidth]{figures/bloc_diagram_NMF_EN_2.pdf}
\caption{Bloc diagram of the experimental protocol}
\label{fig:bloc_diagram_nmf}
\end{figure}

\subsection{Dictionary building}\label{part:dictionary_building}
The dictionary building is designed from a second sound database specially dedicated to this task to prevent any overfitting issues. It contains the 53 audio files of the 2 other cars (Dacia Sandero and Renault Clio) recorded on the runway, see part \ref{part:simScene}.

First the spectrogram of each audio file is computed (window $w = 2^{12}$ with 50 $\%$ overlap). The spectrogram is then cut in multiple temporal frames of $w_t  = \lbrace$0.5, 1$\rbrace$ second duration. In each cut spectrogram, the root mean square on each frequency bin is calculated to obtained a spectrum of $F \times 1$ dimension. This method allows the description of the audio sample with finer spectra and then having the different characteristic pitches of the traffic spectra. An illustrative example on a 3 seconds sample is displayed in Figure \ref{fig:example_dictionary}. From the 53 audio files, we obtain respectively for $w_t\in$ $\lbrace$0.5, 1$\rbrace$ second, 2218 and 1109 elements. A $K$-mean clustering algorithm is applied to reduce these dimensions to $K \in \lbrace$25, 50, 100, 200$\rbrace$ in order to avoid redundant information and decrease the computation time. The obtained $K$ clusters compose then the $K$ elements of $\mathbf{W}$. Furthermore, the case where each audio generates one spectrum from its spectrogram is added ($w_t = all$). Here, by the added approach, the dictionary elements are based on the spectral envelops of the audio samples. For this case, having 53 audio samples, the number of elements in $\mathbf{W}$ with the $K$-mean clustering algorithm is reduced to $K \in \lbrace$25, 50$\rbrace$.

The obtained dictionary is expressed in third octave bands to decrease the dimensionality without deteriorate the spectral content. A previous experimental validation revealed that the use of third octave bands, instead of linear spectra, does not impact the performances of the chosen estimator in this study.  Finally, each basis of $\mathbf{W}$ is normalized as $ \Vert \mathbf{w}_k \Vert = 1$ where $\Vert \bullet \Vert$ is the $\ell_1$ norm. Table \ref{tab:experimental_factorsNMF} summarizes the different modalities of the two experimental factors ($K$ and $w_t$).

\begin{table*}[t]
\caption{Summary of the different experimental factors and their modalities taken into account in the frequency low-pass filter estimator}
\centering
\begin{tabularx}{17.5cm}{L{3cm}@{}C{12cm}@{}C{2cm}@{}}
	\hline
    \textbf{\begin{tabular}[c]{@{}l@{}}experimental \\ factors\end{tabular}} & \textbf{modalities} & \begin{tabular}[c]{@{}C{2cm}@{}}\textbf{number of}\\ \textbf{modalities}\end{tabular}\\ \toprule
\end{tabularx}

\begin{tabularx}{17.5cm}{L{3cm}@{}@{}C{3cm}@{}@{}C{3cm}@{}@{}C{3cm}@{}@{}C{3cm}@{}C{2cm}@{}}
    \textbf{sound environment} & park & quiet street & noisy street & very noisy street & 4 \\
\end{tabularx}

\begin{tabularx}{17.5cm}{L{3cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}C{2cm}@{}}
\rowcolor[HTML]{C0C0C0} 
   $\mathbf{f_c}$ (kHz) & 0.5 & 1 & 2 &  5 & 10 & 20 & 6\\
   \bottomrule
\end{tabularx}

\label{tab:experimental_factorsFilter}
\end{table*}

\begin{table*}[t]
\centering
\caption{Summary of the different experimental factors and their modalities taken into account in NMF estimator}
\begin{tabularx}{17.5cm}{L{3cm}@{}C{12cm}@{}C{2cm}@{}}
	\hline
    \textbf{\begin{tabular}[c]{@{}l@{}}experimental \\ factors\end{tabular}} & \textbf{modalities} & \begin{tabular}[c]{@{}C{2cm}@{}}\textbf{number of}\\ \textbf{modalities}\end{tabular}\\ \toprule
\end{tabularx}

\begin{tabularx}{17.5cm}{L{3cm}@{}C{3cm}@{}@{}C{3cm}@{}@{}C{3cm}@{}@{}C{3cm}@{}C{2cm}@{}}
    \textbf{sound environment} & park & quiet street & noisy street & very noisy street & 4\\
\end{tabularx}

\begin{tabularx}{17.5cm}{L{3cm}@{}C{4cm}@{}@{}@{}C{4cm}@{}@{}C{4cm}@{}C{2cm}@{}}	
	\rowcolor[HTML]{C0C0C0} 
  \textbf{method} & Sup NMF & Sem NMF & TI NMF & 3\\
\end{tabularx}

\begin{tabularx}{17.5cm}{L{3cm}@{}C{4cm}@{}@{}C{4cm}@{}@{}C{4cm}@{}C{2cm}@{}}
    $\mathbf{w_t}$ (s)& 0.5 & 1 & \textit{all} & 3 \\
\end{tabularx}

\begin{tabularx}{17.5cm}{L{3cm}@{}C{3cm}@{}@{}C{3cm}@{}@{}C{3cm}@{}@{}C{3cm}@{}C{2cm}@{}}
	\rowcolor[HTML]{C0C0C0} 
    $\mathbf{K}$ & 25 & 50 & 100 & 200 & 4\\
\end{tabularx}

\begin{tabularx}{17.5cm}{L{3cm}@{}C{4cm}@{}@{}C{4cm}@{}@{}C{4cm}@{}C{2cm}@{}}
   $\mathbf{\beta}$ & 0 & 1 & 2 & 3\\
\end{tabularx}

\begin{tabularx}{17.5cm}{L{3cm}@{}C{12cm}@{}C{2cm}@{}}
	\rowcolor[HTML]{C0C0C0} 
   hard threshold $\mathbf{t_h}$ & from 0.30 to 0.60 with a 0.01 step & 31\\
\end{tabularx}

\begin{tabularx}{17.5cm}{L{3cm}@{}C{12cm}@{}C{2cm}@{}}
   high firm threshold $\mathbf{t_{f,1}}$ & from 0.30 to 0.60 with a 0.01 step & 31\\
\end{tabularx}

\begin{tabularx}{17.5cm}{L{3cm}@{}C{12cm}@{}C{2cm}@{}}
	\rowcolor[HTML]{C0C0C0} 
      low firm threshold $\mathbf{t_{f,2}}$ & from 0.10 to 0.45 with a 0.01 step & 36\\
   \bottomrule
\end{tabularx}

\label{tab:experimental_factorsNMF}
\end{table*}

\subsection{NMF experimental factors}

NMF is performed for 3 $\beta$-divergences: $\beta$ = 2 (euclidean distance), $\beta$ = 1 (Kullback-Leibler divergence) and $\beta = 0$ (Itakura-Sa\"ito divergence). The spectrogram $\mathbf{V}$ and the dictionary $\mathbf{W}$ are displayed in a logarithmic scale through a third band octave representation that reduces the high frequency predominance where the traffic component is absent. In addition, as the number of frequency bins is reduced ($F$ = 29), the computation time is reduced too. 400 iterations are performed to get a stabilized results. For SEM-NMF, the number of elements in $\mathbf{W_r}$ is fixed at $J = 2$. For hard thresholding, the threshold value, $t_h$ is defined between 0.30 and 0.60. For firm thresholding, the high threshold, $t_{f,1}$, is fixed between 0.45 and 0.60, the low threshold, $t_{f,2}$, is fixed between 0.10 and 0.45. All the intervals are defined with a 0.01 step. The cases $t_{f,1} < t_{f,2}$ are, naturally, discarded, see part \ref{part:NMF_TI}.
Each unique association of modalities between each experimental factor forms a setting. For the filter estimator, 24 settings are computed (4 $\times$ 6). For SUP and SEM-NMF, 240 settings are computed (4 $\times$ 2 $\times$ (2 $\times$ 4 + 1 $\times$ 2) $\times$ 3). Finally, for TI-NMF, the number of settings is much higher (115 320) due to the multiple possible combinations between the threshold values (4 $\times$ 1 $\times$ (2 $\times$ 4 + 1 $\times$ 2) $\times$ 3 $\times$ (31 + (20 $\times$ 31 + $\sum_{i = 15}^{30} i $))).
The summarize of the experimental factors and their different modalities is displayed on Tables \ref{tab:experimental_factorsFilter} and \ref{tab:experimental_factorsNMF}.

The approximated traffic spectrograms $\mathbf{\tilde{V}}_{traffic}$ are obtained after 400 iterations. The estimated traffic sound level in dB, $\tilde{L}_{p, traffic}$, is deducted,

\begin{equation}
\tilde{L}_{p, traffic} = 20\log_{10}\frac{p_{rms}}{p_0}, 
\end{equation}

where $p_0$ is the reference sound pressure, $p_0 = 2\times 10^{-5}$ Pa. For each setting, $M$ traffic sound levels, corresponding to the $M$ scenes of each sound environment, are then calculated.

\subsection{Metrics}

The traffic sound levels, $\tilde{L}_{p,traffic}$, are compared to the exact values, $L_{p,traffic}$, through the Mean Absolute Error ($MAE$) \cite{willmott2005advantages}. The $MAE$ expresses the quality of the long-term reconstruction of the signal and consists in the average of the absolute difference between the exact and the estimated sound levels,

\begin{equation}
MAE = \frac{\sum_{i = 1}^{M} \vert L_{p,traffic}^i - \tilde{L}_{p,traffic}^i \vert}{M}.
\end{equation}

It is then possible to express the $MAE$ error for each unique setting but, also, to average this metric according to the 4 sound environments to be able to estimate the optimal setting that offers the lowest error for all the sound environments:

\begin{equation}
mMAE = \frac{\sum_{i = 1}^4 MAE_i}{4}, 
\end{equation}

where the other experimental factors (method, $f_c$, $K$, $w_t$, $\beta$, threshold value(s)) are fixed.

\section{Results and discussion}\label{part:results}

Table \ref{tab:results} summarizes the lowest $mMAE$ errors according to the method (LP filter, SUP-NMF, SEM-NMF and Hard/Firm TI-NMF) and $\beta$ with the others corresponding modalities.

\begin{table*}[t]
\centering
\caption{Best $mMAE$ errors according to the experimental factors $\beta$ and \textit{method} (in bold letter, the lowest error).}
\begin{tabular}{@{}ccccccccc@{}}
\toprule
\textbf{method} & $f_c$ (kHz) & $\mathbf{\beta}$ & $\mathbf{K}$ & $\mathbf{w_t}$ (s) &  $\mathbf{t_h}$ &  $\mathbf{t_{f,1}}$ &  $\mathbf{t_{f,2}}$ & \textbf{$mMAE$} (dB) \\ \midrule
filter & 20 & - & - & - & - & - & - & 3.76 ($\pm$ 4.35) \\
filter & 0.5 & - & - & - & - & - & - & 2.14 ($\pm$ 1.83) \\ 
\hline \hline
SUP-NMF & - & 0 & 200 & 0.5 & - & - & - & 4.06 ($\pm$ 4.69) \\
SUP-NMF & - & 1 & 200 & 0.5 & - & - & - & 2.79 ($\pm$ 3.38) \\
SUP-NMF & - & 2 & 25 & 1 & - & - & - & 2.32  ($\pm$ 2.80) \\ 
\hline \hline
SEM-NMF & - & 0 & 200 & 1 & - & - & - & 2.05 ($\pm$ 0.70) \\
SEM-NMF & - & 1 & 200 & 1 & - & - & - & 1.94 ($\pm$ 0.38) \\
SEM-NMF & - & 2 & 200 & 1 & - & - & - & 2.39 ($\pm$ 1.23) \\ 
\hline \hline
Hard TI-NMF & - & 0 & 50 & all & 0.39 & - & - & 1.44 ($\pm$ 1.09)\\
Hard TI-NMF & - & 1 & 200 & 0.5 & 0.35 & - & - & 1.40 ($\pm$ 1.22)\\
Hard TI-NMF & - & 2 & 200 & 0.5 & 0.33 & - & - & 1.32 ($\pm$ 1.25)\\
\hline \hline
Firm TI-NMF & - & 0 & 50 & all & - & 0.45 & 0.32 & 1.42 ($\pm$ 1.11)\\
Firm TI-NMF & - & 1 & 50 & all & - & 0.41 & 0.30 & 1.37 ($\pm$ 1.20) \\
\textbf{Firm TI-NMF} & - & \textbf{2} & \textbf{200} & \textbf{1} & - & \textbf{0.38} & \textbf{0.29} & \textbf{1.31 ($\pm$ 1.26)}\\ 
\bottomrule
\end{tabular}
\label{tab:results}
\end{table*}

The LP filter with $f_c = 20$ kHz cut-off frequency is equivalent to considerate the sound level of the entire scene without specific distinction between the sound sources. The error is then important with a high standard deviation ($mMAE =$ 3.76 ($\pm$ 4.35) dB). The lowest error for a LP filter is obtained with $f_c$ = 500 Hz ($mMAE = $2.14 ($\pm$ 1.83) dB). It allows a balance between the discarded and remaining energy through the sound environments.

When considering all the sound scenes, SUP-NMF does not succeed to have a lower error than the 500 Hz LP filter for all the $\beta$ values. By adding the mobile part $\mathbf{W_r}$ in the dictionary, SEM-NMF with $\beta = 0$ and $\beta = 1$ allows a lower error than 500 Hz LP filter with a reduced standard deviation especially for $\beta = 1$ ($mMAE = 1.94$ ($\pm$ 0.38) dB).

Hard and Firm TI-NMF are the approaches with the lowest global error ($<$ 1.50 dB). The best result is obtained for firm TI-NMF ($MAE$ = 1.31 $\pm$ 1.26 dB) with $\beta = 2$, $K$ = 200, $w_t$ = 1 s and as threshold values $t_{f,1}$ = 0.38 and $t_{f,2}$ = 0.29. This combination of settings offers the most fitted method to be adapted to all the sound environments. One observes that for firm TI-NMF and for each $\beta$, the thresholds values $t_{f,1}$ and $t_{f,2}$ frame the hard threshold value $t_h$.

The elements located between these threshold values, included in $\mathbf{W}_{traffic}$, despite a highest distance with the traffic spectra, their impacts can be reduced by the weighting. Firm TI-NMF allows then to take into account more elements as traffic component than hard thresholding and then decrease the $mMAE$ error.

Furthermore, on the dictionary creation, except SEM-NMF where all the best methods according to $\beta$ use the same dictionary, no specific dictionary form through all the method, is used. SUP and SEM-NMF privilege a high number of element ($K$ = 200) with a fine description of the audio samples ($w_t \in \lbrace 0.5$,$1 \rbrace$ second). For TI-NMF, the composition of the initial dictionary is more diverse both on the $K$ value and on the finesse  of the description ($w_t \in \lbrace all$,$0.5 \rbrace$ second).

From these global results, the $MAE$ errors are compared to the LP filter and each method for the 4 sound environments, see Figure \ref{fig:mae_env}.

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{figures/mea_grafic_bar.pdf}
\caption{$MAE$ errors according each sound environment for the best combination of the LP filter ($f_c$ = 500 Hz), SUP-NMF ($\beta$ = 2, $K$ = 25, $w_t$ = 1 second), SEM-NMF ($\beta$ = 1, $K$ = 200, $w_t$ = 1 second) and firm TI-NMF ($\beta$ = 2, $K$ = 200, $w_t$ = 1 second, $t_{f,1} = 0.38$ and $t_{f,2}$ = 0.29)}
\label{fig:mae_env}
\end{figure}

Except SEM-NMF, all the methods show the same error evolution the same error evolution: a decrease of the error with the increase of the traffic predominance. SEM-NMF show an almost constant error for all 4 sound environments. The LP filter error is mainly important for environments where the traffic is less present. As this approach considers the remaining energy as the traffic component, no distinction can be made between the different sound sources not related to the sound sources of interest, here traffic. In the opposite, for noisy and very noisy environments, the performances of the LP filter are good ($MAE$ $<$ 1 dB). The errors are then due to a high deletion of the traffic energy by the filter while it becomes the main sound source.

Despite a fixed dictionary composed of traffic spectra, SUP-NMF fails to identified correctly the traffic component particularly for \textit{park} ($MAE$ = 6.42 dB) environments. With this method, as NMF minimizes the cost function, eq. \ref{eq:min-D-WH}, the dictionary's elements are used to model the other sound sources. In the opposite, for \textit{noisy} and \textit{very noisy} environments, SUP-NMF identifies correctly the traffic components ($MAE$ $<$ 0.6 dB) as it is the main source. 
In the case of SEM-NMF, adding the mobile dictionary, $\mathbf{W_r}$, makes it possible to include the other sound sources not present in the dictionary. If this behavior is advantageous for the \textit{park} environment ($MAE$ = 1.87 dB) where lot of different kind of sources are present, it is less advantageous for the rest of the environments where the traffic becomes predominant resulting in the highest errors. Indeed, this degree of freedom generates higher error as $\mathbf{W_r}$ is not constrained and is free to include traffic component in it, penalizing the traffic sound level estimation.

Besides having the lowest $mMAE$ error, TI-NMF presents the most performing results. The park environment is the only case where an other NMF approach out-performed significantly firm TI-NMF ($MAE$ = 3.05 dB). In this sound environment, the traffic dictionary is then composed, in average, of more than half the 200 elements in $\mathbf{W'}$ (136 elements). For the rest of the sound environments, firm TI-NMF has the lowest error. For very noisy environment the error is even very low ($MAE=$ 0.30 dB). In this case, in average, $\mathbf{W'}$ is composed of 198 traffic elements. This method out-performed SUP-NMF since, as $\mathbf{W_0}$ is updated, $\mathbf{W'}$ is then most suited to the sound scene than a fixed dictionary. The advantage to have an unique dictionary suited to each sound scene makes TI-NMF very performing when traffic is predominant while, by the thresholding, it limits the dictionary derivation when the traffic is more quiet.

\section{Conclusion}

A non-negative matrix factorization frame has been applied as a source separation tool to estimate the traffic sound level  from a corpus of urban sound scenes, which consists of scenes built artificially in which the level and position of each source is controled. In addition, the realism of the scenes has been verified thanks to a perceptual test. As a result, the proposed protocol mimics the outputs given by a deployed sensor network, but knowing the actual sound level of each source composing the sound mixture.

The results confirm the interest of this method on such sound environments. It easily takes into account the overlap between the multiple sound sources present in cities and is suited to monophonic sensor networks. Different NMF formula have been studied through the supervised and semi-supervised approach. On all the sound environments, these common approaches reveal to be not sufficiently efficient: supervised NMF approach, with its fixed dictionary, does not succeed to estimate correctly the traffic sound level especially when this sound source is quiet, while semi-supervised approach with the presence of a mobile part in the dictionary is the best estimator for \textit{park} environments but failed on heavily traffic scenes. Finally, the proposed approach, namely Thresholded Initialized NMF, achieved the lowest error with an initialized updated dictionary. By considering the elements the most similar to traffic spectra with a firm thresholding, it makes it possible to adapt specifically the dictionary to the sound scenes. Consequently, in the case where the emplacement or the sound environment of the sensors cannot be identified (for instance within a mobile measurement framework), firm TI-NMF, with $\beta = 2$, $K$ = 200, $w_t = 1$ second, $t_{f,1} = 0.38$ and $t_{f,2}$ = 0.29, seems the best generic framework, for the sound corpus studied. But if the sound environment can be identified through a prior analysis, or based on positioning data \cite{can2015noise,lavandier2016urban}, it seems possible to adapt the proposed work by selecting the most efficient approach (for instance SEM-NMF within parks), in order to further reduce the error in the estimated road traffic sound levels.
Further analyses are required to extend the proposed method to other sound sources, such as birds or voices sounds, by replacing or adding elements in the built dictionary. This would prove useful in the context of muti-source noise mapping that is  gaining interest \cite{aumond2017Probabilistic, aletta2015soundscape}. Finally, the selected parameters stand for the corpus of sound mixtures of this study. Further analyses on various corpus of sound scenes are needed to test the robustness of the method, and select the most relevant approaches for specific sound environments (predominance of water or industrial sounds, rural environments \dots).

For reproducible purposes, the experimental protocol and the programs developed under the Matlab software are available online. This study proposes also a realistic urban sound corpus that could be greatly appreciate for research communities dedicated to the detection, identification or source separation tasks.

\section*{References}
\bibliographystyle{elsarticle-num}
\bibliography{bibliographie_applied}

\end{document}

### 2018-02-27 paper/table/classificationSoundEnvTab.tex
add paper and experiment

\centering

\begin{tabular}{cc*{19}{l|}}
\cline{3-21}
\multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{19}{|c|}{\textbf{point}} \\
\hline
\multicolumn{1}{|l}{\textbf{day}} & \multicolumn{1}{|l|}{\textbf{journey}}   & 1                        & 2                        & 3                        & 4                        & 5                        & 6                        & 7                        & 8                        & 9                        & 10                       & 11                       & 12                       & 13                       & 14                       & 15                       & 16                       & 17                       & 18                       & 19                       \\ \hline
\multicolumn{1}{|c|}{\textbf{1}} & \textbf{EW} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{5AB25A} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{5AB25A} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{5AB25A} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{9A0000} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{5AB25A} \\
\hline
\multicolumn{1}{|c|}{\textbf{1}} & \textbf{WE} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{F56B00} &                          & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{9A0000} & \cellcolor[HTML]{9A0000} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} &  \\
 \hline
\multicolumn{1}{|c|}{\textbf{2}} & \textbf{EW} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{5AB25A} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{9A0000} & \cellcolor[HTML]{9A0000} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{5AB25A} \\
\hline
\multicolumn{1}{|c|}{\textbf{2}} & \textbf{WE} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{5AB25A} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{9A0000} & \cellcolor[HTML]{9A0000} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{9A0000} & \cellcolor[HTML]{5AB25A} \\ \hline
\end{tabular}
%
\vspace{0.5cm}

\begin{tabular}{|p{1cm}|l|p{0.001cm}|p{2cm}|l|p{0.001cm}|p{2cm}|l|p{0.001cm}|p{2.75cm}|l|p{0.001cm}|p{2cm}|l|}

\hhline{|-|-|~|-|-|~|-|-|~|-|-|~|-|-|}
park & {\cellcolor[HTML]{5AB25A}} & & quiet street & {\cellcolor[HTML]{FFCB2F}} & & noisy street & {\cellcolor[HTML]{F56B00}} & &  very noisy street & {\cellcolor[HTML]{9A0000}} & & unclassified & \\
\hhline{|-|-|~|-|-|~|-|-|~|-|-|~|-|-|}

\end{tabular}

### 2018-02-27 experiment/NMFThresholdGrafic.m

@@ -1,0 +1,190 @@
+function NMFThresholdGrafic(varargin)                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                          
+% Welcome to the main entry point of NMFThresholdGrafic                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                        
+% Please DO NOT modify this file unless you have a precise intent.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                             
+                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                               
+shortExperimentName = 'nmfthgr';                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                               
+[p, experimentName] = fileparts(mfilename('fullpath'));                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                        
[…]

### 2018-02-27 experiment/functions/MFCC/audspec.m

@@ -1,0 +1,40 @@
+function [aspectrum,wts,binfrqs] = audspec(pspectrum, sr, nfilts, fbtype, minfreq, maxfreq, sumpower, bwidth)
+%[aspectrum,wts] = audspec(pspectrum, sr, nfilts, fbtype, minfreq, maxfreq, sumpower, bwidth)
+%
+% perform critical band analysis (see PLP)
+% takes power spectrogram as input
+
+if nargin < 2;  sr = 16000;                          end
+if nargin < 3;  nfilts = ceil(hz2bark(sr/2))+1;      end
+if nargin < 4;  fbtype = 'bark';  end
+if nargin < 5;  minfreq = 0;    end
+if nargin < 6;  maxfreq = sr/2; end
+if nargin < 7;  sumpower = 1;   end
+if nargin < 8;  bwidth = 1.0;   end
+
+[nfreqs,~] = size(pspectrum);
+
+nfft = (nfreqs-1)*2;
+
+if strcmp(fbtype, 'bark')
+  [wts,binfrqs] = fft2barkmx(nfft, sr, nfilts, bwidth, minfreq, maxfreq);
+elseif strcmp(fbtype, 'mel')
+  [wts,binfrqs] = fft2melmx(nfft, sr, nfilts, bwidth, minfreq, maxfreq);
+elseif strcmp(fbtype, 'htkmel')
+  [wts,binfrqs] = fft2melmx(nfft, sr, nfilts, bwidth, minfreq, maxfreq, 1, 1);
+elseif strcmp(fbtype, 'fcmel')
+  [wts,binfrqs] = fft2melmx(nfft, sr, nfilts, bwidth, minfreq, maxfreq, 1, 0);
+else
+  disp(['fbtype ', fbtype, ' not recognized']);
+  error;
+end
+
+wts = wts(:, 1:nfreqs);
+
+% Integrate FFT bins into Mel bins, in abs or abs^2 domains:
+if (sumpower)
+  aspectrum = wts * pspectrum;
+else
+  aspectrum = (wts * sqrt(pspectrum)).^2;
+end
+

### 2018-02-27 experiment/functions/MFCC/ispecgram.m

@@ -1,0 +1,48 @@
+function x = ispecgram(d, ftsize, sr, win, nov)
+% X = ispecgram(D, F, SR, WIN, NOV)           Inverse specgram
+%    Overlap-add the inverse of the output of specgram
+%    ftsize is implied by sizeof d, sr is ignored, nov defaults to ftsize/2
+% dpwe 2005may16.  after istft
+
+[nspec,ncol] = size(d);
+
+if nargin < 2
+  ftsize = 2*(nspec-1);
+end
+if nargin < 3
+  % who cares?
+end
+if nargin < 4
+  win = ftsize;  % doesn't matter either - assume it added up OK
+end
+if nargin < 5
+  nov = ftsize/2;
+end
+
+hop = win - nov;
+
+if nspec ~= (ftsize/2)+1
+  error('number of rows should be fftsize/2+1')
+end
+
+xlen = ftsize + (ncol-1) * hop;
+x = zeros(xlen,1);
+
+halff = ftsize/2;   % midpoint of win
+
+% No reconstruction win (for now...)
+
+for c = 1:ncol
+  ft = d(:,c);
+  ft = [ft(1:(ftsize/2+1)); conj(ft([(ftsize/2):-1:2]))];
+
+  if max(imag(ifft(ft))) > 1e-5
+    disp('imag oflow');
+  end
+  
+  px = real(ifft(ft));  % no shift in specgram
+  
+  b = (c-1)*hop;
+  x(b+[1:ftsize]) = x(b+[1:ftsize]) + px;
+end;
+

### 2018-02-27 experiment/functions/NarrowToNthOctaveEXP.m

@@ -1,0 +1,94 @@
+function [sortedData,Fc,Flow,Fhigh] = NarrowToNthOctaveEXP(arrayOfNBCenterFreqs,arrayOfdBToConvert,n)
+
+df=arrayOfNBCenterFreqs(2)-arrayOfNBCenterFreqs(1);
+
+% %Use correctionFactor=1 if you did not window or you used ECF
+correctionFactor=1;
+% %Use this if you used a hanning window with ACF
+% ACF=(1/mean(window(@hann,2^11)));
+% ECF=(1/mean(window(@hann,2^11)));
+% correctionFactor=1/ACF;
+arrayOfdBToConvert=arrayOfdBToConvert*correctionFactor;
+
+%Determine Initial Center frequency (Fc)
+previous=1000*2^(1/n);
+lowFc=1000; %Start at 1000 Hz and find lowest Fc
+while (previous-lowFc) > df
+    previous=lowFc;
+    lowFc=lowFc/2^(1/n);
+end
+
+%Compute center frequencies
+ii=1;
+Fc(ii)=lowFc;
+while Fc(ii) <= max(arrayOfNBCenterFreqs) 
+    ii=ii+1;
+    Fc(ii)=2^(1/n)*Fc(ii-1);
+end
+
+%Compute high, low frequencies, and bandwidth (BW) from center frequencies
+    Flow=Fc/sqrt(2^(1/n));
+    Fhigh=Fc*sqrt(2^(1/n));
+    BW=Fhigh-Flow;
+
+%Sort data in frequency bins
+sortedData=zeros(length(Fc),size(arrayOfdBToConvert,2)); zeroFreqValue=0;
+for jj=1:size(arrayOfdBToConvert,1) %Do for each value given
+    for kk=1:length(Fc) %Check each Fc bin to see where value should be placed
+        if arrayOfNBCenterFreqs(jj)>=Flow(kk) && arrayOfNBCenterFreqs(jj)<Fhigh(kk) %Find place. In the rare case a given center freq = flow(ii)/fhigh(ii+1) then sum the value in flow
+            if sortedData(kk,:)==0 %if no values has been added to the band then set initial value
+                sortedData(kk,:)=arrayOfdBToConvert(jj,:);
+                N(kk)=1;
+            else %else sum in the value
+                sortedData(kk,:)=sqrt(sum(([sortedData(kk,:); arrayOfdBToConvert(jj,:)].^2))); %dB add value to that bin
+                N(kk)=N(kk)+1;
+            end
[…]

### 2018-02-27 experiment/functions/audio2SpectrogramEXP.m

@@ -1,0 +1,32 @@
+function [X,Xlinear,time,frequency,XScale] = audio2SpectrogramEXP(x,setting)
+%% [X,Xlinear,time,XScale,frequency] = audio2Spectrogram(x,setting)
+
+sr = setting.sr;
+nfft = setting.nfft;
+window = setting.nfft;
+noverlap = floor(window*setting.noverlap/100);
+
+if size(x,1)>size(x,2)
+    x = x';
+end
+
+X = cell(1,size(x,1));
+XScale = X;
+frequency = X;
+time = X;
+Xlinear = X;
+
+for ii = 1:size(x,1)
+    [spec,frequency{ii},time{ii}] = spectrogram(x(ii,:),window,noverlap,nfft,sr);
+    spec = spec./nfft;
+    Xlinear{ii} = abs(spec);
+    
+    if strcmp('thirdOctave',setting.domain)
+        [X{ii},XScale{ii},~,~] = NarrowToNthOctaveEXP(frequency{ii},Xlinear{ii},3);
+    elseif strcmp('mel',setting.domain)
+        [X{ii},XScale{ii}] = spectre2MelEXP(Xlinear{ii},setting.numberMel,setting.cutOffFreq,setting.sr);
+    else
+        X{ii} = Xlinear{ii};
+        XScale{ii} = frequency{ii};
+    end
+end

### 2018-02-27 experiment/functions/creationDictionaryEXP.m

@@ -1,0 +1,45 @@
+function [W,Xscale,indTraffic] = creationDictionaryEXP(file,indClass,setting)
+
+
+if min(size(file)) == 2        % stereo to mono
+    file = mean(file,2)';
+else
+    file = file';
+end
+
+spec = audio2SpectrogramEXP(file,setting);
+
+switch setting.frameSize
+    case 0 % W composed of fft of sounds get from the rms of the spectrogram
+        X = rms(spec{1},2);
+        
+    otherwise   % W composed of several temporal window get by the rms of the selected window
+        step = (setting.window*(setting.noverlap/100))/setting.sr;
+        time = step:step:size(spec{1},2)*step;
+        frameSize_temp = setting.frameSize*1e-3;    % size of the temporal window
+        timeEnd = floor(time(end));
+        if time(end) < 1
+           timeEnd = frameSize_temp; 
+        end
+        [~,idxBeg] = min(abs(repmat((0:frameSize_temp:timeEnd-frameSize_temp)',1,length(time))-repmat(time,timeEnd/frameSize_temp,1)),[],2);
+        [~,idxEnd] = min(abs(repmat((frameSize_temp:frameSize_temp:timeEnd)',1,length(time))-repmat(time,timeEnd/frameSize_temp,1)),[],2);
+
+        if length(idxEnd) < floor(time(end)/frameSize_temp)
+            idxBeg = [idxBeg; idxEnd(end)+1];
+            idxEnd = [idxEnd; length(time)];
+        end
+        X = zeros(size(spec{1},1),length(idxEnd));
+        
+        for ii = 1:length(idxBeg)
+            X(:,ii) = rms(spec{1}(:,idxBeg(ii):idxEnd(ii)-1),2);
+        end
+end
+
+Xscale = linspace(0,setting.sr/2,size(X,1));
+W = X./repmat(sum(X),size(X,1),1);
+
+if indClass == 1
+    indTraffic = ones(1,size(W,2));
+else
+    indTraffic = zeros(1,size(W,2));
+end

### 2018-02-27 experiment/functions/cutSpectrogramEXP.m

@@ -1,0 +1,23 @@
+function [soundMix] = cutSpectrogramEXP(W,setting)
+
+switch setting.domain
+    case 'mel'
+        sr = setting.sr;
+        numberMel = setting.numberMel; 
+
+        [W,~] = spectre2MelEXP(W,numberMel,setting.cutOffFreq,sr);
+        ind = numberMel;
+        
+    case 'thirdOctave'
+        frequency = linspace(0,setting.sr/2,size(W,1));
+        [W,~,~,~] = NarrowToNthOctaveEXP(frequency,W,3);
+        ind = size(W,1);
+        
+    otherwise
+        frequency = linspace(0,setting.sr/2,size(W,1));
+        [~,ind] = min(abs(setting.cutOffFreq-frequency));
+end
+
+soundMix.ind = ind;
+soundMix.seed = rng;
+soundMix.W = W;

### 2018-02-27 experiment/functions/estimationLpEXP.m

@@ -1,0 +1,40 @@
+function [Lp,Leq] = estimationLpEXP(Xinput,setting)
+
+if ~iscell(Xinput)
+    Xin{1} = Xinput;
+else
+    Xin = Xinput;
+end
+
+sr = setting.sr;
+window = setting.window;
+noverlap = setting.noverlap;
+step = (window*(noverlap/100))/sr;
+
+Lp = cell(1,size(Xin,2));
+Leq = zeros(1,size(Xin,2));
+
+for ii = 1:size(Xin,2)
+    X = Xin{ii};
+    X(2:end,:) = X(2:end,:)*2;
+    time =  step:step:size(X,2)*step;
+    
+    B = sqrt(abs(X).^2);
+    B(2:end,:) = sqrt(2)*B(2:end,:);
+    Bw = B/sqrt(1.5);           %hanning window effect "scaling effect" see https://www.ap.com/technical-library/fft-scaling-for-noise/
+
+    a = sqrt(sum(Bw.^2));
+    Leq(ii) = rms(a);
+    
+    T_tau = floor(time(end)/setting.temporalInterval); % nombre interval
+    N_tau = floor(length(time)/T_tau); % nombre echantilon par interval
+    ind_beg = linspace(0,(T_tau-1)*N_tau,T_tau);
+    ind_end = linspace(N_tau,T_tau*N_tau,T_tau);
+    
+    Lp{ii} = zeros(1,T_tau);
+    for jj = 1:T_tau
+        Lp{ii}(:,jj) = rms(a(ind_beg(jj)+1:ind_end(jj)));
+    end
+    Lp{ii}(isnan(Lp{ii}))=0;
+    Lp{ii}(isinf(Lp{ii}))=0;
+end

### 2018-02-27 experiment/functions/filtrePasseBasEXP.m

@@ -1,0 +1,22 @@
+function [LeqFiltre,LpFiltre] = filtrePasseBasEXP(X,setting)
+%% Y = filtrePasseBas(X,soundMix,varargin)
+% Fonction filtre pour un signal temporel x filtrer par un filtre de
+% Butterworth. Les paramètres se trouvent dans le structure soundMix :
+% - la fréquence de coupure, .cutOffFreq,
+% - l'ordre du filtre, .order.
+% Si ils ne sont pas présent : par défaut cutOffFreq = 1000 et order = 2.
+% Si on renseigne un troisième argument d'une valeur quelconque, on trace
+% la fonction filtre par la fonction fvtool.
+
+sr = setting.sr;
+nfft = setting.nfft;
+fc = setting.cutOffFreq;
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% FONCTION %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+frequency = linspace(0,sr/2,nfft/2+1);
+[~,ind] = min(abs(fc-frequency));
+X{1} = X{1}(1:ind,:);
+
+[LpFiltre,LeqFiltre] = estimationLpEXP(X,setting);
+

### 2018-02-27 experiment/functions/indiceEstimationEXP.m

@@ -1,0 +1,56 @@
+function indice = indiceEstimationEXP(x)
+
+
+indice.Leq = 20*log10(rms(x)/2e-5);
+
+%% LAeq %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+X = 2/sqrt(2)*abs(fft(x)/length(x));
+frequency = linspace(0,44100,length(X));
+
+X = X(frequency>20);
+frequency = frequency(frequency>20);
+X = X(frequency<20000);
+frequency = frequency(frequency<20000);
+
+[X_third,~] = NarrowToNthOctaveEXP(frequency,X,3);
+
+A_weight = [0.0012 0.0021 0.0037 0.0062 0.0098 0.0150 0.0222 0.0313 0.0428...
+    0.0570 0.0743 0.0935 0.1151 0.1384 0.1607 0.1824 0.2000 0.2143 0.2244 0.2296...
+    0.2323 0.2296 0.2244 0.2119 0.1977 0.1762 0.1500 0.1219 0.0935 0.0686]'.*1e-4;
+
+indice.LAeq = 20*log10(sqrt(sum((X_third(2:end)+A_weight).^2))/2e-5);
+
+%% L50, L10, L90 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+t = 0:1/44100:length(x)/44100-1/44100;
+dt = 0.125;
+
+T_tau = floor(t(end)/dt); % nombre interval
+N_tau = floor(length(t)/T_tau); % nombre echantilon par interval
+ind_beg = linspace(0,(T_tau-1)*N_tau,T_tau);
+ind_end = linspace(N_tau,T_tau*N_tau,T_tau);
+
+tt = linspace(0,t(end),T_tau)+(dt/2);
+Lp = zeros(1,T_tau);
+for jj = 1:T_tau
+    Lp(:,jj) = rms(x(ind_beg(jj)+1:ind_end(jj)));
+end
+
+Lp_sort = sort(20*log10(Lp/2e-5),'ascend');
+t_10 = (1-0.1)*tt(end);
+t_50 = 0.5*tt(end);
+t_90 = (1-0.9)*tt(end);
+
+[~,idx_90] = min(abs(tt-t_90));
[…]

### 2018-02-27 experiment/functions/mel2SpectreEXP.m

@@ -1,0 +1,13 @@
+function Y = mel2SpectreEXP(X,sr,nfft,cutOffFreq)
+
+if iscell(X)
+    Y = cell(size(X,1),size(X,2));
+    for ii = 1:size(X,1)
+        for jj = 1:size(X,2)
+            [Y{ii,jj}] = invaudspec(X{ii,jj}, sr, nfft, 'mel', 0, cutOffFreq-1, 0, 1);
+        end
+        
+    end
+else
+    [Y] = invaudspec(X, sr, nfft, 'mel', 0, cutOffFreq-1, 0, 1);
+end

### 2018-02-27 experiment/functions/nmf/NMFestimationEXP.m

@@ -1,0 +1,33 @@
+function [NMF,sequentialData] = NMFestimationEXP(V,dictionary,setting,sequentialData)
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% limit the spectrogram and the dictionary to cutOffFreq %%%%%%%%%%%%%%%%%
+[soundMix] = cutSpectrogramEXP(dictionary.W,setting);
+soundMix.indTraffic = dictionary.indTraffic;
+
+W0 = soundMix.W(1:soundMix.ind,:);      % initial dictionary
+V = V{1}(1:soundMix.ind,:);    % spectrogram to approximate
+
+if isempty(sequentialData)
+    iteration = setting.iteration;
+    W = W0;                             % first iteration W = W0
+    
+    binTemp = size(V,2);
+    rng(soundMix.seed)
+    H = rand(size(W,2),binTemp);    % H initiate randomly
+else
+    % continuing step of the sequential run
+    iteration = setting.iteration-sequentialData.numberIteration;
+    H = sequentialData.H;
+    W = sequentialData.W; 
+end
+% 
+NMF = algo_nmfUnsupervisedEXP(H,W,W0,V,iteration,setting);
+
+sequentialData.numberIteration = setting.iteration;
+sequentialData.H = NMF.H;
+sequentialData.W = NMF.W;
+
+
+
+

### 2018-02-27 experiment/functions/nmf/algo_nmfSemiSupervisedEXP.m

@@ -1,0 +1,183 @@
+function NMF = algo_nmfSemiSupervisedEXP(H,W,Y,Z,V,iteration,soundMix,setting)
+
+beta = setting.beta;
+sparsity = setting.sparsity;
+smoothness = setting.SM_weight;
+J = setting.SS_sizeWrand;
+K = size(W,2);
+
+if setting.SS_weight ~= 0
+    mu = setting.SS_weight;
+    betaM = setting.SS_betaM;
+    lambdaSS = setting.SS_lambda;
+else
+    mu = 0;
+end
+
+if ~isequal(beta,0) || ~isequal(beta,1) || ~isequal(beta,2)
+    if beta < 1
+        gamma = 1/(2-beta);
+    elseif beta >= 1 && beta <= 2
+        gamma = 1;
+    else
+        gamma = 1/(beta-1);
+    end
+end
+
+if smoothness~=0
+    smoothForm = setting.smoothnessForm;
+    
+    switch smoothForm
+        case 'all'
+            lambdaSM = [sum(W,1) sum(Y,1)];
+            smoothnessTot(1:K+J,1) = smoothness;
+        case 'traffic'
+            lambdaSM = sum(W,1);
+            smoothnessTot = zeros(K,1);
+            smoothnessTot(soundMix.indTraffic==1) = smoothness;
+    end
+else
+    smoothnessTot = 0;
+    lambdaSM = NaN;
+end
+
+Vap = W*H;
+Vtot = Vap+Y*Z;
+
+NMF = [];
+
+switch beta
+    case 2
+        for iter = 1:iteration
+            if mu == 0
+                Y = Y.*(((V*Z')./(Vtot*Z')).^gamma);
+            else
+                div = betadivEXP(Y(:)*ones(1,size(W,2)),repmat(W,size(Y,2),1),betaM);
+                Cm = exp((-1/lambdaSS)*sum(div));
+
+                num = lambdaSS*V*Z'+mu*Y.^(betaM-1)*Cm;
+                den = (lambdaSS*Vtot)*Z'+mu*Y.^(betaM-2).*Cm.*sum(W,2);
+                Y = Y.*(num./den);
+                Y(Y<2e-5) = 2e-5;
+            end
+            Y = Y./repmat(sum(Y),size(Y,1),1);
+            Vtot = Vap+Y*Z;
+            
+            if smoothness == 0 || strcmp(smoothForm,'traffic')
+                Z = Z.*(((Y'*V)./(Y'*Vtot)).^gamma);
[…]

### 2018-02-27 experiment/functions/nmf/algo_nmfSupervisedEXP.m

@@ -1,0 +1,100 @@
+function NMF = algo_nmfSupervisedEXP(H,W,V,iteration,soundMix,setting)
+
+sparsity = setting.sparsity; 
+smoothness = setting.SM_weight; 
+beta = setting.beta;
+Vap = W*H;
+
+NMF = [];
+
+if ~isequal(beta,0) || ~isequal(beta,1) || ~isequal(beta,2)
+    if beta < 1
+        gamma = 1/(2-beta);
+    elseif beta >= 1 && beta <= 2
+        gamma = 1;
+    else
+        gamma = 1/(beta-1);
+    end
+end
+
+if smoothness~=0
+    lambda = sum(W,1);
+    smoothForm = setting.smoothnessForm;
+    
+    switch smoothForm
+        case 'all'
+            smoothnessTot(1:size(W,2),1) = smoothness;
+        case 'traffic'
+            smoothnessTot = zeros(size(W,2),1);
+            smoothnessTot(soundMix.indTraffic==1) = smoothness;
+    end
+else
+    smoothnessTot = 0;
+    lambda = NaN;
+end
+
+switch beta
+    case 2
+        for iter = 1:iteration(end)
+            if smoothness==0
+                H = H.*(W'*V-sparsity)./(W'*Vap);
+                H(H<0)=0;
+            else
+                H = updateSmoothEssidEXP(H,W,V,Vap,beta,smoothnessTot,sparsity,lambda);
+            end
+            H(isnan(H)) = 0;
+            Vap = W*H;
+        end
+    case 1
+        for iter = 1:iteration
+            
+            if smoothness==0
+                H = H.*((W'*(V.*Vap.^(-1)))./(sparsity+W'*Vap.^(0)));
+            else
+                H = updateSmoothEssidEXP(H,W,V,Vap,beta,smoothnessTot,sparsity,lambda);
+            end
+            H(isnan(H)) = 0;
+            Vap = W*H;
+            
+        end
+    case 0
+        for iter = 1:iteration
+            
+            if smoothness==0
+                H = H.*((W'*(V.*Vap.^(-2)))./(sparsity+W'*Vap.^(-1))).^(1/2);
+            else
+                H = updateSmoothEssidEXP(H,W,V,Vap,beta,smoothnessTot,sparsity,lambda);
+            end
+            H(isnan(H)) = 0;
+            Vap = W*H;
+
+        end
+    otherwise
+        if beta < 1
+            for iter = 1:iteration
+                if smoothness==0
[…]

### 2018-02-27 experiment/functions/nmf/algo_nmfUnsupervisedEXP.m

@@ -1,0 +1,124 @@
+function [NMF] = algo_nmfUnsupervisedEXP(H,W,W0,V,iteration,setting)
+
+sparsity = setting.sparsity;
+beta = setting.beta;
+Vap = W*H;
+K = size(W,2); 
+
+if beta == 0
+    smoothness = setting.smoothnessBeta0;
+elseif beta == 1
+    smoothness = setting.smoothnessBeta1;
+elseif beta == 2
+    smoothness = setting.smoothnessBeta2;
+end
+
+if ~isequal(beta,0) || ~isequal(beta,1) || ~isequal(beta,2)
+    if beta < 1
+        gamma = 1/(2-beta);
+    elseif beta >= 1 && beta <= 2
+        gamma = 1;
+    else
+        gamma = 1/(beta-1);
+    end
+end
+
+cost = zeros(1,iteration);
+switch beta
+    case 2
+        for iter = 1:iteration
+            
+            W = W.*(V*H')./(Vap*H');
+            W = (1-setting.weightW)*W0+setting.weightW*W;
+            W(isnan(W)) = 0;
+            W = W./repmat(sum(W),size(W,1),1);
+            Vap = W*H;
+            
+            if smoothness == 0
+                H = H.*(W'*V-sparsity)./(W'*Vap);
+            else
+                H_1 = [zeros(K,1) H(:,1:end-1)];
+                H_2 = [H(:,2:end) zeros(K,1)];
+                H_12 = [zeros(K,1) H(:,2:end-1) zeros(K,1)];
+                H = H.*(W'*V-sparsity+2*smoothness.*(H_1+H_2))./(W'*Vap+2*smoothness.*(H+H_12));
+            end
+            
+            H(isnan(H)) = 0;
+
+            Vap = W*H;
+            cost(iter) = betadivEXP(V,Vap,beta);
+        end
+    case 1
+        for iter = 1:iteration
+            W = W .* ((V./Vap)*H')./(repmat(sum(H,2)',size(V,1),1));
+            W(isnan(W)) = 0;
+            W = W./repmat(sum(W),size(W,1),1);
+            Vap = W*H;
+
+            if smoothness == 0
+                H = H.*((W'*(V./Vap))./(sparsity+W'*Vap.^(0)));
+            else
+                H_1 = [zeros(K,1) H(:,1:end-1)];
+                H_2 = [H(:,2:end) zeros(K,1)];
+                H_12 = [zeros(K,1) H(:,2:end-1) zeros(K,1)];
[…]

### 2018-02-27 experiment/functions/nmf/betadivEXP.m

@@ -1,0 +1,41 @@
+function cost = betadivEXP(V,Vap,beta,varargin)
+
+if nargin > 3
+    H = varargin{1};
+    sparsity = varargin{2};
+    smoothness = varargin{3};
+else
+    H = [];
+    sparsity = 0;
+    smoothness = 0;
+end
+
+switch beta
+    case 2
+        cost = sum((V(:)-Vap(:)).^2)/2;
+    case 1
+        ind_0 = find(V(:)<=eps);
+        ind_1 = 1:length(V(:));
+        ind_1(ind_0) = [];
+        
+        cost = sum(V(ind_1).*log(V(ind_1)./Vap(ind_1)) - V(ind_1) + Vap(ind_1) ) + sum(Vap(ind_0));        
+    case 0
+        ind_0 = find(V(:)<=eps);
+        ind_1 = 1:length(V(:));
+        ind_1(ind_0) = [];
+        
+        cost = sum(V(ind_1)./Vap(ind_1)-log(V(ind_1)./Vap(ind_1)))-length(V(:))+sum(Vap(ind_0));
+
+    otherwise
+        cost = sum(V(:).^beta+(beta-1)*Vap(:).^beta-beta*V(:).*Vap(:).^(beta-1))/(beta*(beta-1));
+end
+
+if sparsity~=0
+    cost = cost+sparsity*sum(H(:));
+end
+
+if smoothness~=0
+    cost = cost+0.5*smoothness*sum(sum((H(:,1:end-1)-H(:,2:end)).^2));
+end
+    
+

### 2018-02-27 experiment/functions/nmf/betadivSemiEXP.m

@@ -1,0 +1,57 @@
+function cost = betadivSemiEXP(V,NMF,setting,varargin)
+
+
+beta = setting.beta;
+sparsity = setting.sparsity;
+smoothness = setting.SM_weight;
+betaM = setting.SS_betaM;
+mu = setting.SS_weight;
+lambda = setting.SS_sensitivity;
+
+H = NMF.H;
+W = NMF.W;
+Y = NMF.Y;
+Vtot = NMF.Vtot;
+
+term = zeros(1,size(W,2));
+if mu == 0
+    cost = betadivEXP(V,Vtot,beta,H,sparsity,smoothness);
+else
+
+    div = betadivEXP(Y(:)*ones(1,size(W,2)),repmat(W,size(Y,2),1),betaM);
+	Cm = exp((-1/lambda)*sum(div));
+
+    switch beta
+        case 2
+            cost = sum((V(:)-Vtot(:)).^2)/2+mu*sum(term);
+        case 1
+            ind_0 = find(V(:)<=eps);
+            ind_1 = 1:length(V(:));
+            ind_1(ind_0) = [];
+            cost = sum(V(ind_1).*log(V(ind_1)./Vtot(ind_1))-V(ind_1)+Vtot(ind_1))+sum(Vtot(ind_0))...
+                +mu*Cm;
+        case 0
+            ind_0 = find(V(:)<=eps);
+            ind_1 = 1:length(V(:));
+            ind_1(ind_0) = [];
+
+            cost = sum(V(ind_1)./Vtot(ind_1)-log(V(ind_1)./Vtot(ind_1)))-length(V(:))+sum(Vtot(ind_0))...
+                +mu*Cm;
+
+        otherwise
+            cost = sum(V(:).^beta + (beta-1)*Vtot(:).^beta - beta*V(:).*Vtot(:).^(beta-1) )/(beta*(beta-1))...
+                +mu*Cm;
+    end
+end
+if sparsity~=0
+    cost = cost+sparsity*sum(H(:));
+end
+
+if smoothness~=0
+    if strcmp(setting.smoothnessForm,'all')
+        basis = [H;NMF.Z];
+        cost = cost+0.5*smoothness*sum((sum((basis(:,1:end-1)-basis(:,2:end)).^2)));
+    else
+        cost = cost+0.5*smoothness*sum((sum((H(:,1:end-1)-H(:,2:end)).^2)));
+    end
+end

### 2018-02-27 experiment/functions/nmf/extractionLevelEXP.m

@@ -1,0 +1,58 @@
+function NMF = extractionLevelEXP(NMF,soundMix,V,beta,setting,varargin)
+
+sparsity = setting.sparsity; 
+smoothness = setting.smoothness; 
+nmfType = setting.nmfType; 
+
+Vap = NMF.Vap;
+H = NMF.H;
+
+if strcmp(setting.domain,'mel')
+    Y = mel2SpectreEXP([{V} {Vap}],setting.sr,setting.nfft,setting.cutOffFreq);
+    V = Y{1};
+    Vap = Y{2};
+end
+
+switch nmfType
+    case 'supervised'
+        costTemp = betadivEXP(V,Vap,beta,H,sparsity,smoothness);
+        dictionary = soundMix.W;
+        activator = H;
+        
+    case 'semi-supervised'
+        Y = varargin{1};
+        Z = varargin{2};
+        size_Wrand = setting.SS_sizeWrand; 
+        constraint = setting.SS_constraint; 
+
+        YFull = [Y; zeros(size(soundMix.Wfull,1)-size(Y,1),size_Wrand)];
+        dictionary = [soundMix.W YFull];
+        activator = [H; Z];
+
+        switch constraint
+            case 0
+                costTemp = betadivEXP(V,Vap,beta,H,sparsity,smoothness);
+
+            otherwise
+                betaM = setting.SS_betaM;
+                lambda = setting.SS_lambda;
+
+                div = betadivEXP(repmat(soundMix.Wcut,size(Y,2),1),Y(:)*ones(1,size(soundMix.Wcut,2)),betaM);
+                Cp = exp((-1/lambda)*sum(div));
+                costTemp = betadivEXP(V,Vap,beta,H,sparsity,smoothness)+Cp;
+        end
+        NMF.Y = Y;
+        NMF.Z = Z;
+end
+
+Vap_full = dictionary*activator;
+
+[Lp,Leq] = estimationLpEXP([{Vap} {Vap_full}],setting);
+
+NMF.H = H;
+NMF.cost = costTemp;
+NMF.LeqGlobalEstimate = Leq;
+NMF.LpGlobalEstimate = Lp;
+
+NMF = separationTrafficEXP(NMF,soundMix,setting);
+NMF = orderfields(NMF);

### 2018-02-27 experiment/functions/nmf/nmfSupervise_MMEXP.m

@@ -1,0 +1,19 @@
+function [NMF] = nmfSupervise_MMEXP(setting,sequentialData,soundMix,beta)
+%% function [NMFOUT] = beta_nmfSupervise_MM(V, W, HInit, beta, NMF, setting, sample, ind)
+
+W = soundMix.Wcut;
+V = soundMix.V_scene;
+
+binTemp = size(V,2);
+
+if isempty(sequentialData)
+    iteration = setting.iteration;
+    rng(soundMix.seed)
+    HInit = rand(size(W,2),binTemp);
+else
+    % continuing step of the sequential run
+    iteration = setting.iteration-sequentialData.numberIteration;
+    HInit = sequentialData.H;
+end
+
+NMF = algo_nmfSupervised(HInit,W,V,beta,iteration,soundMix,setting);

### 2018-02-27 experiment/functions/nmf/separationTrafficEXP.m

@@ -1,0 +1,26 @@
+function [NMF] = separationTrafficEXP(NMF,soundMix,setting)
+%% [NMF] = separationActivateur(NMF,soundMix,ind)
+% séparation des activateurs pour chaque classe du dictionnaire
+%
+
+Htraffic = NMF.H(soundMix.indTraffic==1,:);
+Wtraffic = soundMix.W(:,soundMix.indTraffic==1);
+
+Vtraffic_full = Wtraffic*Htraffic;
+
+if strcmp(setting.domain,'mel')
+    
+    Vtraffic_full = mel2SpectreEXP(Vtraffic_full,setting.sr,setting.nfft);
+    [Lp,Leq] = estimationLpEXP(Vtraffic_full,setting);
+    
+elseif strcmp(setting.domain,'thirdOctave')
+    
+    [Lp,Leq] = estimationLpEXP(Vtraffic_full,setting);
+else
+    
+    Vtraffic_cut = Wtraffic(1:soundMix.ind,:)*Htraffic;
+    [Lp,Leq] = estimationLpEXP([{Vtraffic_full} {Vtraffic_cut}],setting);
+end
+
+NMF.LeqTrafficEstimate = Leq;
+NMF.LpTrafficEstimate = Lp;

### 2018-02-27 experiment/functions/nmf/updateSmoothEssidEXP.m

@@ -1,0 +1,80 @@
+function [H] = updateSmoothEssidEXP(H,W,V,V_ap,beta,smoothness,sparsity,lambda)
+
+N = size(H,2);
+
+switch beta
+    case 0
+        Ht = H;
+        a = 2*smoothness.*(lambda'.^2);
+        c = 0;
+        d = -W'*(V./(V_ap.^2)).*(Ht.^2);
+        
+        % n = 1
+        b = W'*(V_ap(:,1).^(-1))-smoothness.*(lambda'.^2).*(H(:,2));
+        H(:,1) = resolutionThirdOrder(a,b,c,d(:,1),smoothness);
+        
+        % n = 2:N-1
+        for n = 2:N-1
+            b = W'*(V_ap(:,n).^(-1))-smoothness.*(lambda'.^2).*(H(:,n-1)+H(:,n+1));
+            H(:,n) = resolutionThirdOrder(a,b,c,d(:,n),smoothness);
+        end
+        
+        % n = N
+        b = W'*(V_ap(:,N).^(-1))-smoothness.*(lambda'.^2).*(H(:,N-1));
+        H(:,N) = resolutionThirdOrder(a,b,c,d(:,N),smoothness);
+        
+        
+    case 1
+        if any(smoothness==0)
+            smoothness(smoothness==0)=1e-12;
+        end
+        
+        Ht = H;
+        a_n = 2*smoothness.*(lambda'.^2);
+        a_1 = a_n/2;
+        a_N = a_n/2;
+        bt = sparsity+lambda';
+        c = Ht.*(W'*(V./V_ap));
+        
+        % n = 1
+        b(:,1) = bt-smoothness.*(lambda'.^2).*H(:,2);
+        H(:,1) = (sqrt(b(:,1).^2+4*a_1.*c(:,1))-b(:,1))./a_1;
+        
+        % n = 2:N-1
+        b(:,2:N-1) = repmat(bt,1,size(H,2)-2)-repmat(a_n,1,size(H,2)-2).*(H(:,1:N-2)+H(:,3:N));
+        for n = 2:N-1
+            H(:,n) = (sqrt(b(:,n).^2+4*a_n.*c(:,n))-b(:,n))./(2*a_n);
+        end
+        
+        % n = N
+        b(:,N) = bt-smoothness.*(lambda'.^2).*(H(:,N-1));
+        H(:,N) = (sqrt(b(:,N).^2+4*a_N.*c(:,N))-b(:,N))./a_N;
+        
+        H(H<0)=0;
[…]

### 2018-02-27 experiment/functions/resolutionThirdOrder.m

@@ -1,0 +1,57 @@
+function X = resolutionThirdOrder(a,b,c,d,alphaCoeff)
+
+K = length(a);
+
+if length(alphaCoeff)==1
+    alphaCoeff(1:K,1) = alphaCoeff;
+end
+
+alphaNonNull = find(alphaCoeff~=0);
+alphaNull = find(alphaCoeff==0);
+
+%% alphaNonNull ~= 0
+if ~isempty(alphaNonNull)
+    p = c./a(alphaNonNull)-(b(alphaNonNull).^2)./(3*(a(alphaNonNull).^2));
+    q = (2*(b(alphaNonNull).^3))./(27*(a(alphaNonNull).^3))+...
+        d(alphaNonNull)./a(alphaNonNull)-b(alphaNonNull)*c./(3*a(alphaNonNull).^2);
+
+    delta = q.^2+4*((p./3).^3);
+
+    pos = find(delta > 0);
+    nul = find(delta == 0);
+    neg = find(delta <0);
+
+    X = zeros(K,1);
+    if ~isempty(pos)
+        C1 = nthroot((-q(pos)+sqrt(delta(pos)))/2,3);
+        C2 = nthroot((-q(pos)-sqrt(delta(pos)))/2,3);
+
+        X(pos) = C1+C2-(b(pos)./(3*a(pos)));
+    end
+    if ~isempty(nul)
+
+        Xnul(:,1) = (3*q(nul))./p(nul)-(b(nul)./(3*a(nul)));
+        Xnul(:,2) = -(3*q(nul))./(2*p(nul))-(b(nul)./(3*a(nul)));
+        Xnul(:,3) = Xnul(:,2);
+        X(nul) = min(abs(Xnul),[],2);
+    end
+    if~isempty(neg)
+        deltaLoop = abs(delta(neg));
+        jj = -0.5+1j*(sqrt(3)/2);
+        c1 = (-q(neg)+1j*sqrt(deltaLoop))/2;
+        c2 = (-q(neg)-1j*sqrt(deltaLoop))/2;
+        C1 = nthroot(abs(c1),3).*exp(1j*angle(c1)/3);
+        C2 = nthroot(abs(c2),3).*exp(1j*angle(c2)/3);
+
+        Xneg(:,1) = C1+C2-(b(neg)./(3*a(neg)));
+        Xneg(:,2) = jj*C1+(jj)^2*C2-(b(neg)./(3*a(neg)));
[…]

### 2018-02-27 experiment/functions/rmseEXP.m

@@ -1,0 +1,24 @@
+function errorRMSE = rmseEXP(X,Y,varargin)
+%% rmse(X,Y,DIMENSION)
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% CONDITIONS %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+if size(X,2)>size(Y,2)
+    X = X(:,1:size(Y,2));
+elseif size(X,2)<size(Y,2)
+    Y = Y(:,1:size(X,2));
+end
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% FONCTION %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% if ~isfinite(X)
+%     X(ii) = 0;
+% elseif ~isfinite(Y)
+%     Y(ii) = 0;
+% end
+if size(X,1)~=1
+    errorRMSE = sqrt(mean(bsxfun(@minus,Y,X).^2));
+else
+    errorRMSE = sqrt(sum((X-Y).^2));
+end

### 2018-02-27 experiment/functions/spectre2MelEXP.m

@@ -1,0 +1,13 @@
+function [Y,binMel,wts] = spectre2MelEXP(X,numberMel,cutOffFreq,sr)
+
+if iscell(X)
+    Y = cell(size(X,1),size(X,2));
+    for ii = 1:size(X,1)
+        for jj = 1:size(X,2)
+            [Y{ii,jj},wts,binMel] = audspec(X{ii,jj}, sr, numberMel,'mel', 20, cutOffFreq, 0, 1);
+        end
+    end
+else 
+    [Y,wts,binMel] = audspec(X, sr, numberMel,'mel', 20, cutOffFreq, 0, 1);
+
+end

### 2018-02-27 experiment/nmfthgr1dictionary.m

@@ -1,0 +1,79 @@
+function [config, store, obs] = nmfthgr1dictionary(config, setting, data)            
+% nmfth1dictionary DICTIONARY step of the expLanes experiment NMFThreshold         
+%    [config, store, obs] = nmfth1dictionary(config, setting, data)                
+%      - config : expLanes configuration state                                     
+%      - setting   : set of factors to be evaluated                                
+%      - data   : processing data stored during the previous step                  
+%      -- store  : processing data to be saved for the other steps                 
+%      -- obs    : observations to be saved for analysis                           
+                                                                                   
+% Copyright: <userName>                                                            
+% Date: 01-Dec-2017                                                                
+                                                                                   
+% Set behavior for debug mode                                                      
+if nargin==0, NMFThresholdGrafic('do', 1, 'mask',{0 0 0 0 1}); return; else store=[]; obs=[]; end
+                                                                                   
+if strcmp(setting.type,'nmf')
+    K = setting.numberElement;
+    inputPath = config.inputPath;
+    
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    %% NAME & NUMBER SAMPLE EXTRACTION %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    
+    cityCarSamples = dir(strcat(inputPath,'dictionary',filesep,'*cityCar*'));
+    roadCarSamples = dir(strcat(inputPath,'dictionary',filesep,'*roadCar*'));
+    stopCarSamples = dir(strcat(inputPath,'dictionary',filesep,'*stopCar*'));
+    sample = [cityCarSamples; roadCarSamples; stopCarSamples];
+    indiceClass = ones(1,length(sample));
+    
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    %% AUDIO SAMPLE EXTRACTION %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    
[…]

### 2018-02-27 experiment/nmfthgr2estimation.m

@@ -1,0 +1,104 @@
+function [config, store, obs] = nmfthgr2estimation(config, setting, data)            
+% nmfth2estimation ESTIMATION step of the expLanes experiment NMFThreshold         
+%    [config, store, obs] = nmfth2estimation(config, setting, data)                
+%      - config : expLanes configuration state                                     
+%      - setting   : set of factors to be evaluated                                
+%      - data   : processing data stored during the previous step                  
+%      -- store  : processing data to be saved for the other steps                 
+%      -- obs    : observations to be saved for analysis                           
+                                                                                   
+% Copyright: <userName>                                                            
+% Date: 01-Dec-2017                                                                
+                                                                                   
+% Set behavior for debug mode                                                      
+if nargin==0, NMFThresholdGrafic('do', 2, 'mask', {}); return; else store=[]; obs=[]; end
+
+dataset = setting.dataset;
+type = setting.type;
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% DICTIONARY %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+if strcmp(type,'nmf')
+    dictionary.W = data.dictionary.W;
+    dictionary.frequency = data.dictionary.frequency;
+    dictionary.indTraffic = data.dictionary.indTraffic;
+else
+    dictionary = [];
+end
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% BASELINE GLOBALE ERROR + FILTRE %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+creationSceneDir = strcat(config.inputPath, dataset, filesep, setting.gType, filesep);
+
+files = dir(strcat(creationSceneDir,filesep,'*.wav'));
+globalName = cell(1,length(files)/2);
+ind = 1;
+
+for ii = 1:length(files)
+    if strfind(files(ii).name,'traffic')
[…]

### 2018-02-27 experiment/nmfthgr3threshold.m

@@ -1,0 +1,122 @@
+function [config, store, obs] = nmfthgr3threshold(config, setting, data)
+% nmfth3threshold THRESHOLD step of the expLanes experiment NMFThreshold
+%    [config, store, obs] = nmfth3threshold(config, setting, data)
+%      - config : expLanes configuration state
+%      - setting   : set of factors to be evaluated
+%      - data   : processing data stored during the previous step
+%      -- store  : processing data to be saved for the other steps
+%      -- obs    : observations to be saved for analysis
+
+% Copyright: <userName>
+% Date: 01-Dec-2017
+
+% Set behavior for debug mode
+if nargin==0, NMFThresholdGrafic('do', 3, 'mask',{}); return; else store=[]; obs=[]; end
+
+nmf =  data.nmf;
+levels = cell(1,length(data.nmf));
+
+if strcmp(setting.type,'nmf') && strcmp(setting.nmfType,'thresholded')
+    distanceMethod = setting.distanceMethod;
+    methodThreshold = setting.methodThreshold;
+    
+    valueThresholdCheck = 1;
+    switch  methodThreshold
+        case 'firm'
+            threshold(1) = setting.thresholdFirmHigh;
+            threshold(2) = setting.thresholdFirmLow;
+            if threshold(1) < threshold(2)
+                valueThresholdCheck = 0;
+            end
+        otherwise
+            threshold = setting.threshold;
+    end
+    
+    [soundMix] = cutSpectrogramEXP(nmf{1}.W0,setting);
+    W0 = soundMix.W(1:soundMix.ind,:);
+    [F,K] = size(W0);
+    
+    if valueThresholdCheck == 1
+        for ii = 1:length(nmf)
+            W = nmf{ii}.W(1:F,:);
+            H = nmf{ii}.H;
+            
+            switch distanceMethod
+                case 'cosine'
+                    dist = sum(W.*W0)./(sqrt(sum(W.^2,1)).*sqrt(sum(W0.^2,1)));
+                    dist(isnan(dist)) = 0;
+                    [~, order] = sort(dist,'descend');
+                    
+                case 'none'
+                    order = 1:K;
+            end
+            Wn = W(:,order);
+            Hn = H(order,:);
+            
+            switch distanceMethod
+                case 'none'
+                    Wtraffic = Wn;
+                    Htraffic = Hn;
+                    
[…]

### 2018-02-27 experiment/nmfthgr4metric.m

@@ -1,0 +1,94 @@
+function [config, store, obs] = nmfthgr4metric(config, setting, data)                
+% nmfth4metric METRIC step of the expLanes experiment NMFThreshold                 
+%    [config, store, obs] = nmfth4metric(config, setting, data)                    
+%      - config : expLanes configuration state                                     
+%      - setting   : set of factors to be evaluated                                
+%      - data   : processing data stored during the previous step                  
+%      -- store  : processing data to be saved for the other steps                 
+%      -- obs    : obsrmservations to be saved for analysis                           
+                                                                                   
+% Copyright: <userName>                                                            
+% Date: 01-Dec-2017                                                                
+                                                                                   
+% Set behavior for debug mode                                                      
+if nargin==0, NMFThresholdGrafic('do', 4, 'mask', {2 0 0 0 1, ...
+	3 0 6 0 3, 0 3 11 1 1, 1 1 10 1 3, 1 0 42:47 20:40}); return; else store=[]; obs=[]; end                                                                                   
+
+levels = data.levels;
+pref = setting.p0;
+
+LeqTrafficEstimate = zeros(length(levels),1);
+LeqTrafficEstimatedB = zeros(length(levels),1);
+LeqTrafficExact = zeros(length(levels),1);
+LeqTrafficExactdB = zeros(length(levels),1);
+LeqGlobal = zeros(length(levels),1);
+rmsedB = zeros(length(levels),1);
+cost = zeros(length(levels),1);
+rmse = zeros(length(levels),1);
+nrmse = zeros(length(levels),1);
+
+%% EXTRACTION LEVELS
+for ii = 1:length(levels)
+    LpTrafficExact = levels{ii}.LpTraffic;
+    LpTrafficExact(LpTrafficExact<2e-5) = pref;
+    
[…]

### 2018-02-27 experiment/nmfthgrInit.m

@@ -1,0 +1,10 @@
+function [config, store] = nmfthgrInit(config)                            
+% nmfthgrInit INITIALIZATION of the expLanes experiment NMFThresholdGrafic
+%    [config, store] = nmfthgrInit(config)                                
+%      - config : expLanes configuration state                            
+%      -- store  : processing data to be saved for the other steps        
+                                                                          
+% Copyright: <userName>                                                   
+% Date: 21-Dec-2017                                                       
+                                                                          
+if nargin==0, NMFThresholdGrafic(); return; else store=[];  end           

### 2018-02-27 experiment/nmfthgrReport.m

@@ -1,0 +1,116 @@
+function config = nmfthgrReport(config)                                
+% nmfthgrReport REPORTING of the expLanes experiment NMFThresholdGrafic
+%    config = nmfthgrInitReport(config)                                
+%       config : expLanes configuration state                          
+                                                                       
+% Copyright: <userName>                                                
+% Date: 21-Dec-2017                                                    
+                                                                       
+if nargin==0, NMFThresholdGrafic('report', 'r'); return; end         
+                                                                       
+%% mask 
+% 2 0 0 0 1, ...
+% 	3 0 6 0 3, 0 3 11 1 1, 1 1 10 1 [1 3], 1 25:45 31:47 20:40
+% 2 0 4 3 1, ...
+% 	3 0 6 0 3, 3 3 11 1 1, 1 6:7 10 1 [1 3], 1 1:31 1:30 1:25
+% 1 0 0 0 0, 0 0 0 0 3
+%% hard thresholding results
+
+% config = expExpose(config, 't', 'mask', {{1 0 0 0 0, 0 0 0 0 3},{2 0 0 0 1, ...
+% 	3 0 6 0 3, 0 0 11 1 1, 1 1 10 1 1, 1 25:45}},...
+%     'obs','mae','integrate',{'gType'},'negativeRank',-2,'sort',1);
+
+% config = expExpose(config, 't', 'mask', {2 0 [3 4] [1 2] 1, ...
+% 	3 0 6 0 3, 0 3 11 1 1, 1 1 10 1 1, 1},...
+%     'obs','mae','integrate',{'gType'},'negativeRank',-2,'sort',1);
+% 
+% config = expExpose(config, 'b', 'mask', {1 0 0 0 0, 0 0 1 0 3},...
+%     'obs','mae','negativeRank',-2,'sort',1,'save','bar_filter_grafic');
+
+%%% optimum
+% for env = 1:4
[…]

### 2018-02-27 paper/gloaguenAppliedNMF.tex
update missing references

@@
 Three variants of NMF are tested, namely supervised, semi-supervised, and threshold initialized NMF.
-While the semi-supervised approach is the most appropriated for park environments, threshold initialized NMF, which is developed especially for this research purpose, appears to be the best generic approach, allowing road traffic noise level estimation with mean average errors of 1.3 dB over the tested corpus of sound scenes.
+While the semi-supervised approach is the most appropriated for park environments, threshold initialized NMF, which is developed especially for this research purpose, appears to be the best generic approach, allowing road traffic noise level estimation with average errors of 1.3 dB over the tested corpus of sound scenes.
 
@@
 Therefore, noise measurements are increasingly used in addition to simulation to describe urban noise environments.
-Several measurement set-ups have been proposed in the last years, including  mobile measurements with high quality microphones \cite{manvell2004sadmam, can2014measurement}, participative sensing through dedicated smartphone applications \cite{picaut:halshs-01565214, maisonneuve2009noisetube}, or the development of fixed-sensor networks.
+Several measurement set-ups have been proposed in the last years, including  mobile measurements with high quality microphones \cite{manvell2004sadmam, can2014measurement}, participative sensing through dedicated smartphone applications \cite{picaut:halshs-01565214, ventura2017evaluation}, or the development of fixed-sensor networks.
 In this latter case, the sensor networks can be based either on high-quality sensors as in \cite{mietlicki2012innovative, maijala2018environmental}, or low-cost sensors as in the DYNAMAP project \cite{dynamap_2016} or the CENSE project \cite{picaut2017characterization}.
@@
 Wei et al.
-modify the acoustical parameters of the simulation thanks to noise measurements.
-Mallet et al.
-call for data assimilation techniques between models and measurements to reduce the uncertainty of the produced noise maps.
+\cite{wei_dynamic_2016} modify the acoustical parameters of the simulation thanks to noise measurements, while  Mallet et al.
+\cite{ventura2017estimation} call for data assimilation techniques between models and measurements to reduce the uncertainty of the produced noise maps.
 However, these works make the implicit assumption that the noise measurements consist mainly of road traffic.
@@
 A lot of applications can be found for musical \cite{smaragdis_non-negative_2003, benetos2006musical} and speech \cite{wilson_speech_2008, mysore2011non} contents.
-Dedicated to sound separation, Immani and Kasa\"i used NMF in a two steps sound separation with the help of time variant gain features.
-A first study \cite{} has been conducted, in which diverse NMF formula, namely the supervised, the semi-supervised, and the threshold initialized NMF, have been applied on a large set of simulated sound scenes that which mixed traffic component with specific urban sounds at calibrated sound levels.
+Dedicated to sound separation, Immani and Kasa\"i \cite{satoshi_innami_nmf-based_2012} used NMF in a two steps sound separation with the help of time variant gain features.
+A first study [article JASM] has been conducted, in which diverse NMF formula, namely the supervised, the semi-supervised, and the threshold initialized NMF, have been applied on a large set of simulated sound scenes that which mixed traffic component with specific urban sounds at calibrated sound levels.
 The study proved the interest of NMF for urban sound environments and compared the benefits of each approach.
@@
 
+\subsection{NMF experimental factors}
+
+NMF is performed for 3 $\beta$-divergences: $\beta$ = 2 (euclidean distance), $\beta$ = 1 (Kullback-Leibler divergence) and $\beta = 0$ (Itakura-Sa\"ito divergence).
+The spectrogram $\mathbf{V}$ and the dictionary $\mathbf{W}$ are displayed in a logarithmic scale through a third band octave representation that reduces the high frequency predominance where the traffic component is absent.
+In addition, as the number of frequency bins is reduced ($F$ = 29), the computation time is reduced too.
+400 iterations are performed to get a stabilized results.
+For SEM-NMF, the number of elements in $\mathbf{W_r}$ is fixed at $J = 2$.
+For hard thresholding, the threshold value, $t_h$ is defined between 0.30 and 0.60.
+For firm thresholding, the high threshold, $t_{f,1}$, is fixed between 0.45 and 0.60, the low threshold, $t_{f,2}$, is fixed between 0.10 and 0.45.
+All the intervals are defined with a 0.01 step.
+The cases $t_{f,1} < t_{f,2}$ are, naturally, discarded, see part \ref{part:NMF_TI}.
+Each unique association of modalities between each experimental factor forms a setting.
+For the filter estimator, 24 settings are computed (4 $\times$ 6).
+For SUP and SEM-NMF, 240 settings are computed (4 $\times$ 2 $\times$ (2 $\times$ 4 + 1 $\times$ 2) $\times$ 3).
+Finally, for TI-NMF, the number of settings is much higher (115 320) due to the multiple possible combinations between the threshold values (4 $\times$ 1 $\times$ (2 $\times$ 4 + 1 $\times$ 2) $\times$ 3 $\times$ (31 + (20 $\times$ 31 + $\sum_{i = 15}^{30} i $))).
+The summarize of the experimental factors and their different modalities is displayed on Tables \ref{tab:experimental_factorsFilter} and \ref{tab:experimental_factorsNMF}.
+
+The approximated traffic spectrograms $\mathbf{\tilde{V}}_{traffic}$ are obtained after 400 iterations. The estimated traffic sound level in dB, $\tilde{L}_{p, traffic}$, is deducted,
+
+\begin{equation}
+\tilde{L}_{p, traffic} = 20\log_{10}\frac{p_{rms}}{p_0}, 
+\end{equation}
+
+where $p_0$ is the reference sound pressure, $p_0 = 2\times 10^{-5}$ Pa. For each setting, $M$ traffic sound levels, corresponding to the $M$ scenes of each sound environment, are then calculated.
+
 \begin{table*}[t]
@@
 \end{tabularx}
-
 \label{tab:experimental_factorsNMF}
 \end{table*}
-
-\subsection{NMF experimental factors}
-
-NMF is performed for 3 $\beta$-divergences: $\beta$ = 2 (euclidean distance), $\beta$ = 1 (Kullback-Leibler divergence) and $\beta = 0$ (Itakura-Sa\"ito divergence).
-The spectrogram $\mathbf{V}$ and the dictionary $\mathbf{W}$ are displayed in a logarithmic scale through a third band octave representation that reduces the high frequency predominance where the traffic component is absent.
-In addition, as the number of frequency bins is reduced ($F$ = 29), the computation time is reduced too.
-400 iterations are performed to get a stabilized results.
-For SEM-NMF, the number of elements in $\mathbf{W_r}$ is fixed at $J = 2$.
-For hard thresholding, the threshold value, $t_h$ is defined between 0.30 and 0.60.
-For firm thresholding, the high threshold, $t_{f,1}$, is fixed between 0.45 and 0.60, the low threshold, $t_{f,2}$, is fixed between 0.10 and 0.45.
-All the intervals are defined with a 0.01 step.
-The cases $t_{f,1} < t_{f,2}$ are, naturally, discarded, see part \ref{part:NMF_TI}.
-Each unique association of modalities between each experimental factor forms a setting.
-For the filter estimator, 24 settings are computed (4 $\times$ 6).
-For SUP and SEM-NMF, 240 settings are computed (4 $\times$ 2 $\times$ (2 $\times$ 4 + 1 $\times$ 2) $\times$ 3).
-Finally, for TI-NMF, the number of settings is much higher (115 320) due to the multiple possible combinations between the threshold values (4 $\times$ 1 $\times$ (2 $\times$ 4 + 1 $\times$ 2) $\times$ 3 $\times$ (31 + (20 $\times$ 31 + $\sum_{i = 15}^{30} i $))).
-The summarize of the experimental factors and their different modalities is displayed on Tables \ref{tab:experimental_factorsFilter} and \ref{tab:experimental_factorsNMF}.
-
-The approximated traffic spectrograms $\mathbf{\tilde{V}}_{traffic}$ are obtained after 400 iterations. The estimated traffic sound level in dB, $\tilde{L}_{p, traffic}$, is deducted,
-
-\begin{equation}
-\tilde{L}_{p, traffic} = 20\log_{10}\frac{p_{rms}}{p_0}, 
-\end{equation}
-
-where $p_0$ is the reference sound pressure, $p_0 = 2\times 10^{-5}$ Pa. For each setting, $M$ traffic sound levels, corresponding to the $M$ scenes of each sound environment, are then calculated.
 

### 2018-03-01 paper/gloaguenAppliedNMF.tex
minor improvements

@@
 
-$d_{\beta}$ is mainly chosen as a $\beta$-divergence \cite{fevotte_algorithms_2011}, a sub-classes belonging to the Bregman divergences \cite{hennequin_beta-divergence_2011} which include 3 specific divergence and distance calculations: the Euclidean distance (eq.
+$d_{\beta}(x\vert y)$ is mainly chosen as a $\beta$-divergence \cite{fevotte_algorithms_2011}, a sub-classes belonging to the Bregman divergences \cite{hennequin_beta-divergence_2011} which include 3 specific divergence and distance calculations: the Euclidean distance (eq.
 \ref{eq:def_distEUC}), the Kullback-Leibler divergence (eq.
@@
 As the main source that interests us can be known \textit{a priori}, semi-supervised NMF (SEM-NMF) \cite{lee_semi-supervised_2010} is considered to better take into account the interfering sound sources.
-This approach proposes to decompose $\mathbf{W}_{F \times (K+J)}$ with two distinctive matrices: $\mathbf{W} = \left[ \mathbf{W_s}~\mathbf{W_r} \right]$ where $\mathbf{W_s}_{F \times K}$ is a fixed part of $\mathbf{W}$ composed of fixed audio spectra and $ \mathbf{W_r}_{F \times J}$, a mobile part which is updated, see eq.
+This approach proposes to decompose $\mathbf{W}_{F \times (K+J)}$ with two distinctive matrices: $\mathbf{W} = \left[ \mathbf{W_s}~\mathbf{W_r} \right]$ where $\mathbf{W_s}_{F \times K}$ is a fixed part of $\mathbf{W}$ composed of audio spectra and $ \mathbf{W_r}_{F \times J}$, a mobile part which is updated, see eq.
 \ref{eq:W_r_SS}.
@@
 \begin{equation}
-\mathbf{V} = \mathbf{W_s H_s}+\mathbf{W_r H_r}.
+\mathbf{V} \approx \mathbf{WH} = \mathbf{W_s H_s}+\mathbf{W_r H_r}.
 \end{equation}
@@
 This approach as the advantage, with the add of the mobile part $\mathbf{W_r}$, to bring more flexibility and then to be more adaptive to the different urban sound environments.
-Applications of Sem-NMF can be found for musical \cite{weninger2012supervised, kitamura_music_2014} and speech content \cite{joder2012real, mysore2011non}.
+Applications of SEM-NMF can be found for musical \cite{weninger2012supervised, kitamura_music_2014} and speech content \cite{joder2012real, mysore2011non}.
 
@@
 
-The urban sound scenes are taken from 76 recordings from 2 to 5 min, achieved in the \nth{13} district of Paris (France) at 19 different locations \footnote{Recordings were made as part of the Grafic project funded by Ademe}, which cover four various sound environments (Figure \ref{fig:map_grafic}).
+The urban sound scenes are taken from 76 recordings from 2 to 5 min, achieved in the \nth{13} district of Paris (France) at 19 different locations \footnote{Recordings were made as part of the Grafic project funded by Ademe}, which cover four various sound environments (Figure \ref{fig:map_grafic}), see Figure \ref{fig:map_grafic}.
 A complete description of the experimental protocol can be found in \cite{aumond_modelling_2017}.
@@
 The recordings are listened and categorized within four different sound environments, as proposed in \cite{can_describing_2015}: park (P, 8 audio files), quiet street (Q, 35 audio files), noisy street (N, 23 audio files) and very noisy street (vN, 8 audio files).
-Then, each audio file is annotated, noting the start and end time of each sound event along with its sound class (described here below).
+Then, each audio file is annotated, noting the start and end time of each sound event along with its sound class.
 The aim of the annotation phase is next to transcribe the recordings, in order to obtain simulated sound scenes with the same distribution of sound events as the recordings and therefore as close as possible to the realistic scenes.
@@
 A $K$-mean clustering algorithm is applied to reduce these dimensions to $K \in \lbrace$25, 50, 100, 200$\rbrace$ in order to avoid redundant information and decrease the computation time.
-The obtained $K$ clusters compose then the $K$ elements of $\mathbf{W}$.
+The obtained $K$ clusters compose then the elements of $\mathbf{W}$.
 Furthermore, the case where each audio generates one spectrum from its spectrogram is added ($w_t = all$).
@@
 	\rowcolor[HTML]{C0C0C0} 
-  \textbf{method} & Sup NMF & Sem NMF & TI NMF & 3\\
+  \textbf{method} & SUP NMF & SEM NMF & TI NMF & 3\\
 \end{tabularx}
@@
 The error is then important with a high standard deviation ($mMAE =$ 3.76 ($\pm$ 4.35) dB).
-The lowest error for a LP filter is obtained with $f_c$ = 500 Hz ($mMAE = $2.14 ($\pm$ 1.83) dB).
+The lowest error for a LP filter is obtained with $f_c$ = 500 Hz ($mMAE =$ 2.14 ($\pm$ 1.83) dB).
 It allows a balance between the discarded and remaining energy through the sound environments.
@@
 SUP and SEM-NMF privilege a high number of element ($K$ = 200) with a fine description of the audio samples ($w_t \in \lbrace 0.5$,$1 \rbrace$ second).
-For TI-NMF, the composition of the initial dictionary is more diverse both on the $K$ value and on the finesse  of the description ($w_t \in \lbrace all$,$0.5 \rbrace$ second).
+For TI-NMF, the composition of the initial dictionary is more diverse both on the $K$ value and on the finesse  of the description ($w_t \in \lbrace all$,$0.5$, $1 \rbrace$ second).
 
@@
 The LP filter error is mainly important for environments where the traffic is less present.
-As this approach considers the remaining energy as the traffic component, no distinction can be made between the different sound sources not related to the sound sources of interest, here traffic.
+As this approach considers the remaining energy as the traffic component, no distinction is made between the different sound sources not related to the traffic component.
 In the opposite, for noisy and very noisy environments, the performances of the LP filter are good ($MAE$ $<$ 1 dB).
@@
 In this case, in average, $\mathbf{W'}$ is composed of 198 traffic elements.
-This method out-performed SUP-NMF since, as $\mathbf{W_0}$ is updated, $\mathbf{W'}$ is then most suited to the sound scene than a fixed dictionary.
-The advantage to have an unique dictionary suited to each sound scene makes TI-NMF very performing when traffic is predominant while, by the thresholding, it limits the dictionary derivation when the traffic is more quiet.
+This method out-performed SUP-NMF since, as $\mathbf{W_0}$ is updated, the final dictionary $\mathbf{W'}$ is then most suited to the sound scene than a fixed dictionary.
+The advantage to have an unique dictionary suited to each sound scene makes TI-NMF very performing when traffic is predominant while, by the thresholding, it limits the dictionary derivation when the traffic is less present.
 
@@
 
-A non-negative matrix factorization frame has been applied as a source separation tool to estimate the traffic sound level  from a corpus of urban sound scenes, which consists of scenes built artificially in which the level and position of each source is controled.
+A non-negative matrix factorization frame has been applied as a source separation tool to estimate the traffic sound level  from a corpus of urban sound scenes artifically built.
 In addition, the realism of the scenes has been verified thanks to a perceptual test.
-As a result, the proposed protocol mimics the outputs given by a deployed sensor network, but knowing the actual sound level of each source composing the sound mixture.
+The proposed protocol generate audio sample similar than the outputs of a deployed sensor networks with tha advantages of the simulation process (sound level and position of each source controlled and known).
 

### 2018-03-02 paper/gloaguenAppliedNMF.tex
remove firm tresholding parts

@@
 Next, the similarities are sorted in descending order.
-The extraction of traffic elements in $\mathbf{W'}$ is carried out by two thresholding methods.
-For both, the traffic elements are estimated by weighting $\mathbf{W'}$ according to $D_{\theta}\left(\mathbf{w_0} \Vert \mathbf{w'} \right)$ and a threshold value such as :
+The extraction of traffic elements in $\mathbf{W'}$ is carried out by a hard thresholding method \cite{donoho1994threshold}.
+It consists in weigthing in a binary way the traffic elements $\mathbf{W'}$ according to $D_{\theta}\left(\mathbf{w_0} \Vert \mathbf{w'} \right)$ and a threshold value such as :
 
@@
 \end{equation}
-
-The first method is the hard thresholding \cite{donoho1995noising}. It selects the traffic elements from $\mathbf{W'}$ in a binary way by comparing their similarities to a fixed threshold $t_h$
-
+with 
 \begin{subequations}\label{eq:WH-TIupdateHard}
@@
 
-A second approach is based on firm thresholding \cite{gao1997waveshrink} which considers two thresholds, $t_{f,1}$ and $t_{f,2}$ with $t_{f,1}$ $>$ $t_{f,2}$. The weight $\alpha_k$ is then defined as
-
-\begin{subequations}\label{eq:WH-TIupdateFirm}
-\begin{numcases}{\alpha_k =}
-	1 & {\scriptsize iff $D_{\theta}\left(\mathbf{w}_{0} \Vert \mathbf{w'} \right) > t_{f,1}$} \\
-    0 & {\scriptsize iff $D_{\theta}\left(\mathbf{w}_{0} \Vert \mathbf{w'} \right) \leq t_{f,2}$} \\
-	\mathcal{N} \left(D_{\theta}(\mathbf{w_0}\Vert \mathbf{w'}) \right)& else.
-\end{numcases}
-\end{subequations}
-
-% \dfrac{d_{\beta}(\mathbf{w_0}\Vert \mathbf{w'}) - \min(d_{\beta}(\mathbf{w_0}\Vert \mathbf{w'}))}{\max(d_{\beta}(\mathbf{w_0}\Vert \mathbf{w'}))-\min(d_{\beta}(\mathbf{w_0}\Vert \mathbf{w'}))}
-where $\mathcal{N}\left(D_{\theta}(\mathbf{w_0}\Vert \mathbf{w'}) \right)$ is a normalization between 0 and 1 of the similarity. 
-The most similar elements to $\mathbf{W}_0$ are then keeping as they are whereas the elements whose the similarities are located between the two thresholds are then weighted.
 The allure of $D_{\theta}\left(\mathbf{W_0} \Vert \mathbf{W'} \right)$ and $\alpha$ are displayed in Figure \ref{fig:example_threshold}.\\
@@
        \includegraphics[width=0.9\linewidth]{./figures/thresh.pdf}
-    \caption{Evolution of the $\alpha$ weight according to $D_{\theta}\left(\mathbf{W_0} \Vert \mathbf{W'} \right)$ for hard thresholding with $t_h$ = 0.6 (in dashed line) and firm thresholding with $t_{f,1}$ = 0.67 and $t_{f,2}$ = 0.40 (in full line)}
+    \caption{Evolution of the $\alpha$ weight according to $D_{\theta}\left(\mathbf{W_0} \Vert \mathbf{W'} \right)$ for hard thresholding with $t_h$ = 0.6 (in dashed line)}
     \label{fig:example_threshold}
@@
 For SEM-NMF, the number of elements in $\mathbf{W_r}$ is fixed at $J = 2$.
-For hard thresholding, the threshold value, $t_h$ is defined between 0.30 and 0.60.
-For firm thresholding, the high threshold, $t_{f,1}$, is fixed between 0.45 and 0.60, the low threshold, $t_{f,2}$, is fixed between 0.10 and 0.45.
-All the intervals are defined with a 0.01 step.
-The cases $t_{f,1} < t_{f,2}$ are, naturally, discarded, see part \ref{part:NMF_TI}.
+For hard thresholding, the threshold value, $t_h$ is defined between 0.30 and 0.60 with a 0.01 step.
 Each unique association of modalities between each experimental factor forms a setting.
@@
 For SUP and SEM-NMF, 240 settings are computed (4 $\times$ 2 $\times$ (2 $\times$ 4 + 1 $\times$ 2) $\times$ 3).
-Finally, for TI-NMF, the number of settings is much higher (115 320) due to the multiple possible combinations between the threshold values (4 $\times$ 1 $\times$ (2 $\times$ 4 + 1 $\times$ 2) $\times$ 3 $\times$ (31 + (20 $\times$ 31 + $\sum_{i = 15}^{30} i $))).
+Finally, for TI-NMF, the number of settings is much higher (3720) due to the multiple possible combinations between the threshold values (4 $\times$ 1 $\times$ (2 $\times$ 4 + 1 $\times$ 2) $\times$ 3 $\times$ 31).
 The summarize of the experimental factors and their different modalities is displayed on Tables \ref{tab:experimental_factorsFilter} and \ref{tab:experimental_factorsNMF}.
@@
    hard threshold $\mathbf{t_h}$ & from 0.30 to 0.60 with a 0.01 step & 31\\
-\end{tabularx}
-
-\begin{tabularx}{17.5cm}{L{3cm}@{}C{12cm}@{}C{2cm}@{}}
-   high firm threshold $\mathbf{t_{f,1}}$ & from 0.30 to 0.60 with a 0.01 step & 31\\
-\end{tabularx}
-
-\begin{tabularx}{17.5cm}{L{3cm}@{}C{12cm}@{}C{2cm}@{}}
-	\rowcolor[HTML]{C0C0C0} 
-      low firm threshold $\mathbf{t_{f,2}}$ & from 0.10 to 0.45 with a 0.01 step & 36\\
    \bottomrule
@@
 
-where the other experimental factors (method, $f_c$, $K$, $w_t$, $\beta$, threshold value(s)) are fixed.
+where the other experimental factors (method, $f_c$, $K$, $w_t$, $\beta$, threshold value $t_h$) are fixed.
 
@@
 
-Table \ref{tab:results} summarizes the lowest $mMAE$ errors according to the method (LP filter, SUP-NMF, SEM-NMF and Hard/Firm TI-NMF) and $\beta$ with the others corresponding modalities.
+Table \ref{tab:results} summarizes the lowest $mMAE$ errors according to the method (LP filter, SUP-NMF, SEM-NMF and TI-NMF) and $\beta$ with the others corresponding modalities.
 
@@
 \caption{Best $mMAE$ errors according to the experimental factors $\beta$ and \textit{method} (in bold letter, the lowest error).}
-\begin{tabular}{@{}ccccccccc@{}}
+\begin{tabular}{@{}ccccccc@{}}
 \toprule
-\textbf{method} & $f_c$ (kHz) & $\mathbf{\beta}$ & $\mathbf{K}$ & $\mathbf{w_t}$ (s) &  $\mathbf{t_h}$ &  $\mathbf{t_{f,1}}$ &  $\mathbf{t_{f,2}}$ & \textbf{$mMAE$} (dB) \\ \midrule
-filter & 20 & - & - & - & - & - & - & 3.76 ($\pm$ 4.35) \\
-filter & 0.5 & - & - & - & - & - & - & 2.14 ($\pm$ 1.83) \\ 
+\textbf{method} & $f_c$ (kHz) & $\mathbf{\beta}$ & $\mathbf{K}$ & $\mathbf{w_t}$ (s) &  $\mathbf{t_h}$ &  \textbf{$mMAE$} (dB) \\ \midrule
+filter & 20 & - & - & - & - & 3.76 ($\pm$ 4.35) \\
+filter & 0.5 & - & - & - & - & 2.14 ($\pm$ 1.83) \\ 
 \hline \hline
-SUP-NMF & - & 0 & 200 & 0.5 & - & - & - & 4.06 ($\pm$ 4.69) \\
-SUP-NMF & - & 1 & 200 & 0.5 & - & - & - & 2.79 ($\pm$ 3.38) \\
-SUP-NMF & - & 2 & 25 & 1 & - & - & - & 2.32  ($\pm$ 2.80) \\ 
+SUP-NMF & - & 0 & 200 & 0.5 & - & 4.06 ($\pm$ 4.69) \\
+SUP-NMF & - & 1 & 200 & 0.5 & - & 2.79 ($\pm$ 3.38) \\
+SUP-NMF & - & 2 & 25 & 1 & - & 2.32  ($\pm$ 2.80) \\ 
 \hline \hline
-SEM-NMF & - & 0 & 200 & 1 & - & - & - & 2.05 ($\pm$ 0.70) \\
-SEM-NMF & - & 1 & 200 & 1 & - & - & - & 1.94 ($\pm$ 0.38) \\
-SEM-NMF & - & 2 & 200 & 1 & - & - & - & 2.39 ($\pm$ 1.23) \\ 
+SEM-NMF & - & 0 & 200 & 1 & - & 2.05 ($\pm$ 0.70) \\
+SEM-NMF & - & 1 & 200 & 1 & - & 1.94 ($\pm$ 0.38) \\
+SEM-NMF & - & 2 & 200 & 1 & - & 2.39 ($\pm$ 1.23) \\ 
 \hline \hline
-Hard TI-NMF & - & 0 & 50 & all & 0.39 & - & - & 1.44 ($\pm$ 1.09)\\
-Hard TI-NMF & - & 1 & 200 & 0.5 & 0.35 & - & - & 1.40 ($\pm$ 1.22)\\
-Hard TI-NMF & - & 2 & 200 & 0.5 & 0.33 & - & - & 1.32 ($\pm$ 1.25)\\
-\hline \hline
-Firm TI-NMF & - & 0 & 50 & all & - & 0.45 & 0.32 & 1.42 ($\pm$ 1.11)\\
-Firm TI-NMF & - & 1 & 50 & all & - & 0.41 & 0.30 & 1.37 ($\pm$ 1.20) \\
-\textbf{Firm TI-NMF} & - & \textbf{2} & \textbf{200} & \textbf{1} & - & \textbf{0.38} & \textbf{0.29} & \textbf{1.31 ($\pm$ 1.26)}\\ 
+TI-NMF & - & 0 & 50 & all & 0.39 & 1.44 ($\pm$ 1.09)\\
+TI-NMF & - & 1 & 200 & 0.5 & 0.35 & 1.40 ($\pm$ 1.22)\\
+\textbf{TI-NMF }& - & \textbf{2} & \textbf{200} & \textbf{0.5} & \textbf{0.33} & \textbf{1.32 ($\pm$ 1.25)}\\
 \bottomrule
@@
 
-Hard and Firm TI-NMF are the approaches with the lowest global error ($<$ 1.50 dB).
-The best result is obtained for firm TI-NMF ($MAE$ = 1.31 $\pm$ 1.26 dB) with $\beta = 2$, $K$ = 200, $w_t$ = 1 s and as threshold values $t_{f,1}$ = 0.38 and $t_{f,2}$ = 0.29.
+TI-NMF is the approach with the lowest global error ($<$ 1.50 dB).
+The best result is obtained for TI-NMF ($MAE$ = 1.31 $\pm$ 1.26 dB) with $\beta = 2$, $K$ = 200, $w_t$ = 0.5 s and as threshold value $t_h$ = 0.33.
 This combination of settings offers the most fitted method to be adapted to all the sound environments.
-One observes that for firm TI-NMF and for each $\beta$, the thresholds values $t_{f,1}$ and $t_{f,2}$ frame the hard threshold value $t_h$.
-
-The elements located between these threshold values, included in $\mathbf{W}_{traffic}$, despite a highest distance with the traffic spectra, their impacts can be reduced by the weighting.
-Firm TI-NMF allows then to take into account more elements as traffic component than hard thresholding and then decrease the $mMAE$ error.
-
 Furthermore, on the dictionary creation, except SEM-NMF where all the best methods according to $\beta$ use the same dictionary, no specific dictionary form through all the method, is used.
@@
 \includegraphics[width=\linewidth]{figures/mea_grafic_bar.pdf}
-\caption{$MAE$ errors according each sound environment for the best combination of the LP filter ($f_c$ = 500 Hz), SUP-NMF ($\beta$ = 2, $K$ = 25, $w_t$ = 1 second), SEM-NMF ($\beta$ = 1, $K$ = 200, $w_t$ = 1 second) and firm TI-NMF ($\beta$ = 2, $K$ = 200, $w_t$ = 1 second, $t_{f,1} = 0.38$ and $t_{f,2}$ = 0.29)}
+\caption{$MAE$ errors according each sound environment for the best combination of the LP filter ($f_c$ = 500 Hz), SUP-NMF ($\beta$ = 2, $K$ = 25, $w_t$ = 1 second), SEM-NMF ($\beta$ = 1, $K$ = 200, $w_t$ = 1 second) and TI-NMF ($\beta$ = 2, $K$ = 200, $w_t$ = 0.5 second, $t_{h} = 0.33$)}
 \label{fig:mae_env}
@@
 In the case of SEM-NMF, adding the mobile dictionary, $\mathbf{W_r}$, makes it possible to include the other sound sources not present in the dictionary.
-If this behavior is advantageous for the \textit{park} environment ($MAE$ = 1.87 dB) where lot of different kind of sources are present, it is less advantageous for the rest of the environments where the traffic becomes predominant resulting in the highest errors.
+If this behavior is advantageous for the \textit{park} environment ($MAE$ = 2.10 dB) where lot of different kind of sources are present, it is less advantageous for the rest of the environments where the traffic becomes predominant resulting in the highest errors.
 Indeed, this degree of freedom generates higher error as $\mathbf{W_r}$ is not constrained and is free to include traffic component in it, penalizing the traffic sound level estimation.
@@
 Besides having the lowest $mMAE$ error, TI-NMF presents the most performing results.
-The park environment is the only case where an other NMF approach out-performed significantly firm TI-NMF ($MAE$ = 3.05 dB).
+The park environment is the only case where an other NMF approach out-performed significantly TI-NMF ($MAE$ = 3.05 dB).
 In this sound environment, the traffic dictionary is then composed, in average, of more than half the 200 elements in $\mathbf{W'}$ (136 elements).
-For the rest of the sound environments, firm TI-NMF has the lowest error.
-For very noisy environment the error is even very low ($MAE=$ 0.30 dB).
+For the rest of the sound environments, TI-NMF has the lowest error.
+For very noisy environment the error is even very low ($MAE=$ 0.31 dB).
 In this case, in average, $\mathbf{W'}$ is composed of 198 traffic elements.
@@
 Finally, the proposed approach, namely Thresholded Initialized NMF, achieved the lowest error with an initialized updated dictionary.
-By considering the elements the most similar to traffic spectra with a firm thresholding, it makes it possible to adapt specifically the dictionary to the sound scenes.
-Consequently, in the case where the emplacement or the sound environment of the sensors cannot be identified (for instance within a mobile measurement framework), firm TI-NMF, with $\beta = 2$, $K$ = 200, $w_t = 1$ second, $t_{f,1} = 0.38$ and $t_{f,2}$ = 0.29, seems the best generic framework, for the sound corpus studied.
+By considering the elements the most similar to traffic spectra with a hard thresholding, it makes it possible to adapt specifically the dictionary to the sound scenes.
+Consequently, in the case where the emplacement or the sound environment of the sensors cannot be identified (for instance within a mobile measurement framework), TI-NMF, with $\beta = 2$, $K$ = 200, $w_t = 1$ second, $t_{h} = 0.33$, seems the best generic framework, for the sound corpus studied.
 But if the sound environment can be identified through a prior analysis, or based on positioning data \cite{can2015noise,lavandier2016urban}, it seems possible to adapt the proposed work by selecting the most efficient approach (for instance SEM-NMF within parks), in order to further reduce the error in the estimated road traffic sound levels.

### 2018-03-05 paper/gloaguenAppliedNMF.tex
minor updates

@@
 Two of the 76 recordings are rejected for the analysis because the audio files were corrupted, resulting in 74 valid audio files assumed as representative of the variety of sound environments.
-The recordings are listened and categorized within four different sound environments, as proposed in \cite{can_describing_2015}: park (P, 8 audio files), quiet street (Q, 35 audio files), noisy street (N, 23 audio files) and very noisy street (vN, 8 audio files).
+The recordings are listened and categorized within four different sound environments, as proposed in \cite{can_describing_2015}: park (P, 8 audio files), quiet street (Q, 33 audio files), noisy street (N, 24 audio files) and very noisy street (vN, 9 audio files).
 Then, each audio file is annotated, noting the start and end time of each sound event along with its sound class.
@@
 An illustrative example on a 3 seconds sample is displayed in Figure \ref{fig:example_dictionary}.
-From the 53 audio files, we obtain respectively for $w_t\in$ $\lbrace$0.5, 1$\rbrace$ second, 2218 and 1109 elements.
-A $K$-mean clustering algorithm is applied to reduce these dimensions to $K \in \lbrace$25, 50, 100, 200$\rbrace$ in order to avoid redundant information and decrease the computation time.
-The obtained $K$ clusters compose then the elements of $\mathbf{W}$.
-Furthermore, the case where each audio generates one spectrum from its spectrogram is added ($w_t = all$).
+From the 53 audio files, we obtain respectively for $\mathbf{w_t}\in$ $\lbrace$0.5, 1$\rbrace$ second, 2218 and 1109 elements.
+A $\mathbf{K}$-mean clustering algorithm is applied to reduce these dimensions to $\mathbf{K} \in \lbrace$25, 50, 100, 200$\rbrace$ in order to avoid redundant information and decrease the computation time.
+The obtained $\mathbf{K}$ clusters compose then the elements of $\mathbf{W}$.
+Furthermore, the case where each audio generates one spectrum from its spectrogram is added ($\mathbf{w_t} = all$).
 Here, by the added approach, the dictionary elements are based on the spectral envelops of the audio samples.
-For this case, having 53 audio samples, the number of elements in $\mathbf{W}$ with the $K$-mean clustering algorithm is reduced to $K \in \lbrace$25, 50$\rbrace$.
+For this case, having 53 audio samples, the number of elements in $\mathbf{W}$ with the $\mathbf{K}$-mean clustering algorithm is reduced to $\mathbf{K} \in \lbrace$25, 50$\rbrace$.
 
@@
 Finally, each basis of $\mathbf{W}$ is normalized as $ \Vert \mathbf{w}_k \Vert = 1$ where $\Vert \bullet \Vert$ is the $\ell_1$ norm.
-Table \ref{tab:experimental_factorsNMF} summarizes the different modalities of the two experimental factors ($K$ and $w_t$).
+Table \ref{tab:experimental_factorsNMF} summarizes the different modalities of the two experimental factors ($\mathbf{K}$ and $\mathbf{w_t}$).
 
@@
 For SEM-NMF, the number of elements in $\mathbf{W_r}$ is fixed at $J = 2$.
-For hard thresholding, the threshold value, $t_h$ is defined between 0.30 and 0.60 with a 0.01 step.
+For hard thresholding, the threshold value, $\mathbf{t_h}$ is defined between 0.30 and 0.60 with a 0.01 step.
 Each unique association of modalities between each experimental factor forms a setting.
@@
 
-where the other experimental factors (method, $f_c$, $K$, $w_t$, $\beta$, threshold value $t_h$) are fixed.
+where the other experimental factors (method, $f_c$, $\mathbf{K}$, $\mathbf{w_t}$, $\beta$, threshold value $\mathbf{t_h}$) are fixed.
 
@@
 \hline \hline
-TI-NMF & - & 0 & 50 & all & 0.39 & 1.44 ($\pm$ 1.09)\\
-TI-NMF & - & 1 & 200 & 0.5 & 0.35 & 1.40 ($\pm$ 1.22)\\
-\textbf{TI-NMF }& - & \textbf{2} & \textbf{200} & \textbf{0.5} & \textbf{0.33} & \textbf{1.32 ($\pm$ 1.25)}\\
+TI-NMF & - & 0 & 25 & 1 & 0.39 & 1.42 ($\pm$ 0.89)\\
+TI-NMF & - & 1 & 100 & 1 & 0.35 & 1.38 ($\pm$ 0.88)\\
+\textbf{TI-NMF }& - & \textbf{2} & \textbf{200} & \textbf{0.5} & \textbf{0.32} & \textbf{1.24 ($\pm$ 1.24)}\\
 \bottomrule
@@
 TI-NMF is the approach with the lowest global error ($<$ 1.50 dB).
-The best result is obtained for TI-NMF ($MAE$ = 1.31 $\pm$ 1.26 dB) with $\beta = 2$, $K$ = 200, $w_t$ = 0.5 s and as threshold value $t_h$ = 0.33.
+The best result is obtained for TI-NMF ($MAE$ = 1.24 ($\pm$ 1.24) dB) with $\beta = 2$, $\mathbf{K}$ = 200, $\mathbf{w_t}$ = 0.5 s and as threshold value $\mathbf{t_h}$ = 0.32.
 This combination of settings offers the most fitted method to be adapted to all the sound environments.
 Furthermore, on the dictionary creation, except SEM-NMF where all the best methods according to $\beta$ use the same dictionary, no specific dictionary form through all the method, is used.
-SUP and SEM-NMF privilege a high number of element ($K$ = 200) with a fine description of the audio samples ($w_t \in \lbrace 0.5$,$1 \rbrace$ second).
-For TI-NMF, the composition of the initial dictionary is more diverse both on the $K$ value and on the finesse  of the description ($w_t \in \lbrace all$,$0.5$, $1 \rbrace$ second).
+SUP and SEM-NMF privilege a high number of element ($\mathbf{K}$ = 200) with a fine description of the audio samples ($\mathbf{w_t} \in \lbrace 0.5$,$1 \rbrace$ second).
+For TI-NMF, the composition of the initial dictionary is more diverse both on the $\mathbf{K}$ value and on the finesse  of the description ($\mathbf{w_t} \in \lbrace all$,$0.5 \rbrace$ second).
 
@@
 \includegraphics[width=\linewidth]{figures/mea_grafic_bar.pdf}
-\caption{$MAE$ errors according each sound environment for the best combination of the LP filter ($f_c$ = 500 Hz), SUP-NMF ($\beta$ = 2, $K$ = 25, $w_t$ = 1 second), SEM-NMF ($\beta$ = 1, $K$ = 200, $w_t$ = 1 second) and TI-NMF ($\beta$ = 2, $K$ = 200, $w_t$ = 0.5 second, $t_{h} = 0.33$)}
+\caption{$MAE$ errors according each sound environment for the best combination of the LP filter ($f_c$ = 500 Hz), SUP-NMF ($\beta$ = 2, $\mathbf{K}$ = 25, $\mathbf{w_t}$ = 1 second), SEM-NMF ($\beta$ = 1, $\mathbf{K}$ = 200, $\mathbf{w_t}$ = 1 second) and TI-NMF ($\beta$ = 2, $\mathbf{K}$ = 200, $\mathbf{w_t}$ = 0.5 second, $\mathbf{t_{h}} = 0.32$)}
 \label{fig:mae_env}
@@
 Besides having the lowest $mMAE$ error, TI-NMF presents the most performing results.
-The park environment is the only case where an other NMF approach out-performed significantly TI-NMF ($MAE$ = 3.05 dB).
+The park environment is the only case where an other NMF approach out-performed significantly TI-NMF ($MAE$ = 2.95 dB).
 In this sound environment, the traffic dictionary is then composed, in average, of more than half the 200 elements in $\mathbf{W'}$ (136 elements).
 For the rest of the sound environments, TI-NMF has the lowest error.
-For very noisy environment the error is even very low ($MAE=$ 0.31 dB).
+For very noisy environment the error is even very low ($MAE=$ 0.28 dB).
 In this case, in average, $\mathbf{W'}$ is composed of 198 traffic elements.
@@
 By considering the elements the most similar to traffic spectra with a hard thresholding, it makes it possible to adapt specifically the dictionary to the sound scenes.
-Consequently, in the case where the emplacement or the sound environment of the sensors cannot be identified (for instance within a mobile measurement framework), TI-NMF, with $\beta = 2$, $K$ = 200, $w_t = 1$ second, $t_{h} = 0.33$, seems the best generic framework, for the sound corpus studied.
+Consequently, in the case where the emplacement or the sound environment of the sensors cannot be identified (for instance within a mobile measurement framework), TI-NMF, with $\beta = 2$, $\mathbf{K}$ = 200, $\mathbf{w_t} = 0.5$ second, $\mathbf{t_{h}} = 0.32$, seems the best generic framework, for the sound corpus studied.
 But if the sound environment can be identified through a prior analysis, or based on positioning data \cite{can2015noise,lavandier2016urban}, it seems possible to adapt the proposed work by selecting the most efficient approach (for instance SEM-NMF within parks), in order to further reduce the error in the estimated road traffic sound levels.

### 2018-03-05 paper/gloaguenAppliedNMF.tex
add cumulative times

@@
 Two of the 76 recordings are rejected for the analysis because the audio files were corrupted, resulting in 74 valid audio files assumed as representative of the variety of sound environments.
-The recordings are listened and categorized within four different sound environments, as proposed in \cite{can_describing_2015}: park (P, 8 audio files), quiet street (Q, 33 audio files), noisy street (N, 24 audio files) and very noisy street (vN, 9 audio files).
+The recordings are listened and categorized within four different sound environments, as proposed in \cite{can_describing_2015}: park (P, 8 audio files with a cumulative duration of 16min01), quiet street (Q, 33 audio files with a cumulative duration of 77min27), noisy street (N, 24 audio files with a cumulative duration of 56min10) and very noisy street (vN, 9 audio files with a cumulative duration of 21min42).
 Then, each audio file is annotated, noting the start and end time of each sound event along with its sound class.

### 2018-03-06 paper/gloaguenAppliedNMF.tex
add mTIR paragraph

@@
 
+\begin{table}[t]
+\centering
+\caption{$mTIR$ for each sound environment}
+\begin{tabular}{lc}
+ & $mTIR$ (dB)\\ \hline
+ park & -9.10 ($\pm$ 7.35) \\
+ quiet street & 0.88 ($\pm$ 5.92) \\
+ noisy street & 6.96 ($\pm$ 5.16) \\
+ very noisy street & 15.75 ($\pm$ 9.78) \\ \hline
+\end{tabular}
+\label{tab:mTIR}
+\end{table}
+
 With this built-up database, the \textit{SimScene} software and the annotations of the recordings, 74 simulated sound scenes are generated, which have the same temporal structure than the recordings.
 The \textit{ebr} parameter is adjusted manually on each sound scene to be faithful compared to the recorded scenes.
+To validate this adjustment, the mean \textit{Traffic Interfering Ratio} is calculated and summarized on Table \ref{tab:mTIR}.
+It expresses, on all the scenes of an sound environment, the mean difference between the equivalent traffic sound levels of each scene, $L_{p,traffic}$, with the sound level of the \textit{interfering} sound class, $L_{p, interfering}$, which gathered all the other sound sources not related to the traffic, see Eq \ref{eq:mTIR}.
+It quantifies the predominance of the traffic component on the 4 sound environments.
+
+\begin{equation}\label{eq:mTIR}
+mTIR = \frac{\sum_{i = 1}^M L_{p,traffic} - L_{p, interfering}}{M}.
+\end{equation}
+
+As it is the sound environment where the traffic is less present, the $mTIR$ is always negative for the \textit{park}.
+The \textit{interfering} sound class is therefore the main sound source.
+In the case of the \textit{quiet street}, the $TIR$ can be positive or negative depending on the traffic presence on the scene.
+For the 2 others sound environments, when the traffic become the main sound source, $mTIR$ is always positive.
+
 To validate its realism, the corpus of transcribed urban sound scenes is submitted to a perceptual test.

### 2018-03-06 paper/gloaguenAppliedNMF.tex
removed unnecessary figures + add reference in biblio

@@
 Dedicated to sound separation, Immani and Kasa\"i \cite{satoshi_innami_nmf-based_2012} used NMF in a two steps sound separation with the help of time variant gain features.
-A first study [article JASM] has been conducted, in which diverse NMF formula, namely the supervised, the semi-supervised, and the threshold initialized NMF, have been applied on a large set of simulated sound scenes that which mixed traffic component with specific urban sounds at calibrated sound levels.
+A first study \cite{gloaguen2018Estimation} has been conducted, in which diverse NMF formula, namely the supervised, the semi-supervised, and the threshold initialized NMF, have been applied on a large set of simulated sound scenes that which mixed traffic component with specific urban sounds at calibrated sound levels.
 The study proved the interest of NMF for urban sound environments and compared the benefits of each approach.
@@
 \end{subequations}
-
-The allure of $D_{\theta}\left(\mathbf{W_0} \Vert \mathbf{W'} \right)$ and $\alpha$ are displayed in Figure \ref{fig:example_threshold}.\\
-
-\begin{figure}[t]
-    \centering
-       \includegraphics[width=0.9\linewidth]{./figures/thresh.pdf}
-    \caption{Evolution of the $\alpha$ weight according to $D_{\theta}\left(\mathbf{W_0} \Vert \mathbf{W'} \right)$ for hard thresholding with $t_h$ = 0.6 (in dashed line)}
-    \label{fig:example_threshold}
-\end{figure}
 

### 2018-03-06 paper/gloaguenAppliedNMF.tex
layout of legends

@@
 \includegraphics[width=.9\linewidth]{./figures/schema_introduction_nmf.pdf} 
-\caption{NMF for an audio sample with 3 elements ($K$ = 3): passing car (a), car horn (b) and whistling bird (c)}
+\caption{NMF for an audio sample with 3 elements ($K$ = 3): passing car (a), car horn (b) and whistling bird (c).}
 \label{fig:exampleNMF}
@@
 \includegraphics[width=\linewidth]{./figures/trajet_19pts.png}
-\caption{Walked path with the 19 stop points  \cite{aumond_modelling_2017}}.
+\caption{Walked path with the 19 stop points  \cite{aumond_modelling_2017}.}
 \label{fig:map_grafic}
@@
        \includegraphics[width=.8\linewidth]{./figures/exampleSimScene.pdf}
-    \caption{Spectrogram of a simple scene created with the \textit{SimScene} software with a sound background (road traffic in green) and 3 sound events (car horn in purple, passing car in red and whistling bird in blue)}
+    \caption{Spectrogram of a simple scene created with the \textit{SimScene} software with a sound background (road traffic in green) and 3 sound events (car horn in purple, passing car in red and whistling bird in blue).}
     \label{fig:example_simScene}
@@
 \centering
-\caption{$mTIR$ for each sound environment}
+\caption{$mTIR$ for each sound environment.}
 \begin{tabular}{lc}
@@
 \includegraphics[width=\linewidth]{figures/testPerceptif_boxplotType_EN.pdf} 
-\caption{Box-and whiskers plot of the rating of realism according to the type (recorded /transcribed)}
+\caption{Box-and whiskers plot of the rating of realism according to the type (recorded /transcribed).}
 \label{fig:boxPlot_test}
@@
 \includegraphics[width=.9\linewidth]{figures/bloc_diagram_estimator.pdf}
-\caption{Bloc diagram of the experimental protocol}
+\caption{Bloc diagram of the experimental protocol.}
 \label{fig:bloc_diagram_estimator}
@@
 \includegraphics[width=.9\linewidth]{figures/bloc_diagram_NMF_EN_2.pdf}
-\caption{Bloc diagram of the experimental protocol}
+\caption{Bloc diagram of the NMF estimator.}
 \label{fig:bloc_diagram_nmf}
@@
 \begin{table*}[t]
-\caption{Summary of the different experimental factors and their modalities taken into account in the frequency low-pass filter estimator}
+\caption{Summary of the different experimental factors and their modalities taken into account in the frequency low-pass filter estimator.}
 \centering
@@
 \centering
-\caption{Summary of the different experimental factors and their modalities taken into account in NMF estimator}
+\caption{Summary of the different experimental factors and their modalities taken into account in NMF estimator.}
 \begin{tabularx}{17.5cm}{L{3cm}@{}C{12cm}@{}C{2cm}@{}}
@@
 \includegraphics[width=\linewidth]{figures/mea_grafic_bar.pdf}
-\caption{$MAE$ errors according each sound environment for the best combination of the LP filter ($f_c$ = 500 Hz), SUP-NMF ($\beta$ = 2, $\mathbf{K}$ = 25, $\mathbf{w_t}$ = 1 second), SEM-NMF ($\beta$ = 1, $\mathbf{K}$ = 200, $\mathbf{w_t}$ = 1 second) and TI-NMF ($\beta$ = 2, $\mathbf{K}$ = 200, $\mathbf{w_t}$ = 0.5 second, $\mathbf{t_{h}} = 0.32$)}
+\caption{$MAE$ errors according each sound environment for the best combination of the LP filter ($f_c$ = 500 Hz), SUP-NMF ($\beta$ = 2, $\mathbf{K}$ = 25, $\mathbf{w_t}$ = 1 second), SEM-NMF ($\beta$ = 1, $\mathbf{K}$ = 200, $\mathbf{w_t}$ = 1 second) and TI-NMF ($\beta$ = 2, $\mathbf{K}$ = 200, $\mathbf{w_t}$ = 0.5 second, $\mathbf{t_{h}} = 0.32$).}
 \label{fig:mae_env}

### 2018-03-06 paper/gloaguenAppliedNMF.tex
correct minor errors

@@
 The extraction of traffic elements in $\mathbf{W'}$ is carried out by a hard thresholding method \cite{donoho1994threshold}.
-It consists in weigthing in a binary way the traffic elements $\mathbf{W'}$ according to $D_{\theta}\left(\mathbf{w_0} \Vert \mathbf{w'} \right)$ and a threshold value such as :
+It consists in weighting in a binary way the traffic elements $\mathbf{W'}$ according to $D_{\theta}\left(\mathbf{w_0} \Vert \mathbf{w'} \right)$ and a threshold value such as:
 
@@
 These methods are applied on simulated sound scenes in order to compare the estimated sound levels with the exact solutions.
-In \cite{}, the sound corpus was composed of a mix of traffic samples with specific sound classes (\textit{alert, animals, climate, human, mechanics, transportation}.
+In \cite{gloaguen2018Estimation}, the sound corpus was composed of a mix of traffic samples with specific sound classes (\textit{alert, animals, climate, human, mechanics, transportation}).
 Here, in order to implement this method in embedded sensors, a new more realistic sound corpus is generated based on urban recordings mixing all kind of sound sources.
@@
 To transcribe the recordings in simulated scenes, a high quality sound database (wav format, 44.1 kHz sampling rate, high \textit{Signal Noise Ratio}) has been built-up from audio samples found online (\textit{freesound.org}) or with the help of an already existing sound database \cite{salamon2014dataset}.
-The sound database is composed of two categories of sound : the \textit{event} category, which includes 245 brief sound samples considered as salient, with a 1 to 20 seconds duration and classified among 21 sound classes (\textit{ringing bell, whistling bird, car horn, passing car, hammer, barking dog, siren, footstep, metallic noise, voice \dots}) and the \textit{background} (or \textit{texture}) category gathering 154 long duration sounds ($\approx$ 1m30), whose acoustic properties do not vary in time.
+The sound database is composed of two categories of sound: the \textit{event} category, which includes 245 brief sound samples considered as salient, with a 1 to 20 seconds duration and classified among 21 sound classes (\textit{ringing bell, whistling bird, car horn, passing car, hammer, barking dog, siren, footstep, metallic noise, voice \dots}) and the \textit{background} (or \textit{texture}) category gathering 154 long duration sounds ($\approx$ 1m30), whose acoustic properties do not vary in time.
 This category includes among others\textit{whistling bird, crowd noise, rain, children playing in schoolyard, constant traffic noise} \dots sound classes.
@@
 
-A non-negative matrix factorization frame has been applied as a source separation tool to estimate the traffic sound level  from a corpus of urban sound scenes artifically built.
+A non-negative matrix factorization frame has been applied as a source separation tool to estimate the traffic sound level  from a corpus of urban sound scenes artificially built.
 In addition, the realism of the scenes has been verified thanks to a perceptual test.
-The proposed protocol generate audio sample similar than the outputs of a deployed sensor networks with tha advantages of the simulation process (sound level and position of each source controlled and known).
+The proposed protocol generate audio sample similar than the outputs of a deployed sensor networks with the advantages of the simulation process (sound level and position of each source controlled and known).
 

### 2018-03-06 paper/gloaguenAppliedNMF.tex
proofread intro

@@
 
+\newcommand{\ml}[1]{\textcolor{red}{ML : #1}}
 \modulolinenumbers[5]
@@
 
-Acoustic sensor networks are increasingly deployed in cities, and appear more and more as a possible tool to enrich modeled road traffic noise maps through data assimilation techniques.
-This, or more simply the validation of the modeled maps through measures, requires first being able to isolate from the measured sound mixtures the road traffic sound level.
+Experimental acoustic sensor networks are currently tested in large cities, and appear more and more as a useful tool to enrich modeled road traffic noise maps through data assimilation techniques.
+One challenge is to be able to isolate from the measured sound mixtures acoustic quantities of interest such as the  sound level of road traffic.
 This task is anything but trivial because of the multiple sound sources that overlap within urban sound mixtures.
+
 In this paper, a Non-negative Matrix Factorization (NMF) framework is developed to estimate road traffic noise levels within urban sound scenes.
-A corpus of sound scenes is artificially built imitating real annotated recordings.
-The realism of the scenes is validated through a perceptual test, forming a protocol that both reproduces the sensor network outputs, and in which the actual occurrence and sound level of each source is known.
-Three variants of NMF are tested, namely supervised, semi-supervised, and threshold initialized NMF.
-While the semi-supervised approach is the most appropriated for park environments, threshold initialized NMF, which is developed especially for this research purpose, appears to be the best generic approach, allowing road traffic noise level estimation with average errors of 1.3 dB over the tested corpus of sound scenes.
+To evaluate the performance of the proposed approach, a synthetic corpus of sound scenes is designed to cover most common soundscape settings.
+The realism of the produced scenes is validated through a perceptual test, forming a protocol that both reproduces the sensor network outputs, and in which the actual occurrence and sound level of each source is known.
+
+Several variants of NMF are tested.
+The threshold initialized NMF appears to be the most relibale approach, allowing road traffic noise level estimation with average errors of 1.3 dB over the tested corpus of sound scenes.
 
@@
 \begin{keyword}
-non-negative matrix factorization \sep urban sound environment \sep road traffic sound level
+non-negative matrix factorization \sep urban sound environment \sep road traffic sound level estimation
 \end{keyword}
@@
 
-Noise mapping has been recommended as a tool to tackle noise pollution, in response to the growing demand from urban dwellers for a better environment.
+In response to the growing demand from urban dwellers for a better environment, noise mapping has been recommended as a tool to tackle noise pollution.
 The enactment of the European Directive 2002/EC/49 makes such maps mandatory to cities over 100 000 inhabitants.
@@
 Road traffic concentrates particular attention as it is the main urban source of noise annoyance.
-Road traffic noise maps are built from data collection that consist of traffic data collected on the main roads (flow rates, mean speeds and heavy vehicle ratio) and urban geographic data (building heights and location, topology, ground surfaces, etc.).
-Follows sound emission and sound propagation computations, resulting in the production of the two indicators equivalent A-weighted sound levels, $L_{DEN}$ (\textit{Day-Evening-Night}) and $L_N$ (\textit{Night}) \cite{kephalopoulos2012common}.
+Road traffic noise maps are typically built from data collection that consist of traffic data collected on the main roads (flow rates, mean speeds and heavy vehicle ratio) and urban geographic data (building heights and location, topology, ground surfaces, etc.).
+Follows sound emission and sound propagation computational techniques, resulting in the production of the two indicators equivalent A-weighted sound levels, $L_{DEN}$ (\textit{Day-Evening-Night}) and $L_N$ (\textit{Night}) \cite{kephalopoulos2012common}.
 This procedure also enables drawing up action plans to reduce the noise exposure.
 Despite their unanimously recognized interest, noise maps suffer some limitations.
-The computer efficiency required to produce noise maps at the city scale calls simplifications of the numerical tools and the simulation models that both generate uncertainties \cite{van_leeuwen_noise_2015}.
-Data collection is itself a vector of uncertainty.
-Moreover, the produced aggregated indicators mask the sound levels evolution due to the traffic variations throughout the day.
+The computing cost required to produce noise maps at the city scale calls simplifications of the numerical tools and the simulation models that both generate uncertainties \cite{van_leeuwen_noise_2015}.
+Data collection is itself also a vector of uncertainty.
+Moreover, the produced aggregated indicators do not model the sound levels evolution due to the traffic variations throughout the day.
 
-Therefore, noise measurements are increasingly used in addition to simulation to describe urban noise environments.
+Noise measurements are thus increasingly used in addition to simulation to describe urban noise environments.
 Several measurement set-ups have been proposed in the last years, including  mobile measurements with high quality microphones \cite{manvell2004sadmam, can2014measurement}, participative sensing through dedicated smartphone applications \cite{picaut:halshs-01565214, ventura2017evaluation}, or the development of fixed-sensor networks.
@@
 
-All these measurement protocol enables a priori combining measures and modeling to improve the accuracy of the produced noise maps.
+All these measurement protocols allows the combination of measures and predictions to improve the accuracy of the produced noise maps.
 Traffic noise maps and measurements were compared on restrictive areas in \cite{lefebvre2017traffic} and \cite{mioduszewski2011noise}.
@@
 Many works have dealt with the detection \cite{luitel2016sound} or the recognition \cite{defreville_automatic_2006} of sound events in environmental sound scenes.
-In these cases, a usually two-step schemes is followed were audio samples are described with a set of features (Mel Frequency Cepstral Coefficient, MPEG-7 descriptors \dots) and classified them with the help of a classifiers (Gaussian Mixtures Models, Artificial Neural Network \dots) \cite{chu2008environmental, cowling_comparison_2003}.
+In these cases, a two-step scheme is followed where audio samples are described with a set of features (Mel Frequency Cepstral Coefficient, MPEG-7 descriptors \dots) and classified them with the help of a classifiers (Gaussian Mixtures Models, Artificial Neural Network \dots) \cite{chu2008environmental, cowling_comparison_2003}.
 The classifiers is learnt from a learning database and are next applied on a test database to validate the algorithms.
 Dedicated to the traffic, in \cite{socoro_anomalous_2017}, an Anomalous Event Detection, based on MFCC features, is proposed with the specific aim to improve the traffic sound estimation.
-It is based on the detection of sound events in order to discard them.
+It is based on the detection of unwanted sound events in order to discard them.
 
-An other approach, followed in this paper, is to consider the Blind Source Separation paradigm, see Figure \ref{fig:source_separation}, which consists in the extraction of a specific signal inside a set of mixed signals.
-From the different existing methods (CASA, ICA), Non-negative Matrix Factorization (NMF) \cite{lee_learning_1999}, seems the most relevant method for monophonic sensor networks.
-Dealing easily with the overlapping between the sound sources, this method approximes the magnitude spectrogram of a single signal by the product of two non negative matrices.
-A lot of applications can be found for musical \cite{smaragdis_non-negative_2003, benetos2006musical} and speech \cite{wilson_speech_2008, mysore2011non} contents.
+An other approach, followed in this paper, is to consider the blind source separation paradigm which consists in the extraction of a specific signal inside a set of mixed signals, see Figure \ref{fig:source_separation}.
+From the different existing methods, Non-negative Matrix Factorization (NMF) \cite{lee_learning_1999}, appers to be a relevant method for monophonic sensor networks.
+Many applications can be found for musical \cite{smaragdis_non-negative_2003, benetos2006musical} and speech \cite{wilson_speech_2008, mysore2011non} contents.
 Dedicated to sound separation, Immani and Kasa\"i \cite{satoshi_innami_nmf-based_2012} used NMF in a two steps sound separation with the help of time variant gain features.
-A first study \cite{gloaguen2018Estimation} has been conducted, in which diverse NMF formula, namely the supervised, the semi-supervised, and the threshold initialized NMF, have been applied on a large set of simulated sound scenes that which mixed traffic component with specific urban sounds at calibrated sound levels.
-The study proved the interest of NMF for urban sound environments and compared the benefits of each approach.
+A first study \cite{gloaguen2018Estimation} has been conducted, in which diverse NMF estimation rules are compared, namely the supervised, the semi-supervised, and the threshold initialized NMF, have been applied on a large set of simulated sound scenes that which mixed traffic component with specific urban sounds at calibrated sound levels.
+The study demonstrated the interest of NMF for urban sound environments and compared the benefits of each approach on a validation corpus, designed to .
 However, it has now to face to real urban sound scenes.
 
+\ml{la fin de ce par est a revoir car c'est le point crucial du papier. Bien expliquer le corpus du papier d'avant son usage et en quoi on passe ici a l'étape superieure.}
+
 In this paper, the NMF framework is applied on a corpus of simulated sound scenes, generated based on annotated urban recordings, and whose realism is validated by a perceptual test.
-The NMF framework and its different implemented versions are described in section \ref{part:nmf}.
+The NMF framework and its different optimizatio strategies are described in section \ref{part:nmf}.
 Next, the corpus of urban sound scenes is presented in section \ref{part:urban_scene}, from the sound database built-up to its validation through a perceptual test.
-Finally, in section \ref{part:expProtocol} and \ref{part:results}, the experimental protocol and the results are exposed.
+The experimental protocol and the results are then presented and discussed in section \ref{part:expProtocol} and \ref{part:results}.
 

### 2018-03-07 paper/gloaguenAppliedNMF.tex
proofread until db

@@
 
-\ml{la fin de ce par est a revoir car c'est le point crucial du papier. Bien expliquer le corpus du papier d'avant son usage et en quoi on passe ici a l'étape superieure.}
+\ml{la fin de ce par est a revoir car c'est le point crucial du papier.
+Bien expliquer le corpus du papier d'avant, son usage, les questions qui restent en suspens et en quoi on passe ici a l'étape superieure en y repondant.}
 
-In this paper, the NMF framework is applied on a corpus of simulated sound scenes, generated based on annotated urban recordings, and whose realism is validated by a perceptual test.
-The NMF framework and its different optimizatio strategies are described in section \ref{part:nmf}.
-Next, the corpus of urban sound scenes is presented in section \ref{part:urban_scene}, from the sound database built-up to its validation through a perceptual test.
+In this paper, the NMF framework is applied on a corpus of simulated sound scenes, generated based on annotated urban recordings, and whose degree of realism is validated through a perceptual test.
+The NMF framework and its different optimization strategies are described in section \ref{part:nmf}.
+Next, the corpus of urban sound scenes is presented in section \ref{part:urban_scene}, from the sound database built-up to its validation.
 The experimental protocol and the results are then presented and discussed in section \ref{part:expProtocol} and \ref{part:results}.
@@
 This condition however is not mandatory.
-When an audio file is considering, $\mathbf{V}$ is usually considered as the magnitude spectrogram obtained by a Short-Time Frequency Transform, $\mathbf{W}$ includes audio spectra and $\mathbf{H}$ is equivalent to the temporal evolution of each spectrum, see Figure \ref{fig:exampleNMF}.
+When  applying NMF to audio data, $\mathbf{V}$ is usually considered as the magnitude spectrogram obtained by a Short-Time Frequency Transform, $\mathbf{W}$ includes audio spectra and $\mathbf{H}$ is equivalent to the temporal activation of each spectrum, see Figure \ref{fig:exampleNMF}.
 Because of the non-negativity constraint, only additive combinations between the elements of $\mathbf{W}$ are considered.
-The dictionary is then composed of elementary elements providing a part based representation.
+%The dictionary is then composed of elementary elements providing a part based representation.
 
@@
 \includegraphics[width=.9\linewidth]{./figures/schema_introduction_nmf.pdf}
-\caption{NMF for an audio sample with 3 elements ($K$ = 3): passing car (a), car horn (b) and whistling bird (c).}
+\caption{NMF decomposition of an audio spectrogram composed of 3 elements ($K$ = 3): passing car (a), car horn (b) and whistling bird (c).}
 \label{fig:exampleNMF}
@@
 
-$d_{\beta}(x\vert y)$ is mainly chosen as a $\beta$-divergence \cite{fevotte_algorithms_2011}, a sub-classes belonging to the Bregman divergences \cite{hennequin_beta-divergence_2011} which include 3 specific divergence and distance calculations: the Euclidean distance (eq.
+$d_{\beta}(x\vert y)$ is usually chosen as a $\beta$-divergence \cite{fevotte_algorithms_2011}, a sub-classes belonging to the Bregman divergences \cite{hennequin_beta-divergence_2011} which include 3 specific divergence calculations: the Euclidean distance (eq.
 \ref{eq:def_distEUC}), the Kullback-Leibler divergence (eq.
@@
 The minimization problem (\ref{eq:min-D-WH}) is solved iteratively by updating the form of matrices $\mathbf{W}$ and $\mathbf{H}$.
-Different algorithms such as Alternating Least Square Method \cite{cichocki_regularized_2007} or Projected Gradient \cite{lin_projected_2007} have been proposed.
-The most commonly algorithm used, and the chosen method here, is Multiplicative Update \cite{lee_algorithms_2000} as it ensures non-negative results and the convergence of the results \cite{fevotte_algorithms_2011}.
+Different algorithms such as Alternating Least Square Method \cite{cichocki_regularized_2007} or Projected Gradient \cite{lin_projected_2007} have been considered.
+The most commonly used algorithm is the Multiplicative Update \cite{lee_algorithms_2000}.
+The latter method is chosen here, as it ensures non-negative results and the convergence of the results \cite{fevotte_algorithms_2011}.
 
@@
 It leads to \textit{supervised} NMF (SUP-NMF).
-$\mathbf{H}$ is then the only matrix to determine and is updated at every iteration (eq.
+$\mathbf{H}$ is then the only matrix to estimate and is updated at every iteration (eq.
 \ref{eq:updateH}) \cite{fevotte_algorithms_2011}.
@@
 
-Here, in an urban context, the sound sources are known and their audio samples can be obtained to learn $\mathbf{W}$, see section \ref{part:dictionary_building}.
+Here, in an urban context, if the sound sources are known, their audio samples can be obtained to learn $\mathbf{W}$, see section \ref{part:dictionary_building}.
 As the position of each element is indexed, the traffic source separation from the other sound sources is made by extracting, from the dictionary and the activation matrix, the related elements:
@@
 \subsection{Semi-supervised NMF}
-The main issue with the supervised approach is the limit imposed by $\mathbf{W}$: to be successful, all the acoustical sources must be considered in it which is not possible in an urban environment.
-As the main source that interests us can be known \textit{a priori}, semi-supervised NMF (SEM-NMF) \cite{lee_semi-supervised_2010} is considered to better take into account the interfering sound sources.
-This approach proposes to decompose $\mathbf{W}_{F \times (K+J)}$ with two distinctive matrices: $\mathbf{W} = \left[ \mathbf{W_s}~\mathbf{W_r} \right]$ where $\mathbf{W_s}_{F \times K}$ is a fixed part of $\mathbf{W}$ composed of audio spectra and $ \mathbf{W_r}_{F \times J}$, a mobile part which is updated, see eq.
+
+The main issue with the supervised approach is the representational limit imposed by a fixed $\mathbf{W}$.
+To be completely successful, all the acoustical sources must be considered in it which is not possible in an urban environment.
+However, in our application setting, the main source of interest, \textit{i.e.} the traffic, can be known \textit{a priori} \ml{manque une phrase} semi-supervised NMF (SEM-NMF) \cite{lee_semi-supervised_2010} is considered to better take into account the interfering sound sources.
+To do so, $\mathbf{W}_{F \times (K+J)}$ is decomposed into two distinctive matrices: $\mathbf{W} = \left[ \mathbf{W_s}~\mathbf{W_r} \right]$ where $\mathbf{W_s}_{F \times K}$ is a fixed part of $\mathbf{W}$ composed of audio spectra and $ \mathbf{W_r}_{F \times J}$, a mobile part which is updated, see eq.
 \ref{eq:W_r_SS}.
@@
 
-This approach as the advantage, with the add of the mobile part $\mathbf{W_r}$, to bring more flexibility and then to be more adaptive to the different urban sound environments.
+The addition of the mobile part $\mathbf{W_r}$ gives more flexibility.
+The representational capability is increased, thus the approach is more adaptive to the different urban sound environments.
 Applications of SEM-NMF can be found for musical \cite{weninger2012supervised, kitamura_music_2014} and speech content \cite{joder2012real, mysore2011non}.
@@
 
-We propose a third approach based on the unsupervised NMF framework: Threshold Initialized NMF (TI-NMF).
+To allow even more flexibility while still considering prior knowledge of the source of interest, we propose a third approach based on the unsupervised NMF framework: Threshold Initialized NMF (TI-NMF).
 Usually, in unsupervised NMF, the dictionary is initiated randomly when there is no \textit{prior} knowledge on the sound sources present.
@@
 where $\mathbf{w}$ is a $k$ element of $\mathbf{W}$ of $F \times 1$ dimension.
-When $D_{\theta}\left(\mathbf{w_0} \Vert \mathbf{w'} \right)$=1, the element $k$ from $\mathbf{W'}$ is exactly similar than in $\mathbf{W_0}$.
+When $D_{\theta}\left(\mathbf{w_0} \Vert \mathbf{w'} \right)$=1, the element $k$ from $\mathbf{W'}$ is equal to the in $\mathbf{W_0}$ \ml{pas clair, a quoi sert le k ?}.
 If $D_{\theta}\left(\mathbf{w_0} \Vert \mathbf{w'} \right)$=0, the element is fully different.
-Next, the similarities are sorted in descending order.
+Next, the similarities are sorted in descending order \ml{le sort n'est pas necessaire, le seuil suffit}.
 The extraction of traffic elements in $\mathbf{W'}$ is carried out by a hard thresholding method \cite{donoho1994threshold}.
 It consists in weighting in a binary way the traffic elements $\mathbf{W'}$ according to $D_{\theta}\left(\mathbf{w_0} \Vert \mathbf{w'} \right)$ and a threshold value such as:
+
+\ml{on pourrait tout a fait imaginer l'ajout d'élément de dictionaire random pour se rapprocher de la semi.}
 
@@
 
+\ml{faire un par resumant les prop des trois methodes avec un accent sur l'utilisation de la connaissance a priori.}
+
 These methods are applied on simulated sound scenes in order to compare the estimated sound levels with the exact solutions.
 In \cite{gloaguen2018Estimation}, the sound corpus was composed of a mix of traffic samples with specific sound classes (\textit{alert, animals, climate, human, mechanics, transportation}).
-Here, in order to implement this method in embedded sensors, a new more realistic sound corpus is generated based on urban recordings mixing all kind of sound sources.
+Here, in order to implement this method in embedded sensors, a new and more realistic sound corpus is generated based on urban recordings mixing many kind of sound sources.
 

### 2018-03-08 paper/gloaguenAppliedNMF.tex
proof read until protocol

@@
 
-These methods are applied on simulated sound scenes in order to compare the estimated sound levels with the exact solutions.
+These methods are applied on simulated sound scenes in order to compare the estimated sound levels with a non ambiguous reference.
 In \cite{gloaguen2018Estimation}, the sound corpus was composed of a mix of traffic samples with specific sound classes (\textit{alert, animals, climate, human, mechanics, transportation}).
-Here, in order to implement this method in embedded sensors, a new and more realistic sound corpus is generated based on urban recordings mixing many kind of sound sources.
+Here, in order to implement this method in embedded sensors, a new and more realistic sound corpus by considering more realistic urban sound environments with many kind of sound sources.
 
@@
 
-The urban sound scenes are taken from 76 recordings from 2 to 5 min, achieved in the \nth{13} district of Paris (France) at 19 different locations \footnote{Recordings were made as part of the Grafic project funded by Ademe}, which cover four various sound environments (Figure \ref{fig:map_grafic}), see Figure \ref{fig:map_grafic}.
+\ml{Il faut faire une figure et un paragraphe resumant la figure.
+Voici un essai (je pense d'ailleurs que cela devrait faire partie des claims dans l'intro) : In order to validate the above described methods, the evaluation corpus is both required to be realistic and non ambiguous in terms of traffic sound level.
+The former calls for the use recordings of urban sound environments, and the latter imposes the use of controled use of sound sources with known properties: onset and offset time location, and volume.
+Indeed, the precise annotation of the sound level of a given source of interest in a polyphonic mixture is not feasible.
+We thus propose to design such evalution corpus based on two corpora.
+The first, called the reference corpus, is a set of recordings of urban sound environments, from which a sort of "score" is extracted manually, \textit{i.e.} which sound source is present from this time to this time, at which sound level.
+The second, called the elemental corpus, is a set of monophonic recordings of sound sources that are occurs in the scenes of the reference.
+The recordings of the elemental corpus are not taken from the reference corpus.
+The evaluation corpus is then built by sequencing recordings of the elemental corpus according to the scores of the reference corpus.
+(Le reste est a reecrire avec les termes proposés)}
+
+More precisely, the urban sound scenes are taken from 76 recordings from 2 to 5 min, achieved in the \nth{13} district of Paris (France) at 19 different locations \footnote{Recordings were made as part of the Grafic project funded by Ademe}, which cover four various sound environments (Figure \ref{fig:map_grafic}), see Figure \ref{fig:map_grafic}.
 A complete description of the experimental protocol can be found in \cite{aumond_modelling_2017}.
@@
 The recordings are listened and categorized within four different sound environments, as proposed in \cite{can_describing_2015}: park (P, 8 audio files with a cumulative duration of 16min01), quiet street (Q, 33 audio files with a cumulative duration of 77min27), noisy street (N, 24 audio files with a cumulative duration of 56min10) and very noisy street (vN, 9 audio files with a cumulative duration of 21min42).
-Then, each audio file is annotated, noting the start and end time of each sound event along with its sound class.
-The aim of the annotation phase is next to transcribe the recordings, in order to obtain simulated sound scenes with the same distribution of sound events as the recordings and therefore as close as possible to the realistic scenes.
+Then, each audio file is annotated, annotating the start, end time \ml{and level ?} of each sound event along with its sound class.
+\ml{cette phrase ne dis probablement pas ce qsue tu veux dire annotation et Transcription sont des termes equivalents, The aim of the annotation phase is next to transcribe the recordings}, in order to obtain simulated sound scenes with the same \ml{ca peut preter a confusion en implicitant une distribution statistique : distribution} of sound events as the recordings and therefore as close as possible to the realistic scenes.
 
@@
 
-\subsection{Transcription of the recordings}\label{part:simScene}
+\subsection{Generation of the evaluation corpus}\label{part:simScene}
 The sound scenes are generated with the \textit{SimScene} software\footnote{Open-source project available at: \url{https://bitbucket.org/mlagrange/simscene}}, \cite{rossignol_simscene:_2015}, which is a simulation software generating monaural sound mixtures in wav format with a 44.1 kHz sampling rate from an isolated sound database.
 This software have already been used in a wide range of experiments for sound detection algorithm assessment \cite{lafay_new_2014, benetos2016detection}.
-The control of high level parameters can be handle by the user as the sound class presence, the time between each sample of one sound class, the ratio between the sound level of an event class with the background (i.e the \textit{event background ratio} shorten \textit{ebr})\dots Each parameter is completed with a standard deviation to bring random behavior.
+\ml{Pas clair The control of high level parameters can be handled by the user as the sound class occurence, the time between each sample of one sound class, the ratio between the sound level of an event class with the background (i.e the \textit{event background ratio} shorten \textit{ebr})\dots Each parameter is completed with a standard deviation to bring random behavior.
 It allows too the design of a sound mixture from an annotation text file.
+A remplacer par ceci: The \textit{SimScene} software allows for several type of sequencing, from abstract ones (time indexes and amplitudes are drawn from random distributions) to  precise ones, where the time indexes and amplitudes for each event is set by the user.
+The latter type is considered in this study.}
 As output, \textit{SimScene} generates an audio of the global sound mixture and an audio for each sound class present in the scene, which makes it possible to know their exact contributions in the scene, see Figure \ref{fig:example_simScene}.
@@
 This category includes among others\textit{whistling bird, crowd noise, rain, children playing in schoolyard, constant traffic noise} \dots sound classes.
-Each sound class is present in multiple samples (\textit{carHorn01.wav, carHorn02.wav \dots}) to bring diversity.
+Each sound class is composed of multiple samples (\textit{carHorn01.wav, carHorn02.wav \dots}) that are randomly chosen by the \textit{simScene} software to bring diversity.
 As the road traffic is the main component in urban environment and is the sound source of interest, recordings of car passages has been made on the Ifsttar's runway.
-The recordings has been made for 4 cars (Renault Senic, Renault Megane and Renault Clio, Dacia Sandero),  for different speeds and gear ratios.
-In all, 103 car passages have been recorded.
-The audio samples of the first two cars (Renault Senic and Renault Megane) are included in the \textit{SimScene}'s sound database (50 audio files totally).
+The recordings has been made for 4 cars (Renault Scenic, Renault M\'egane, Renault Clio and Dacia Sandero),  at different speeds and gear ratios.
+Overall, 103 car passages have been recorded.
+\ml{ajouter une description sur les problématiques d'overfitting et de generalisation.
+Cela doit etre un rappel d'un texte ecrit dans la partie nmf.
+A ce titre, cela pourrait etre une extension de l'etude, on en reparle} The audio samples of the first two cars (Renault Scenic and Renault M\'egane) are included in the \textit{SimScene}'s sound database (50 audio files in total).
 The last 53 audio samples are dedicated to the dictionary design as part of NMF, see section \ref{part:dictionary_building}.
-A full description of the recording can be found in \cite{gloaguen_creation_2017}.
+A full description of the recordings can be found in \cite{gloaguen_creation_2017}.
 
@@
 \centering
-\caption{$mTIR$ for each sound environment.}
+\caption{$mTIR$ for each sound environment.\ml{les captions doivent etre self-contained, pas d'acronymes non evident}}
 \begin{tabular}{lc}
@@
 
-With this built-up database, the \textit{SimScene} software and the annotations of the recordings, 74 simulated sound scenes are generated, which have the same temporal structure than the recordings.
+With this built-up database, the \textit{SimScene} software and the annotations of the recordings, 74 simulated sound scenes are generated, which have the temporal structure of the reference recordings.
 The \textit{ebr} parameter is adjusted manually on each sound scene to be faithful compared to the recorded scenes.
@@
 It expresses, on all the scenes of an sound environment, the mean difference between the equivalent traffic sound levels of each scene, $L_{p,traffic}$, with the sound level of the \textit{interfering} sound class, $L_{p, interfering}$, which gathered all the other sound sources not related to the traffic, see Eq \ref{eq:mTIR}.
-It quantifies the predominance of the traffic component on the 4 sound environments.
+It quantifies the predominance of the traffic component for the 4 types of sound environments.
 
@@
 
-To validate its realism, the corpus of transcribed urban sound scenes is submitted to a perceptual test.
+\subsection{Perceptual test}
 
-\subsection{Perceptual test}
-The perceptual test is conducted with a panel of 50 listeners that are asked to assess the level of realism on a 7-point scale (1 is \textit{not realistic at all}, 7 is \textit{very realistic}) of a mix of transcribed and recorded scenes.
+To evaluate the level of realism of the corpus of transcribed urban sound scenes, a perceptual test is considered.
+
+The perceptual test is conducted with a panel of 50 subjects that are asked to assess the level of realism on a 7-point scale (1 is \textit{not realistic at all}, 7 is \textit{very realistic}) of a set of transcribed and recorded scenes.
 The total number of sound scenes tested is set at 40.
 The first half includes 20 30-seconds audio files, including 5 scenes that belong to the sound environment \textit{park}, 6 from \textit{quiet street}, 4 from \textit{noisy street} and 5 from \textit{very noisy street} chosen randomly among the recorded scenes.
-The second half is composed of the same 30-second transcribed scenes.
-In order to limit the duration of the test and to preserve the concentration of the listeners, each of them listens a subset of 20 sound scenes, which mix recorded and transcribed samples.
-Furthermore, all the scenes are normalized to the same sound level, chosen at 65 dB, to prevent the listeners from changing the sound level of their speakers.
+The second half is composed of the transcribed scenes coresponding to the 20 reference scenes.
+In order to limit the duration of the test to preserve the concentration of the subjects, each subject listens to only a subset of 20 sound scenes.
+All the scenes are normalized to the same sound level, chosen at 65 dB, and the subjects are not allowed to change the output sound level once set at the beginning of the experiment.
 
@@
 
-The test was administered online on the 8 February 2017 and the number of participant has been reached 12 days later.
+The test was available online on the 8 February 2017 and the needed number of participant has been reached 12 days later.
 During the test, the participants had the possibility to listen to each scene as many times as wanted before assessing, without being able to change their judgment afterwards.
-The participants could also leave a comment on each audio to explain the rating.
+The participants could also leave a comment on each audio to explain the rating \footnote{The interface of the experiment is available \url{http://soundthings.org/research/xpRealism}}.
 Based on the information provided, the panel of 50 listeners was made of 31 males and 18 females (one not documented) with an average age of 36 ($\pm$ 12) years old.
@@
 \includegraphics[width=\linewidth]{figures/testPerceptif_boxplotType_EN.pdf}
-\caption{Box-and whiskers plot of the rating of realism according to the type (recorded /transcribed).}
+\caption{Box and whiskers plot of the rating of realism according to the type (recorded /transcribed). \ml{Il faut tout expliciter, as tu vraiment besoin de toutes ces informations ?}}
 \label{fig:boxPlot_test}
@@
 The mean score for the transcribed scenes is even superior to the recorded ones ($m_{trans.}$ = 5.1 ($\pm 1.6$), $m_{rec}$ = 4.9 ($\pm 1.6$)).
-These results confirm that all the recorded and the transcribed scenes are perceived in a similar way by the panel.
+These results confirm that the recorded and the transcribed scenes are perceived in a similar way by the panel.
 More details on the results can be found in  \cite{gloaguen_creation_2017}.
 
+\ml{ce serait quand meme bien d'avoir un test stat, au moins un ttest}
+
 As the transcribed sound corpus is sufficiently close to the recordings, these sound mixtures can be used to assess the performances of NMF to estimate the traffic sound level.
+
+UPTO
 

### 2018-03-09 paper/gloaguenAppliedNMF.tex
done proofreading

@@
 
-UPTO
-
 \section{Experimental protocol}\label{part:expProtocol}
 
-The experiment protocol consists in estimate, on the 74 scenes available classified on 4 sound environments, the equivalent sound level of the traffic of the entire scenes, $\tilde{L}_{p, traffic}$ (dB) and to compare them to their exact values given by the simulation process, $L_{p,traffic}$, see Figure \ref{fig:bloc_diagram_estimator}.
+The design of the experiment aims at evaluating in a meaningful manner the performance of the different systems.
+To do so, the 74 sound scenes, organized on 4 different types of sound environments are iteratively fed to the processing system which output an estimate of the equivalent sound level of the traffic within each scene $\tilde{L}_{p, traffic}$ (dB).
+This result is then compared to the reference sound level value given by the simulation process, $L_{p,traffic}$, see Figure \ref{fig:bloc_diagram_estimator}.
 
@@
 
+\subsection{Metrics}
+
+The traffic sound levels, $\tilde{L}_{p,traffic}$, are compared to the exact values, $L_{p,traffic}$, through the Mean Absolute Error ($MAE$) \cite{willmott2005advantages}.
+The $MAE$ expresses the quality of the long-term reconstruction of the signal and consists in the average of the absolute difference between the exact and the estimated sound levels,
+
+\begin{equation}
+MAE = \frac{\sum_{i = 1}^{M} \vert L_{p,traffic}^i - \tilde{L}_{p,traffic}^i \vert}{M}.
+\end{equation}
+
+It is then possible to express the $MAE$ error for each unique setting but, also, to average this metric according to the 4 sound environments to be able to estimate the optimal setting that offers the lowest error for all the sound environments:
+
+\begin{equation}
+mMAE = \frac{\sum_{i = 1}^4 MAE_i}{4},
+\end{equation}
+
+where the other experimental factors (method, $f_c$, $\mathbf{K}$, $\mathbf{w_t}$, $\beta$, threshold value $\mathbf{t_h}$) are fixed.
+
 \begin{figure*}[t]
@@
 \caption{Dictionary building on a 3 second example of a passing car with $w_t = 1$ second.
-The dictionary are cut in 3 frames of $w_t$ duration.
-On each the rms value is calculated to obtained 3 spectra destined to $\mathbf{W}$.}
+In this case, the dictionary $\mathbf{W}$ is made of 3 spectra, each representative of a texture frame of $w_t$ duration.}
 \label{fig:example_dictionary}
@@
 
-As the road traffic is mainly composed of a low frequency content, a first estimator is considered through a frequency low-pass filter (LP filter).
+\subsection{Methods}
+
+As the road traffic is mainly composed of a low frequency content, a frequency low-pass filter (LP filter) is considered as baseline.
 It consists in filtering the sound scenes at different cut-off frequencies $f_c \in \lbrace$500, 1, 2k, 5k, 10k, 20k$\rbrace$ Hz.
@@
 The second estimator is based on the three NMF formula presented in part \ref{part:nmf} (see Figure \ref{fig:bloc_diagram_nmf}).
-Between the dictionary building to the metric calculation, multiple experimental factors take part in these cases where each of them having different modalities.
+Multiple experimental factors take part in these cases where each of them having different modalities.
 
@@
 
-\subsection{Dictionary building}\label{part:dictionary_building}
-The dictionary building is designed from a second sound database specially dedicated to this task to prevent any overfitting issues.
-It contains the 53 audio files of the 2 other cars (Dacia Sandero and Renault Clio) recorded on the runway, see part \ref{part:simScene}.
+\subsection{Dictionary building for NMF methods}\label{part:dictionary_building}
+
+The dictionary is designed from a second sound database specially dedicated to this task to prevent any overfitting issues.
+It contains  53 audio files of the 2 cars (Dacia Sandero and Renault Clio) recorded on the runway, see part \ref{part:simScene}.
+Those recordings have not been considered for the creation of the evaluation dataset.
 
@@
 From the 53 audio files, we obtain respectively for $\mathbf{w_t}\in$ $\lbrace$0.5, 1$\rbrace$ second, 2218 and 1109 elements.
-A $\mathbf{K}$-mean clustering algorithm is applied to reduce these dimensions to $\mathbf{K} \in \lbrace$25, 50, 100, 200$\rbrace$ in order to avoid redundant information and decrease the computation time.
-The obtained $\mathbf{K}$ clusters compose then the elements of $\mathbf{W}$.
+
+A $\mathbf{K}$-means clustering algorithm is applied to reduce these dimensions to $\mathbf{K} \in \lbrace$25, 50, 100, 200$\rbrace$ in order to avoid redundant information and decrease the computation time.
+The resulting $\mathbf{K}$ centroids are taken as the elements of $\mathbf{W}$.
+\ml{Vraiment utile ?
 Furthermore, the case where each audio generates one spectrum from its spectrogram is added ($\mathbf{w_t} = all$).
-Here, by the added approach, the dictionary elements are based on the spectral envelops of the audio samples.
-For this case, having 53 audio samples, the number of elements in $\mathbf{W}$ with the $\mathbf{K}$-mean clustering algorithm is reduced to $\mathbf{K} \in \lbrace$25, 50$\rbrace$.
+In this case, the dictionary elements are based on the spectral envelope of the audio samples.
+With only 53 audio samples, the number of elements in $\mathbf{W}$ with the $\mathbf{K}$-mean clustering algorithm is reduced to $\mathbf{K} \in \lbrace$25, 50$\rbrace$.}
 
-The obtained dictionary is expressed in third octave bands to decrease the dimensionality without deteriorate the spectral content.
-A previous experimental validation revealed that the use of third octave bands, instead of linear spectra, does not impact the performances of the chosen estimator in this study.
-Finally, each basis of $\mathbf{W}$ is normalized as $ \Vert \mathbf{w}_k \Vert = 1$ where $\Vert \bullet \Vert$ is the $\ell_1$ norm.
+The obtained dictionary is expressed in third octave bands to decrease the dimensionality.
+\ml{Citer ou enlever, A previous experimental validation revealed that the use of third octave bands does not impact the performances of the chosen estimator in this study.}  Each basis of $\mathbf{W}$ is then normalized as $ \Vert \mathbf{w}_k \Vert = 1$ where $\Vert \bullet \Vert$ is the $\ell_1$ norm.
 Table \ref{tab:experimental_factorsNMF} summarizes the different modalities of the two experimental factors ($\mathbf{K}$ and $\mathbf{w_t}$).
@@
 NMF is performed for 3 $\beta$-divergences: $\beta$ = 2 (euclidean distance), $\beta$ = 1 (Kullback-Leibler divergence) and $\beta = 0$ (Itakura-Sa\"ito divergence).
-The spectrogram $\mathbf{V}$ and the dictionary $\mathbf{W}$ are displayed in a logarithmic scale through a third band octave representation that reduces the high frequency predominance where the traffic component is absent.
+The spectrogram $\mathbf{V}$ and the dictionary $\mathbf{W}$ are represented in a logarithmic scale through a third band octave representation that reduces the high frequency predominance where the traffic component is absent.
 In addition, as the number of frequency bins is reduced ($F$ = 29), the computation time is reduced too.
 400 iterations are performed to get a stabilized results.
-For SEM-NMF, the number of elements in $\mathbf{W_r}$ is fixed at $J = 2$.
-For hard thresholding, the threshold value, $\mathbf{t_h}$ is defined between 0.30 and 0.60 with a 0.01 step.
-Each unique association of modalities between each experimental factor forms a setting.
+For SEM-NMF, the number of elements in $\mathbf{W_r}$ is set to $J = 2$.
+For hard thresholding, the threshold value, $\mathbf{t_h}$ is set between 0.30 and 0.60 with a 0.01 increment.
+Each unique association of modalities between each experimental factor forms an experimental setting.
 For the filter estimator, 24 settings are computed (4 $\times$ 6).
 For SUP and SEM-NMF, 240 settings are computed (4 $\times$ 2 $\times$ (2 $\times$ 4 + 1 $\times$ 2) $\times$ 3).
-Finally, for TI-NMF, the number of settings is much higher (3720) due to the multiple possible combinations between the threshold values (4 $\times$ 1 $\times$ (2 $\times$ 4 + 1 $\times$ 2) $\times$ 3 $\times$ 31).
-The summarize of the experimental factors and their different modalities is displayed on Tables \ref{tab:experimental_factorsFilter} and \ref{tab:experimental_factorsNMF}.
+Finally, for TI-NMF, the number of settings is much higher (3720) due to the higher cardinality of the set of threshold values (4 $\times$ 1 $\times$ (2 $\times$ 4 + 1 $\times$ 2) $\times$ 3 $\times$ 31).
+The experimental factors and their different modalities are displayed on Tables \ref{tab:experimental_factorsFilter} and \ref{tab:experimental_factorsNMF}.
 
-The approximated traffic spectrograms $\mathbf{\tilde{V}}_{traffic}$ are obtained after 400 iterations. The estimated traffic sound level in dB, $\tilde{L}_{p, traffic}$, is deducted,
+The approximated traffic spectrograms $\mathbf{\tilde{V}}_{traffic}$ are obtained after 400 iterations. The estimated traffic sound level in dB, $\tilde{L}_{p, traffic}$, is then computed,
 
@@
 \begin{table*}[t]
-\caption{Summary of the different experimental factors and their modalities taken into account in the frequency low-pass filter estimator.}
+\caption{Experimental factors and their modalities for the frequency low-pass filter baseline estimator.}
 \centering
@@
 \centering
-\caption{Summary of the different experimental factors and their modalities taken into account in NMF estimator.}
+\caption{Experimental factors and their modalities for the NMF estimator.}
 \begin{tabularx}{17.5cm}{L{3cm}@{}C{12cm}@{}C{2cm}@{}}
@@
 
-\subsection{Metrics}
-
-The traffic sound levels, $\tilde{L}_{p,traffic}$, are compared to the exact values, $L_{p,traffic}$, through the Mean Absolute Error ($MAE$) \cite{willmott2005advantages}.
-The $MAE$ expresses the quality of the long-term reconstruction of the signal and consists in the average of the absolute difference between the exact and the estimated sound levels,
-
-\begin{equation}
-MAE = \frac{\sum_{i = 1}^{M} \vert L_{p,traffic}^i - \tilde{L}_{p,traffic}^i \vert}{M}.
-\end{equation}
-
-It is then possible to express the $MAE$ error for each unique setting but, also, to average this metric according to the 4 sound environments to be able to estimate the optimal setting that offers the lowest error for all the sound environments:
-
-\begin{equation}
-mMAE = \frac{\sum_{i = 1}^4 MAE_i}{4},
-\end{equation}
-
-where the other experimental factors (method, $f_c$, $\mathbf{K}$, $\mathbf{w_t}$, $\beta$, threshold value $\mathbf{t_h}$) are fixed.
-
 \section{Results and discussion}\label{part:results}
@@
 The lowest error for a LP filter is obtained with $f_c$ = 500 Hz ($mMAE =$ 2.14 ($\pm$ 1.83) dB).
-It allows a balance between the discarded and remaining energy through the sound environments.
+\ml{It allows a balance between the discarded and remaining energy through the sound environments.}
 
-When considering all the sound scenes, SUP-NMF does not succeed to have a lower error than the 500 Hz LP filter for all the $\beta$ values.
+When considering all the sound scenes, SUP-NMF does not succeed to achieve a lower error than the 500 Hz LP filter for all the $\beta$ values.
 By adding the mobile part $\mathbf{W_r}$ in the dictionary, SEM-NMF with $\beta = 0$ and $\beta = 1$ allows a lower error than 500 Hz LP filter with a reduced standard deviation especially for $\beta = 1$ ($mMAE = 1.94$ ($\pm$ 0.38) dB).
@@
 
-From these global results, the $MAE$ errors are compared to the LP filter and each method for the 4 sound environments, see Figure \ref{fig:mae_env}.
+From these global results, the $MAE$ errors are compared to the LP filter and each method for the 4 types of sound environments, see Figure \ref{fig:mae_env}.
 
@@
 \includegraphics[width=\linewidth]{figures/mea_grafic_bar.pdf}
-\caption{$MAE$ errors according each sound environment for the best combination of the LP filter ($f_c$ = 500 Hz), SUP-NMF ($\beta$ = 2, $\mathbf{K}$ = 25, $\mathbf{w_t}$ = 1 second), SEM-NMF ($\beta$ = 1, $\mathbf{K}$ = 200, $\mathbf{w_t}$ = 1 second) and TI-NMF ($\beta$ = 2, $\mathbf{K}$ = 200, $\mathbf{w_t}$ = 0.5 second, $\mathbf{t_{h}} = 0.32$).}
+\caption{$MAE$ errors according to each sound environment for the best combination of the LP filter ($f_c$ = 500 Hz), SUP-NMF ($\beta$ = 2, $\mathbf{K}$ = 25, $\mathbf{w_t}$ = 1 second), SEM-NMF ($\beta$ = 1, $\mathbf{K}$ = 200, $\mathbf{w_t}$ = 1 second) and TI-NMF ($\beta$ = 2, $\mathbf{K}$ = 200, $\mathbf{w_t}$ = 0.5 second, $\mathbf{t_{h}} = 0.32$).}
 \label{fig:mae_env}
@@
 
-Except SEM-NMF, all the methods show the same error evolution the same error evolution: a decrease of the error with the increase of the traffic predominance.
-SEM-NMF show an almost constant error for all 4 sound environments.
+Aside SEM-NMF, all the methods show the same error evolution the same error evolution: a decrease of the error with the increase of the traffic predominance.
+On contrary, SEM-NMF show an almost constant error for all 4 sound environments.
 The LP filter error is mainly important for environments where the traffic is less present.
@@
 
-Despite a fixed dictionary composed of traffic spectra, SUP-NMF fails to identified correctly the traffic component particularly for \textit{park} ($MAE$ = 6.42 dB) environments.
+Despite a fixed dictionary composed of traffic spectra, SUP-NMF fails to identify correctly the traffic component particularly for \textit{park} ($MAE$ = 6.42 dB) environments.
 With this method, as NMF minimizes the cost function, eq.
@@
 
-Besides having the lowest $mMAE$ error, TI-NMF presents the most performing results.
-The park environment is the only case where an other NMF approach out-performed significantly TI-NMF ($MAE$ = 2.95 dB).
+\ml{Pas clair, enlever ?
+: Besides having the lowest $mMAE$ error, TI-NMF presents the most performing results.
+The park environment is the only case where an other NMF approach out-performed TI-NMF ($MAE$ = 2.95 dB).
 In this sound environment, the traffic dictionary is then composed, in average, of more than half the 200 elements in $\mathbf{W'}$ (136 elements).
@@
 In this case, in average, $\mathbf{W'}$ is composed of 198 traffic elements.
-This method out-performed SUP-NMF since, as $\mathbf{W_0}$ is updated, the final dictionary $\mathbf{W'}$ is then most suited to the sound scene than a fixed dictionary.
-The advantage to have an unique dictionary suited to each sound scene makes TI-NMF very performing when traffic is predominant while, by the thresholding, it limits the dictionary derivation when the traffic is less present.
+This method out-performed SUP-NMF since, as $\mathbf{W_0}$ is updated, the final dictionary $\mathbf{W'}$ is then most suited to the sound scene than a fixed dictionary.}
+
+Considering a unique dictionary fitted to the sound scene under evaluation thus makes TI-NMF very effective when traffic is predominant while, by the use of the thresholding, it discards the elements of the dictionary that deviated too much when the traffic is less present.
 
@@
 
-A non-negative matrix factorization frame has been applied as a source separation tool to estimate the traffic sound level  from a corpus of urban sound scenes artificially built.
-In addition, the realism of the scenes has been verified thanks to a perceptual test.
-The proposed protocol generate audio sample similar than the outputs of a deployed sensor networks with the advantages of the simulation process (sound level and position of each source controlled and known).
+The non-negative matrix factorization framework has been considered as a source separation tool to estimate the traffic sound level  from a corpus of urban sound scenes artificially built.
+Those scenes are designed to be as similar as possible to the outputs of a deployed sensor networks with the advantages of the simulation process (sound level and position of each source controlled and known).
+The realism of the scenes has been verified thanks to a perceptual test.
 
-The results confirm the interest of this method on such sound environments.
-It easily takes into account the overlap between the multiple sound sources present in cities and is suited to monophonic sensor networks.
-Different NMF formula have been studied through the supervised and semi-supervised approach.
-On all the sound environments, these common approaches reveal to be not sufficiently efficient: supervised NMF approach, with its fixed dictionary, does not succeed to estimate correctly the traffic sound level especially when this sound source is quiet, while semi-supervised approach with the presence of a mobile part in the dictionary is the best estimator for \textit{park} environments but failed on heavily traffic scenes.
-Finally, the proposed approach, namely Thresholded Initialized NMF, achieved the lowest error with an initialized updated dictionary.
-By considering the elements the most similar to traffic spectra with a hard thresholding, it makes it possible to adapt specifically the dictionary to the sound scenes.
-Consequently, in the case where the emplacement or the sound environment of the sensors cannot be identified (for instance within a mobile measurement framework), TI-NMF, with $\beta = 2$, $\mathbf{K}$ = 200, $\mathbf{w_t} = 0.5$ second, $\mathbf{t_{h}} = 0.32$, seems the best generic framework, for the sound corpus studied.
-But if the sound environment can be identified through a prior analysis, or based on positioning data \cite{can2015noise,lavandier2016urban}, it seems possible to adapt the proposed work by selecting the most efficient approach (for instance SEM-NMF within parks), in order to further reduce the error in the estimated road traffic sound levels.
-Further analyses are required to extend the proposed method to other sound sources, such as birds or voices sounds, by replacing or adding elements in the built dictionary.
+The results confirm the potential of the NMF method on such application scenario as it takes into account the overlap between the multiple sound sources present in cities and is suited to monophonic sensor networks.
+Different NMF algorithmic schemes have been studied through the supervised and semi-supervised approach.
+On all the sound environments, these common approaches reveal to be not sufficiently efficient: supervised NMF approach, with its fixed dictionary, does not succeed to estimate correctly the traffic sound level especially when this sound source is quiet, while semi-supervised approach with the presence of a mobile part in the dictionary is the best estimator for \textit{park} environments but fails on heavily traffic scenes.
+
+The proposed approach, named Thresholded Initialized NMF, achieved the lowest error in the evaluation corpus.
+Consequently, in the case where the location or the type of sound environments the sensors are monitoring cannot be identified (for instance within a mobile measurement framework), TI-NMF appears to be the method of choice.
+If the sound environment can be identified through a prior analysis, or based on positioning data \cite{can2015noise,lavandier2016urban}, it should be possible to adapt the estimation procedure by selecting the most efficient approach in order to further reduce the error in the estimated road traffic sound levels.
+
+Further analyses are required to extend the proposed method to other sound sources, such as birds or voices sounds, which can conveniently be done by replacing or adding elements in the dictionary.
 This would prove useful in the context of muti-source noise mapping that is  gaining interest \cite{aumond2017Probabilistic, aletta2015soundscape}.
 Finally, the selected parameters stand for the corpus of sound mixtures of this study.
-Further analyses on various corpus of sound scenes are needed to test the robustness of the method, and select the most relevant approaches for specific sound environments (predominance of water or industrial sounds, rural environments \dots).
+Further analyses on various corpus of sound scenes are needed to evaluate the robustness of the method, and select the most relevant approaches for specific sound environments (predominance of water or industrial sounds, rural environments \dots).
 
-For reproducible purposes, the experimental protocol and the programs developed under the Matlab software are available online.
-This study proposes also a realistic urban sound corpus that could be greatly appreciate for research communities dedicated to the detection, identification or source separation tasks.
+For reproducibility purposes, the evaluation corpus, the experimental protocol and the programs developed under the Matlab software are available online \footnote{\url{https://github.com/jean-remyGloaguen/articleApplied2018}}.
+%This study proposes also a realistic urban sound corpus that could be greatly appreciate for research communities dedicated to the detection, identification or source separation tasks.
 

### 2018-03-12 paper/articleNmfTrafficSimScene2018.tex
ML corrections included

--- 2018-03-09 paper/gloaguenAppliedNMF.tex
@@
 Many applications can be found for musical \cite{smaragdis_non-negative_2003, benetos2006musical} and speech \cite{wilson_speech_2008, mysore2011non} contents.
-Dedicated to sound separation, Immani and Kasa\"i \cite{satoshi_innami_nmf-based_2012} used NMF in a two steps sound separation with the help of time variant gain features.
-A first study \cite{gloaguen2018Estimation} has been conducted, in which diverse NMF estimation rules are compared, namely the supervised, the semi-supervised, and the threshold initialized NMF, have been applied on a large set of simulated sound scenes that which mixed traffic component with specific urban sounds at calibrated sound levels.
-The study demonstrated the interest of NMF for urban sound environments and compared the benefits of each approach on a validation corpus, designed to .
-However, it has now to face to real urban sound scenes.
+Dedicated to sound separation with environmental sound, Immani and Kasa\"i \cite{satoshi_innami_nmf-based_2012} used NMF in a two steps sound separation with the help of time variant gain features.
+A first study \cite{gloaguen2018Estimation} has been conducted, in which diverse NMF estimation rules are compared, namely the supervised, the semi-supervised, and the threshold initialized NMF, have been applied on a large set of simulated sound scenes.
+This corpus mixes 6 sound categories (\textit{alert, animals, climate, humans, mechanics, transportation}) with a traffic component calibrated to different sound levels to obtain variable traffic predominance.
+The diversity of this corpus was made to assess the performances and the limits of each NMF formula. 
+However, if it reveals the interest of NMF for urban sound environments, the assessment of its performance on a corpus of realistic sound scenes must be carried out in order to implement it on a sensor network.
+Design realistic urban sound mixtures makes it possible to access to many acoustic proprieties as the onset and offset time and the sound level of each sound class and especially the traffic component.
+However, like all simulated process, the issue of its realism has to be resolved in order to validate the NMF performances.
 
-\ml{la fin de ce par est a revoir car c'est le point crucial du papier.
-Bien expliquer le corpus du papier d'avant, son usage, les questions qui restent en suspens et en quoi on passe ici a l'étape superieure en y repondant.}
-
-In this paper, the NMF framework is applied on a corpus of simulated sound scenes, generated based on annotated urban recordings, and whose degree of realism is validated through a perceptual test.
-The NMF framework and its different optimization strategies are described in section \ref{part:nmf}.
+In this paper, a realistic urban sound corpus based on annotated urban recordings, and whose degree of realism is validated through a perceptual test, is designed in order to estimate the traffic sound level with the help of the NMF framework.
+The different NMF approaches are described in section \ref{part:nmf}.
 Next, the corpus of urban sound scenes is presented in section \ref{part:urban_scene}, from the sound database built-up to its validation.
@@
 To be completely successful, all the acoustical sources must be considered in it which is not possible in an urban environment.
-However, in our application setting, the main source of interest, \textit{i.e.} the traffic, can be known \textit{a priori} \ml{manque une phrase} semi-supervised NMF (SEM-NMF) \cite{lee_semi-supervised_2010} is considered to better take into account the interfering sound sources.
-To do so, $\mathbf{W}_{F \times (K+J)}$ is decomposed into two distinctive matrices: $\mathbf{W} = \left[ \mathbf{W_s}~\mathbf{W_r} \right]$ where $\mathbf{W_s}_{F \times K}$ is a fixed part of $\mathbf{W}$ composed of audio spectra and $ \mathbf{W_r}_{F \times J}$, a mobile part which is updated, see eq.
+To overcome this issue, semi-supervised NMF (SEM-NMF) \cite{lee_semi-supervised_2010} is considered.
+It consists in decomposing, $\mathbf{W}_{F \times (K+J)}$ into two distinctive matrices: $\mathbf{W} = \left[ \mathbf{W_s}~\mathbf{W_r} \right]$ where $\mathbf{W_s}_{F \times K}$ is a fixed part of $\mathbf{W}$ composed of audio spectra and $ \mathbf{W_r}_{F \times J}$, a mobile part which is updated, see eq.
 \ref{eq:W_r_SS}.
@@
 
-In this study, $\mathbf{W_s}$ is composed of traffic audio spectra to include in $\mathbf{W_r}$ all the other sources that can be present in the urban sound scenes.
+In this study, $\mathbf{W_s}$ is composed of traffic audio spectra, as it is the sound source of interest, to include in $\mathbf{W_r}$ all the other sources that can be present in the urban sound scenes.
 The traffic signal estimation is next defined by the fixed part,
@@
 where $\mathbf{w}$ is a $k$ element of $\mathbf{W}$ of $F \times 1$ dimension.
-When $D_{\theta}\left(\mathbf{w_0} \Vert \mathbf{w'} \right)$=1, the element $k$ from $\mathbf{W'}$ is equal to the in $\mathbf{W_0}$ \ml{pas clair, a quoi sert le k ?}.
-If $D_{\theta}\left(\mathbf{w_0} \Vert \mathbf{w'} \right)$=0, the element is fully different.
-Next, the similarities are sorted in descending order \ml{le sort n'est pas necessaire, le seuil suffit}.
+When $D_{\theta}\left(\mathbf{w_0} \Vert \mathbf{w'} \right)$=1, both element are identicals.
+If $D_{\theta}\left(\mathbf{w_0} \Vert \mathbf{w'} \right)$=0, the element $\mathbf{w'}$ is fully different.
 The extraction of traffic elements in $\mathbf{W'}$ is carried out by a hard thresholding method \cite{donoho1994threshold}.
@@
  0 & else.
-\end{numcases}
+\end{numcases}\\
 \end{subequations}
 
-\ml{faire un par resumant les prop des trois methodes avec un accent sur l'utilisation de la connaissance a priori.}
+To sum-up, to approximate the audio spectrogram, $\mathbf{V}$ and estimate the traffic component, 3 methods are used.
+SUP-NMF is only based on the \textit{a priori} knowledge of the traffic to built the dictionary and is constrained to used it to obtain $\mathbf{\tilde{V}}_{traffic}$.
+SEM-NMF completes this dictionary by the add of a mobile dictionary to consider the interfering sound events and then offer more flexibility.
+Finally, TI-NMF, as SUP-NMF, considers first a dictionary composed of \textit{traffic} spectra to then update them.
+Therefore, it is the entire dictionary that can be fully generalized to the sound scenes.
+The initiated dictionary allows the focused on the source of interest during the updates.
 
-These methods are applied on simulated sound scenes in order to compare the estimated sound levels with a non ambiguous reference.
-In \cite{gloaguen2018Estimation}, the sound corpus was composed of a mix of traffic samples with specific sound classes (\textit{alert, animals, climate, human, mechanics, transportation}).
-Here, in order to implement this method in embedded sensors, a new and more realistic sound corpus by considering more realistic urban sound environments with many kind of sound sources.
+These above described methods are applied on an evaluation corpus,  composed of simulated sound scenes, in order to compare the estimated sound levels with a non ambiguous reference.
+This new and more realistic sound corpus is designed by considering realistic urban sound environments with many kind of sound sources mixed.
 
@@
 
-\ml{Il faut faire une figure et un paragraphe resumant la figure.
-Voici un essai (je pense d'ailleurs que cela devrait faire partie des claims dans l'intro) : In order to validate the above described methods, the evaluation corpus is both required to be realistic and non ambiguous in terms of traffic sound level.
-The former calls for the use recordings of urban sound environments, and the latter imposes the use of controled use of sound sources with known properties: onset and offset time location, and volume.
+\begin{figure*}[t]
+\centering
+\includegraphics[width=0.7\linewidth]{figures/realistic_urban_sound_scene_design.pdf}
+\caption{Bloc diagram of the design of evaluation corpus}
+\label{fig:transcription}
+\end{figure*}
+
+The evaluation corpus is both required to be realistic and non ambiguous in terms of traffic sound level.
+The former calls for the use of recordings of urban sound environments, and the latter imposes the use of controlled sound sources with known properties: onset and offset time location, and sound level.
 Indeed, the precise annotation of the sound level of a given source of interest in a polyphonic mixture is not feasible.
-We thus propose to design such evalution corpus based on two corpora.
-The first, called the reference corpus, is a set of recordings of urban sound environments, from which a sort of "score" is extracted manually, \textit{i.e.} which sound source is present from this time to this time, at which sound level.
+We thus propose to design such evaluation corpus based on two corpora.
+The first, called the reference corpus, is a set of recordings of urban sound environments, from which a sort of transcription is made, \textit{i.e.} which sound source is present from this time to this time, at which sound level.
 The second, called the elemental corpus, is a set of monophonic recordings of sound sources that are occurs in the scenes of the reference.
-The recordings of the elemental corpus are not taken from the reference corpus.
-The evaluation corpus is then built by sequencing recordings of the elemental corpus according to the scores of the reference corpus.
-(Le reste est a reecrire avec les termes proposés)}
+The elemental corpus is built-up with still existing sound and with specific recordings, see part \ref{part:simScene}.
+The evaluation corpus is then built by replicating recordings of the elemental corpus according to the transcript of the reference corpus, see Figure \ref{fig:transcription}.
 
@@
 Two of the 76 recordings are rejected for the analysis because the audio files were corrupted, resulting in 74 valid audio files assumed as representative of the variety of sound environments.
-The recordings are listened and categorized within four different sound environments, as proposed in \cite{can_describing_2015}: park (P, 8 audio files with a cumulative duration of 16min01), quiet street (Q, 33 audio files with a cumulative duration of 77min27), noisy street (N, 24 audio files with a cumulative duration of 56min10) and very noisy street (vN, 9 audio files with a cumulative duration of 21min42).
-Then, each audio file is annotated, annotating the start, end time \ml{and level ?} of each sound event along with its sound class.
-\ml{cette phrase ne dis probablement pas ce qsue tu veux dire annotation et Transcription sont des termes equivalents, The aim of the annotation phase is next to transcribe the recordings}, in order to obtain simulated sound scenes with the same \ml{ca peut preter a confusion en implicitant une distribution statistique : distribution} of sound events as the recordings and therefore as close as possible to the realistic scenes.
+The recordings are listened and categorized within four different sound environments, as proposed in \cite{can_describing_2015}: \textit{park} (P, 8 audio files with a cumulative duration of 16min01), \textit{quiet street} (Q, 33 audio files with a cumulative duration of 77min27), \textit{noisy street} (N, 24 audio files with a cumulative duration of 56min10) and \textit{very noisy street} (vN, 9 audio files with a cumulative duration of 21min42).
+Then, each audio file is annotated, noticing the start, end time and level of each sound event along with its sound class.
+This annotation phase makes it possible to obtain simulated sound scenes with the same positions of sound events than the recordings and therefore as close as possible to the realistic scenes.
 
@@
 This software have already been used in a wide range of experiments for sound detection algorithm assessment \cite{lafay_new_2014, benetos2016detection}.
-\ml{Pas clair The control of high level parameters can be handled by the user as the sound class occurence, the time between each sample of one sound class, the ratio between the sound level of an event class with the background (i.e the \textit{event background ratio} shorten \textit{ebr})\dots Each parameter is completed with a standard deviation to bring random behavior.
-It allows too the design of a sound mixture from an annotation text file.
-A remplacer par ceci: The \textit{SimScene} software allows for several type of sequencing, from abstract ones (time indexes and amplitudes are drawn from random distributions) to  precise ones, where the time indexes and amplitudes for each event is set by the user.
-The latter type is considered in this study.}
+The \textit{SimScene} software allows for several type of sequencing, from abstract ones (time indexes and amplitudes are drawn from random distributions) to  precise ones, where the time indexes and amplitudes for each event is set by the user.
+The latter type is considered in this study.
 As output, \textit{SimScene} generates an audio of the global sound mixture and an audio for each sound class present in the scene, which makes it possible to know their exact contributions in the scene, see Figure \ref{fig:example_simScene}.
@@
 Cela doit etre un rappel d'un texte ecrit dans la partie nmf.
-A ce titre, cela pourrait etre une extension de l'etude, on en reparle} The audio samples of the first two cars (Renault Scenic and Renault M\'egane) are included in the \textit{SimScene}'s sound database (50 audio files in total).
+A ce titre, cela pourrait etre une extension de l'etude, on en reparle}.
+In order to avoid overfitting issues, the audio samples of the first two cars (Renault Scenic and Renault M\'egane) are included in the \textit{SimScene}'s sound database (50 audio files in total).
 The last 53 audio samples are dedicated to the dictionary design as part of NMF, see section \ref{part:dictionary_building}.
@@
 \centering
-\caption{$mTIR$ for each sound environment.\ml{les captions doivent etre self-contained, pas d'acronymes non evident}}
+\caption{Mean \textit{Traffic Interfering Ratio} for each sound environment.}
 \begin{tabular}{lc}
  & $mTIR$ (dB)\\ \hline
- park & -9.10 ($\pm$ 7.35) \\
- quiet street & 0.88 ($\pm$ 5.92) \\
- noisy street & 6.96 ($\pm$ 5.16) \\
- very noisy street & 15.75 ($\pm$ 9.78) \\ \hline
+ \textit{park} & -9.10 ($\pm$ 7.35) \\
+ \textit{quiet street} & 0.88 ($\pm$ 5.92) \\
+ \textit{noisy street} & 6.96 ($\pm$ 5.16) \\
+ \textit{very noisy street} & 15.75 ($\pm$ 9.78) \\ \hline
 \end{tabular}
@@
 $62\%$ of the participants declared having no experience in the listening of urban sound mixtures.
-Figure \ref{fig:boxPlot_test} summarizes the score distributions of all the listeners for the recorded (rec.) and the transcribed (trans.) scenes.
 
-\begin{figure}[t]
+A Student's t-test is performed with the $H_0$ hypothesis which considers the similarity between the distribution of the average rates for the recorded and the transcribed scenes of each participant with threshold \textit{p-value} of 5$\%$.
+Results are displayed in Table \ref{tab:ttest_type_of_scene}.
+
+\begin{table}[t]
 \centering
-\includegraphics[width=\linewidth]{figures/testPerceptif_boxplotType_EN.pdf}
-\caption{Box and whiskers plot of the rating of realism according to the type (recorded /transcribed). \ml{Il faut tout expliciter, as tu vraiment besoin de toutes ces informations ?}}
-\label{fig:boxPlot_test}
-\end{figure}
+\begin{tabular}{cccc}
+ 	& \textbf{DOF} & \textbf{\textit{$\mid$ t $\mid$} }   & \textbf{\textit{p-value}} \\ 
+\hline
+ type & 49 & 1.37 &  0.17   \\
+\hline
+\end{tabular}
+\caption{Student's t-test performed on the distribution of the average ratings for the recorded and replicated scenes}
+\label{tab:ttest_type_of_scene}
+\end{table}
 
-The distributions of the notations according to the type are extremely similar.
-The mean score for the transcribed scenes is even superior to the recorded ones ($m_{trans.}$ = 5.1 ($\pm 1.6$), $m_{rec}$ = 4.9 ($\pm 1.6$)).
-These results confirm that the recorded and the transcribed scenes are perceived in a similar way by the panel.
+These results confirm that the recorded and the transcribed scenes are perceived in a similar way by the panel (\textit{p-value} $>$ $\alpha$).
 More details on the results can be found in  \cite{gloaguen_creation_2017}.
-
-\ml{ce serait quand meme bien d'avoir un test stat, au moins un ttest}
 
@@
 The traffic sound levels, $\tilde{L}_{p,traffic}$, are compared to the exact values, $L_{p,traffic}$, through the Mean Absolute Error ($MAE$) \cite{willmott2005advantages}.
-The $MAE$ expresses the quality of the long-term reconstruction of the signal and consists in the average of the absolute difference between the exact and the estimated sound levels,
+The $MAE$ consists in the average of the absolute difference between the exact and the estimated sound levels,
 
@@
 
+and expresses the quality of the long-term reconstruction of the signal.
 It is then possible to express the $MAE$ error for each unique setting but, also, to average this metric according to the 4 sound environments to be able to estimate the optimal setting that offers the lowest error for all the sound environments:
@@
 The resulting $\mathbf{K}$ centroids are taken as the elements of $\mathbf{W}$.
-\ml{Vraiment utile ?
-Furthermore, the case where each audio generates one spectrum from its spectrogram is added ($\mathbf{w_t} = all$).
-In this case, the dictionary elements are based on the spectral envelope of the audio samples.
-With only 53 audio samples, the number of elements in $\mathbf{W}$ with the $\mathbf{K}$-mean clustering algorithm is reduced to $\mathbf{K} \in \lbrace$25, 50$\rbrace$.}
-
 The obtained dictionary is expressed in third octave bands to decrease the dimensionality.
-\ml{Citer ou enlever, A previous experimental validation revealed that the use of third octave bands does not impact the performances of the chosen estimator in this study.}  Each basis of $\mathbf{W}$ is then normalized as $ \Vert \mathbf{w}_k \Vert = 1$ where $\Vert \bullet \Vert$ is the $\ell_1$ norm.
+Each basis of $\mathbf{W}$ is then normalized as $ \Vert \mathbf{w}_k \Vert = 1$ where $\Vert \bullet \Vert$ is the $\ell_1$ norm.
 Table \ref{tab:experimental_factorsNMF} summarizes the different modalities of the two experimental factors ($\mathbf{K}$ and $\mathbf{w_t}$).
@@
 The lowest error for a LP filter is obtained with $f_c$ = 500 Hz ($mMAE =$ 2.14 ($\pm$ 1.83) dB).
-\ml{It allows a balance between the discarded and remaining energy through the sound environments.}
 
@@
 
-\ml{Pas clair, enlever ?
-: Besides having the lowest $mMAE$ error, TI-NMF presents the most performing results.
-The park environment is the only case where an other NMF approach out-performed TI-NMF ($MAE$ = 2.95 dB).
-In this sound environment, the traffic dictionary is then composed, in average, of more than half the 200 elements in $\mathbf{W'}$ (136 elements).
-For the rest of the sound environments, TI-NMF has the lowest error.
+Finally, TI-NMF presents the most performing results, except in the \textit{park} environment ($MAE$ = 2.95 dB).
+In this sound environment, with a threshold value $t_h$ = 0.32, the traffic dictionary is composed, on average, of 136 elements.
+For the rest of the sound environments, TI-NMF has the lowest errors.
 For very noisy environment the error is even very low ($MAE=$ 0.28 dB).
-In this case, in average, $\mathbf{W'}$ is composed of 198 traffic elements.
-This method out-performed SUP-NMF since, as $\mathbf{W_0}$ is updated, the final dictionary $\mathbf{W'}$ is then most suited to the sound scene than a fixed dictionary.}
+With, on average 198 elements in $\mathbf{W}_{traffic}$, almost all the elements present in $\mathbf{W'}$ can be considered as traffic elements.
 
-Considering a unique dictionary fitted to the sound scene under evaluation thus makes TI-NMF very effective when traffic is predominant while, by the use of the thresholding, it discards the elements of the dictionary that deviated too much when the traffic is less present.
+Considering a unique dictionary fitted to the sound scene under evaluation thus makes TI-NMF very effective when traffic is predominant while, by the use of the thresholding, it discards the elements of the dictionary that deviate too much when the traffic is less present.
 
@@
 For reproducibility purposes, the evaluation corpus, the experimental protocol and the programs developed under the Matlab software are available online \footnote{\url{https://github.com/jean-remyGloaguen/articleApplied2018}}.
-%This study proposes also a realistic urban sound corpus that could be greatly appreciate for research communities dedicated to the detection, identification or source separation tasks.
 

### 2018-03-13 paper/articleNmfTrafficSimScene2018.tex
minor updates

@@
 Several variants of NMF are tested.
-The threshold initialized NMF appears to be the most relibale approach, allowing road traffic noise level estimation with average errors of 1.3 dB over the tested corpus of sound scenes.
+The threshold initialized NMF appears to be the most reliable approach, allowing road traffic noise level estimation with average errors of less than 1.3 dB over the tested corpus of sound scenes.
 
@@
 An other approach, followed in this paper, is to consider the blind source separation paradigm which consists in the extraction of a specific signal inside a set of mixed signals, see Figure \ref{fig:source_separation}.
-From the different existing methods, Non-negative Matrix Factorization (NMF) \cite{lee_learning_1999}, appers to be a relevant method for monophonic sensor networks.
+From the different existing methods, Non-negative Matrix Factorization (NMF) \cite{lee_learning_1999}, appears to be a relevant method for monophonic sensor networks.
 Many applications can be found for musical \cite{smaragdis_non-negative_2003, benetos2006musical} and speech \cite{wilson_speech_2008, mysore2011non} contents.
 Dedicated to sound separation with environmental sound, Immani and Kasa\"i \cite{satoshi_innami_nmf-based_2012} used NMF in a two steps sound separation with the help of time variant gain features.
-A first study \cite{gloaguen2018Estimation} has been conducted, in which diverse NMF estimation rules are compared, namely the supervised, the semi-supervised, and the threshold initialized NMF, have been applied on a large set of simulated sound scenes.
-This corpus mixes 6 sound categories (\textit{alert, animals, climate, humans, mechanics, transportation}) with a traffic component calibrated to different sound levels to obtain variable traffic predominance.
+Dedicated to the traffic sound separation, a first study \cite{gloaguen2018Estimation} has been conducted, in which diverse NMF estimation rules are compared, namely the supervised, the semi-supervised, and the threshold initialized NMF, have been applied on a large set of simulated sound scenes.
+This corpus mixes 6 sound categories (\textit{alert, animals, climate, humans, mechanics, transportation}) with a traffic component calibrated to different sound levels, according to the other sound classes (in the rest of the document, these sound classes are resumed as the \textit{interfering} sound class), to obtain variable traffic predominance.
 The diversity of this corpus was made to assess the performances and the limits of each NMF formula. 
-However, if it reveals the interest of NMF for urban sound environments, the assessment of its performance on a corpus of realistic sound scenes must be carried out in order to implement it on a sensor network.
-Design realistic urban sound mixtures makes it possible to access to many acoustic proprieties as the onset and offset time and the sound level of each sound class and especially the traffic component.
-However, like all simulated process, the issue of its realism has to be resolved in order to validate the NMF performances.
+However, if this study reveals the interest of NMF for urban sound environments, the assessment of its performance on a corpus of realistic sound scenes must be carried out in order to implement it on a sensor network.
+Design urban sound mixtures makes it possible to access to many acoustic proprieties as the onset and offset time and the sound level of each sound class and especially the traffic component.
+The realistic aspect of such a corpus is essential to obtain sound scenes similar to recordings and then validate NMF performances.
+However, like all simulated process, the issue of its realism has to be resolved.
 
@@
 Non-negative Matrix Factorization (NMF) is a linear approximation method proposed by Paatero and Tapper \cite{paatero1994positive} and popularized by Lee and Seung \cite{lee_learning_1999}.
-It consists in approximating a non negative matrix $\mathbf{V}$ $\in \mathbf{R}^+_{F \times N}$ by the product of two non negative matrices: $\mathbf{W}$, called \textit{dictionary} (or basis), and $\mathbf{H}$, called the matrix \textit{activation} with the dimensions $F \times K$ and $K \times N$ respectively.
+It consists in approximating a non negative matrix $\mathbf{V}$ $\in \mathbf{R}^+_{F \times N}$ by the product of two non negative matrices: $\mathbf{W}$, called \textit{dictionary} (or basis), and $\mathbf{H}$, called the \textit{matrix activation} with dimensions $F \times K$ and $K \times N$ respectively,
 
@@
 Because of the non-negativity constraint, only additive combinations between the elements of $\mathbf{W}$ are considered.
-%The dictionary is then composed of elementary elements providing a part based representation.
 
@@
 To be completely successful, all the acoustical sources must be considered in it which is not possible in an urban environment.
-To overcome this issue, semi-supervised NMF (SEM-NMF) \cite{lee_semi-supervised_2010} is considered.
+To overcome this issue, semi-supervised NMF (SEM-NMF) \cite{lee_semi-supervised_2010} has been proposed.
 It consists in decomposing, $\mathbf{W}_{F \times (K+J)}$ into two distinctive matrices: $\mathbf{W} = \left[ \mathbf{W_s}~\mathbf{W_r} \right]$ where $\mathbf{W_s}_{F \times K}$ is a fixed part of $\mathbf{W}$ composed of audio spectra and $ \mathbf{W_r}_{F \times J}$, a mobile part which is updated, see eq.
 \ref{eq:W_r_SS}.
-Thus it is possible to include elements not present in $\mathbf{W_s}$.
+Thus it is possible to include in it elements not present in $\mathbf{W_s}$.
 The dimension of $\mathbf{W_r}$ is set up as $J << K$ in order to best considered the sound source present in $\mathbf{W_s}$.
@@
 \begin{align}
-\mathbf{W_r}^{(i+1)} &\leftarrow \mathbf{W_r}^{(i)}\otimes\left(\frac{\left[\left(\mathbf{W_r H_r}^{(i)} \right)^{(\beta-2)}\otimes\mathbf{V} \right]\mathbf{H_r}^T}{\left(\mathbf{W_r H_r}^{(i)} \right)^{(\beta-1)}\mathbf{H_r}^T}\right)^{\gamma(\beta)}.\label{eq:W_r_SS}\\
+\mathbf{W_r}^{(i+1)} &\leftarrow \mathbf{W_r}^{(i)}\otimes\left(\frac{\left[\left(\mathbf{W_r H_r}^{(i)} \right)^{(\beta-2)}\otimes\mathbf{V} \right]\mathbf{H_r}^T}{\left(\mathbf{W_r H_r}^{(i)} \right)^{(\beta-1)}\mathbf{H_r}^T}\right)^{\gamma(\beta)}, \label{eq:W_r_SS}\\
 \mathbf{H_r}^{(i+1)} &\leftarrow \mathbf{H_r}^{(i)}\otimes\left(\frac{\mathbf{W_r}^T \left[\left(\mathbf{W_r H_r}^{(i)} \right)^{(\beta-2)}\otimes\mathbf{V} \right]}{\mathbf{W_r}^T \left(\mathbf{W_r H_r}^{(i)} \right)^{(\beta-1)}}\right)^{\gamma(\beta)},\label{eq:H_r_SS}\\
-\mathbf{H_s}^{(i+1)} &\leftarrow \mathbf{H_s}^{(i)}\otimes\left(\frac{\mathbf{W_s}^T \left[\left(\mathbf{W_s H_s}^{(i)} \right)^{(\beta-2)}\otimes\mathbf{V} \right]}{\mathbf{W_s}^T \left(\mathbf{W_s H_s}^{(i)} \right)^{(\beta-1)}}\right)^{\gamma(\beta)},\label{eq:H_s_SS}
+\mathbf{H_s}^{(i+1)} &\leftarrow \mathbf{H_s}^{(i)}\otimes\left(\frac{\mathbf{W_s}^T \left[\left(\mathbf{W_s H_s}^{(i)} \right)^{(\beta-2)}\otimes\mathbf{V} \right]}{\mathbf{W_s}^T \left(\mathbf{W_s H_s}^{(i)} \right)^{(\beta-1)}}\right)^{\gamma(\beta)}.\label{eq:H_s_SS}
 \end{align}
@@
 
-In this study, $\mathbf{W_s}$ is composed of traffic audio spectra, as it is the sound source of interest, to include in $\mathbf{W_r}$ all the other sources that can be present in the urban sound scenes.
+In this study, $\mathbf{W_s}$ is composed of traffic audio spectra, as it is the sound source of interest, to include in $\mathbf{W_r}$ the interfering class that can be present in the urban sound scenes.
 The traffic signal estimation is next defined by the fixed part,
@@
 
-where $\mathbf{w}$ is a $k$ element of $\mathbf{W}$ of $F \times 1$ dimension.
+where $\mathbf{w}$ is a $k$ element of $\mathbf{W}$ of $F \times 1$ dimensions.
 When $D_{\theta}\left(\mathbf{w_0} \Vert \mathbf{w'} \right)$=1, both element are identicals.
-If $D_{\theta}\left(\mathbf{w_0} \Vert \mathbf{w'} \right)$=0, the element $\mathbf{w'}$ is fully different.
+If $D_{\theta}\left(\mathbf{w_0} \Vert \mathbf{w'} \right)$=0, the element $\mathbf{w'}$ is fully different from $\mathbf{w_0}$.
 The extraction of traffic elements in $\mathbf{W'}$ is carried out by a hard thresholding method \cite{donoho1994threshold}.
-It consists in weighting in a binary way the traffic elements $\mathbf{W'}$ according to $D_{\theta}\left(\mathbf{w_0} \Vert \mathbf{w'} \right)$ and a threshold value such as:
+It consists in weighting in a binary way $\mathbf{W'}$ according to $D_{\theta}\left(\mathbf{w_0} \Vert \mathbf{w'} \right)$ and a threshold value such as:
 
-\ml{on pourrait tout a fait imaginer l'ajout d'élément de dictionaire random pour se rapprocher de la semi.}
+%\ml{on pourrait tout a fait imaginer l'ajout d'élément de dictionaire random pour se rapprocher de la semi.}
 
@@
 
-To sum-up, to approximate the audio spectrogram, $\mathbf{V}$ and estimate the traffic component, 3 methods are used.
-SUP-NMF is only based on the \textit{a priori} knowledge of the traffic to built the dictionary and is constrained to used it to obtain $\mathbf{\tilde{V}}_{traffic}$.
+To summarize, to approximate the audio spectrogram, $\mathbf{V}$ and estimate the traffic component, 3 methods are used which deal differently the \textit{a priori} knowledge on the traffic.
+SUP-NMF is only based on fixed \textit{traffic} elements and is constrained to used it to obtain $\mathbf{\tilde{V}}_{traffic}$.
 SEM-NMF completes this dictionary by the add of a mobile dictionary to consider the interfering sound events and then offer more flexibility.
-Finally, TI-NMF, as SUP-NMF, considers first a dictionary composed of \textit{traffic} spectra to then update them.
-Therefore, it is the entire dictionary that can be fully generalized to the sound scenes.
-The initiated dictionary allows the focused on the source of interest during the updates.
+Finally, TI-NMF considers first a dictionary composed of \textit{traffic} spectra, as SUP-NMF, but to then update them.
+This method therefore makes it the most suitable for solving the generalization issues as it is the entire dictionary that can be fully adapted to the sound scenes.
+The interest of an initiated dictionary is then to focus on the source of interest during the updates.
 
@@
 Indeed, the precise annotation of the sound level of a given source of interest in a polyphonic mixture is not feasible.
-We thus propose to design such evaluation corpus based on two corpora.
+Two corpora are proposed to design such a corpus of  evaluation.
 The first, called the reference corpus, is a set of recordings of urban sound environments, from which a sort of transcription is made, \textit{i.e.} which sound source is present from this time to this time, at which sound level.
-The second, called the elemental corpus, is a set of monophonic recordings of sound sources that are occurs in the scenes of the reference.
-The elemental corpus is built-up with still existing sound and with specific recordings, see part \ref{part:simScene}.
-The evaluation corpus is then built by replicating recordings of the elemental corpus according to the transcript of the reference corpus, see Figure \ref{fig:transcription}.
-
+The second, called the elemental corpus, is a set of monophonic recordings of isolated sound sources that are occurs in the scenes of the reference.
+The elemental corpus is built with still existing sound and with specific recordings, see part \ref{part:simScene}.
+The evaluation corpus is then designed by replicating the reference corpus according to the transcription with the help of the elemental corpus, see Figure \ref{fig:transcription}.
 More precisely, the urban sound scenes are taken from 76 recordings from 2 to 5 min, achieved in the \nth{13} district of Paris (France) at 19 different locations \footnote{Recordings were made as part of the Grafic project funded by Ademe}, which cover four various sound environments (Figure \ref{fig:map_grafic}), see Figure \ref{fig:map_grafic}.
@@
 Two of the 76 recordings are rejected for the analysis because the audio files were corrupted, resulting in 74 valid audio files assumed as representative of the variety of sound environments.
-The recordings are listened and categorized within four different sound environments, as proposed in \cite{can_describing_2015}: \textit{park} (P, 8 audio files with a cumulative duration of 16min01), \textit{quiet street} (Q, 33 audio files with a cumulative duration of 77min27), \textit{noisy street} (N, 24 audio files with a cumulative duration of 56min10) and \textit{very noisy street} (vN, 9 audio files with a cumulative duration of 21min42).
-Then, each audio file is annotated, noticing the start, end time and level of each sound event along with its sound class.
+The recordings are listened and categorized within four different sound environments, as proposed in \cite{can_describing_2015}: \textit{park} (P, 8 audio files with a cumulative duration of 16min01), \textit{quiet street} (Q, 35 audio files with a cumulative duration of 77min27), \textit{noisy street} (N, 23 audio files with a cumulative duration of 56min10) and \textit{very noisy street} (vN, 8 audio files with a cumulative duration of 21min42).
+Then, each audio file is transcripted, noticing the start, end time and level of each sound event along with its sound class.
 This annotation phase makes it possible to obtain simulated sound scenes with the same positions of sound events than the recordings and therefore as close as possible to the realistic scenes.
@@
 The sound database is composed of two categories of sound: the \textit{event} category, which includes 245 brief sound samples considered as salient, with a 1 to 20 seconds duration and classified among 21 sound classes (\textit{ringing bell, whistling bird, car horn, passing car, hammer, barking dog, siren, footstep, metallic noise, voice \dots}) and the \textit{background} (or \textit{texture}) category gathering 154 long duration sounds ($\approx$ 1m30), whose acoustic properties do not vary in time.
-This category includes among others\textit{whistling bird, crowd noise, rain, children playing in schoolyard, constant traffic noise} \dots sound classes.
+This category includes among others \textit{whistling bird, crowd noise, rain, children playing in schoolyard, constant traffic noise} \dots~sound classes.
 Each sound class is composed of multiple samples (\textit{carHorn01.wav, carHorn02.wav \dots}) that are randomly chosen by the \textit{simScene} software to bring diversity.
@@
 Overall, 103 car passages have been recorded.
-\ml{ajouter une description sur les problématiques d'overfitting et de generalisation.
-Cela doit etre un rappel d'un texte ecrit dans la partie nmf.
-A ce titre, cela pourrait etre une extension de l'etude, on en reparle}.
 In order to avoid overfitting issues, the audio samples of the first two cars (Renault Scenic and Renault M\'egane) are included in the \textit{SimScene}'s sound database (50 audio files in total).
@@
 A full description of the recordings can be found in \cite{gloaguen_creation_2017}.
+
+% \ml{ajouter une description sur les problématiques d'overfitting et de generalisation.
+Cela doit etre un rappel d'un texte ecrit dans la partie nmf.
+A ce titre, cela pourrait etre une extension de l'etude, on en reparle}.
 
@@
 
-With this built-up database, the \textit{SimScene} software and the annotations of the recordings, 74 simulated sound scenes are generated, which have the temporal structure of the reference recordings.
-The \textit{ebr} parameter is adjusted manually on each sound scene to be faithful compared to the recorded scenes.
-To validate this adjustment, the mean \textit{Traffic Interfering Ratio} is calculated and summarized on Table \ref{tab:mTIR}.
+With this built-up database, the \textit{SimScene} software and the transcriptions of the recordings, 74 simulated sound scenes are generated, which have the temporal structure of the reference recordings.
+The sound level of each sound class is adjusted manually on each sound scene to be faithful compared to the recorded scenes.
+To validate this adjustment, the mean \textit{Traffic Interfering Ratio} ($mTIR$) is calculated, see Table \ref{tab:mTIR}.
 It expresses, on all the scenes of an sound environment, the mean difference between the equivalent traffic sound levels of each scene, $L_{p,traffic}$, with the sound level of the \textit{interfering} sound class, $L_{p, interfering}$, which gathered all the other sound sources not related to the traffic, see Eq \ref{eq:mTIR}.
@@
 As it is the sound environment where the traffic is less present, the $mTIR$ is always negative for the \textit{park}.
-The \textit{interfering} sound class is therefore the main sound source.
+The \textit{interfering} sound class is therefore the main sound source which is coherent with this kind of environment.
 In the case of the \textit{quiet street}, the $TIR$ can be positive or negative depending on the traffic presence on the scene.
@@
 The first half includes 20 30-seconds audio files, including 5 scenes that belong to the sound environment \textit{park}, 6 from \textit{quiet street}, 4 from \textit{noisy street} and 5 from \textit{very noisy street} chosen randomly among the recorded scenes.
-The second half is composed of the transcribed scenes coresponding to the 20 reference scenes.
+The second half is composed of the transcribed scenes corresponding to the same 20 reference scenes.
 In order to limit the duration of the test to preserve the concentration of the subjects, each subject listens to only a subset of 20 sound scenes.
@@
 The experimental design is elaborated following a partially Balanced Incomplete Block Design (PBIBD) \cite{john1977optimal} that determines the listening order of each participant based on fixed parameters (number of listeners, total number of audio, number of audio assessed by each listeners).
-The listening order is then built in a such way than each listener listens the same amount of recorded scenes and transcribed scenes to avoide statistical biais.
+The listening order is then built in a such way than each listener listens almost the same amount of recorded scenes and transcribed scenes to avoid statistical biais.
 The experimental design and the listening order per participant are performed with the package \textit{sensoMineR} on the \textit{R} software \cite{le_sensominer:_2008}.\\
@@
 $62\%$ of the participants declared having no experience in the listening of urban sound mixtures.
-
-A Student's t-test is performed with the $H_0$ hypothesis which considers the similarity between the distribution of the average rates for the recorded and the transcribed scenes of each participant with threshold \textit{p-value} of 5$\%$.
+The average rate of all the transcribed scenes ($m_{transcribed} = 5.1 (\pm 1.6)$) is then similar to the recorded ones ($m_{recorded} = 4.9 (\pm 1.6)$).
+In complement, a Student's t-test is performed with the $H_0$ hypothesis which considers the similarity between the distribution of the average rates for the recorded and the transcribed scenes of each participant with a \textit{p-value} threshold of 5$\%$.
 Results are displayed in Table \ref{tab:ttest_type_of_scene}.
@@
 
-These results confirm that the recorded and the transcribed scenes are perceived in a similar way by the panel (\textit{p-value} $>$ $\alpha$).
+This result confirms that the recorded and the transcribed scenes are perceived in a similar way by the panel (\textit{p-value} $>$ $\alpha$).
 More details on the results can be found in  \cite{gloaguen_creation_2017}.
@@
 
+The experimental factors related to this estimator are summarized in Table \ref{tab:experimental_factorsFilter}.
+
 The second estimator is based on the three NMF formula presented in part \ref{part:nmf} (see Figure \ref{fig:bloc_diagram_nmf}).
@@
 
-The dictionary is designed from a second sound database specially dedicated to this task to prevent any overfitting issues.
-It contains  53 audio files of the 2 cars (Dacia Sandero and Renault Clio) recorded on the runway, see part \ref{part:simScene}.
+The dictionary is designed from a sound database specially dedicated to this task.
+To prevent any overfitting issues, it contains 53 audio files of the 2 other cars (Dacia Sandero and Renault Clio) recorded on the runway, see part \ref{part:simScene}.
 Those recordings have not been considered for the creation of the evaluation dataset.
@@
 A $\mathbf{K}$-means clustering algorithm is applied to reduce these dimensions to $\mathbf{K} \in \lbrace$25, 50, 100, 200$\rbrace$ in order to avoid redundant information and decrease the computation time.
-The resulting $\mathbf{K}$ centroids are taken as the elements of $\mathbf{W}$.
+The resulting $\mathbf{K}$ clusters are taken as the elements of $\mathbf{W}$.
 The obtained dictionary is expressed in third octave bands to decrease the dimensionality.
@@
 For the filter estimator, 24 settings are computed (4 $\times$ 6).
-For SUP and SEM-NMF, 240 settings are computed (4 $\times$ 2 $\times$ (2 $\times$ 4 + 1 $\times$ 2) $\times$ 3).
-Finally, for TI-NMF, the number of settings is much higher (3720) due to the higher cardinality of the set of threshold values (4 $\times$ 1 $\times$ (2 $\times$ 4 + 1 $\times$ 2) $\times$ 3 $\times$ 31).
+For SUP and SEM-NMF, 192 settings are computed (4 $\times$ 2 $\times$ 2 $\times$ 4 $\times$ 3).
+Finally, for TI-NMF, the number of settings is much higher (2976) due to the higher cardinality of the set of threshold values (4 $\times$ 1 $\times$ 2 $\times$ 4  $\times$ 3 $\times$ 31).
 The experimental factors and their different modalities are displayed on Tables \ref{tab:experimental_factorsFilter} and \ref{tab:experimental_factorsNMF}.
@@
 
-\begin{tabularx}{17.5cm}{L{3cm}@{}@{}C{3cm}@{}@{}C{3cm}@{}@{}C{3cm}@{}@{}C{3cm}@{}C{2cm}@{}}
-    \textbf{sound environment} & park & quiet street & noisy street & very noisy street & 4 \\
+\begin{tabularx}{17.5cm}{L{3cm}@{}C{3cm}@{}@{}C{3cm}@{}@{}C{3cm}@{}@{}C{3cm}@{}C{2cm}@{}}
+    \textbf{sound environment} & \begin{tabular}[c]{@{}c@{}}park\\ 'P'\end{tabular} & \begin{tabular}[c]{@{}c@{}}quiet street\\ 'Q'\end{tabular} & \begin{tabular}[c]{@{}c@{}}noisy street\\ 'N' \end{tabular}& \begin{tabular}[c]{@{}c@{}}very noisy street\\ 'vN'\end{tabular} & 4\\
 \end{tabularx}
@@
 \begin{tabularx}{17.5cm}{L{3cm}@{}C{3cm}@{}@{}C{3cm}@{}@{}C{3cm}@{}@{}C{3cm}@{}C{2cm}@{}}
-    \textbf{sound environment} & park & quiet street & noisy street & very noisy street & 4\\
+    \textbf{sound environment} & \begin{tabular}[c]{@{}c@{}}park\\ 'P'\end{tabular} & \begin{tabular}[c]{@{}c@{}}quiet street\\ 'Q'\end{tabular} & \begin{tabular}[c]{@{}c@{}}noisy street\\ 'N' \end{tabular}& \begin{tabular}[c]{@{}c@{}}very noisy street\\ 'vN'\end{tabular} & 4\\
 \end{tabularx}
@@
 
-\begin{tabularx}{17.5cm}{L{3cm}@{}C{4cm}@{}@{}C{4cm}@{}@{}C{4cm}@{}C{2cm}@{}}
-    $\mathbf{w_t}$ (s)& 0.5 & 1 & \textit{all} & 3 \\
+\begin{tabularx}{17.5cm}{L{3cm}@{}C{6cm}@{}@{}C{6cm}@{}C{2cm}@{}}
+    $\mathbf{w_t}$ (s)& 0.5 & 1 & 2 \\
 \end{tabularx}
@@
 This combination of settings offers the most fitted method to be adapted to all the sound environments.
-Furthermore, on the dictionary creation, except SEM-NMF where all the best methods according to $\beta$ use the same dictionary, no specific dictionary form through all the method, is used.
-SUP and SEM-NMF privilege a high number of element ($\mathbf{K}$ = 200) with a fine description of the audio samples ($\mathbf{w_t} \in \lbrace 0.5$,$1 \rbrace$ second).
-For TI-NMF, the composition of the initial dictionary is more diverse both on the $\mathbf{K}$ value and on the finesse  of the description ($\mathbf{w_t} \in \lbrace all$,$0.5 \rbrace$ second).
+Furthermore, on the dictionary creation, only SEM-NMF proposes the same dictionary design for all the best methods according to $\beta$.
+In the opposite, SUP and TI-NMF propose different associations between $\mathbf{K}$ and $\mathbf{w_t}$.
+One can notice that, through the 3 methods, it is mainly a high number of elements in $\mathbf{K}$ (100, 200) that is preferred.
+With more elements, it makes it possible to resolve easily the generalization issue.
 
@@
 As this approach considers the remaining energy as the traffic component, no distinction is made between the different sound sources not related to the traffic component.
+The interfering sound class is then wrongly considered as traffic component.
 In the opposite, for noisy and very noisy environments, the performances of the LP filter are good ($MAE$ $<$ 1 dB).
-The errors are then due to a high deletion of the traffic energy by the filter while it becomes the main sound source.
+The errors are here due to a high deletion of the traffic energy by the filter while it becomes the main sound source.
+Consequently, the 500 Hz LP filter estimator provides a low $MAE$ error through the sound environments thank to a balance between the remaining and discarded energy.
 
@@
 With this method, as NMF minimizes the cost function, eq.
-\ref{eq:min-D-WH}, the dictionary's elements are used to model the other sound sources.
+\ref{eq:min-D-WH}, the dictionary's elements are used to model the other sound sources which can not allow a rightful approximation of the traffic component.
 In the opposite, for \textit{noisy} and \textit{very noisy} environments, SUP-NMF identifies correctly the traffic components ($MAE$ $<$ 0.6 dB) as it is the main source.
@@
 Finally, TI-NMF presents the most performing results, except in the \textit{park} environment ($MAE$ = 2.95 dB).
-In this sound environment, with a threshold value $t_h$ = 0.32, the traffic dictionary is composed, on average, of 136 elements.
+In this sound environment, with the threshold value $t_h$ = 0.32, the traffic dictionary is composed, on average, of 136 elements.
+By increasing this value, it would be possible to put aside the spectra furthest away of the traffic elements and then decrease the error.
 For the rest of the sound environments, TI-NMF has the lowest errors.
@@
 
-Considering a unique dictionary fitted to the sound scene under evaluation thus makes TI-NMF very effective when traffic is predominant while, by the use of the thresholding, it discards the elements of the dictionary that deviate too much when the traffic is less present.
+Considering a unique dictionary fitted to the sound scene under evaluation thus makes TI-NMF very effective when traffic is predominant, while the thresholding step make it possible to discards the elements of the dictionary that deviate too much when the traffic is less present.
 
@@
 
-For reproducibility purposes, the evaluation corpus, the experimental protocol and the programs developed under the Matlab software are available online \footnote{\url{https://github.com/jean-remyGloaguen/articleApplied2018}}.
+For reproducibility purposes, the evaluation corpus, the experimental protocol and the programs developed under the Matlab software are available online \footnote{\url{https://github.com/jean-remyGloaguen/articleNmfTrafficSimScene2018}} and as the sound dataset \footnote{\url{https://zenodo.org/record/1184443#.WqfVb3wiGpo}}.
 

### 2018-03-13 experiment/functions/nmf/algo_nmfSemiSupervisedEXP.m

@@ -2,17 +2,8 @@ function NMF = algo_nmfSemiSupervisedEXP(H,W,Y,Z,V,iteration,soundMix,setting)
 
 beta = setting.beta;
 sparsity = setting.sparsity;
-smoothness = setting.SM_weight;
 J = setting.SS_sizeWrand;
 K = size(W,2);
-
-if setting.SS_weight ~= 0
-    mu = setting.SS_weight;
-    betaM = setting.SS_betaM;
-    lambdaSS = setting.SS_lambda;
-else
-    mu = 0;
-end
 
 if ~isequal(beta,0) || ~isequal(beta,1) || ~isequal(beta,2)
     if beta < 1
@@ -24,21 +15,16 @@ if ~isequal(beta,0) || ~isequal(beta,1) || ~isequal(beta,2)
     end
 end
 
-if smoothness~=0
-    smoothForm = setting.smoothnessForm;
-    
-    switch smoothForm
+if setting.SM_weight~=0
+    switch setting.smoothnessForm
         case 'all'
-            lambdaSM = [sum(W,1) sum(Y,1)];
-            smoothnessTot(1:K+J,1) = smoothness;
+            smoothness(1:size(W,2),1) = setting.SM_weight;
         case 'traffic'
-            lambdaSM = sum(W,1);
-            smoothnessTot = zeros(K,1);
-            smoothnessTot(soundMix.indTraffic==1) = smoothness;
+            smoothness = zeros(size(W,2),1);
+            smoothness(soundMix.indTraffic==1) = setting.SM_weight;
     end
 else
-    smoothnessTot = 0;
-    lambdaSM = NaN;
+    smoothness = 0;
 end
 
 Vap = W*H;
@@ -49,24 +35,18 @@ NMF = [];
 switch beta
     case 2
         for iter = 1:iteration
-            if mu == 0
-                Y = Y.*(((V*Z')./(Vtot*Z')).^gamma);
-            else
-                div = betadivEXP(Y(:)*ones(1,size(W,2)),repmat(W,size(Y,2),1),betaM);
-                Cm = exp((-1/lambdaSS)*sum(div));
-
-                num = lambdaSS*V*Z'+mu*Y.^(betaM-1)*Cm;
-                den = (lambdaSS*Vtot)*Z'+mu*Y.^(betaM-2).*Cm.*sum(W,2);
-                Y = Y.*(num./den);
[…]

### 2018-03-13 experiment/functions/nmf/algo_nmfSupervisedEXP.m

@@ -1,7 +1,6 @@
 function NMF = algo_nmfSupervisedEXP(H,W,V,iteration,soundMix,setting)
 
 sparsity = setting.sparsity; 
-smoothness = setting.SM_weight; 
 beta = setting.beta;
 Vap = W*H;
 
@@ -17,20 +16,16 @@ if ~isequal(beta,0) || ~isequal(beta,1) || ~isequal(beta,2)
     end
 end
 
-if smoothness~=0
-    lambda = sum(W,1);
-    smoothForm = setting.smoothnessForm;
-    
-    switch smoothForm
+if setting.SM_weight~=0
+    switch setting.smoothnessForm
         case 'all'
-            smoothnessTot(1:size(W,2),1) = smoothness;
+            smoothness(1:size(W,2),1) = setting.SM_weight;
         case 'traffic'
-            smoothnessTot = zeros(size(W,2),1);
-            smoothnessTot(soundMix.indTraffic==1) = smoothness;
+            smoothness = zeros(size(W,2),1);
+            smoothness(soundMix.indTraffic==1) = setting.SM_weight;
     end
 else
-    smoothnessTot = 0;
-    lambda = NaN;
+    smoothness = 0;
 end
 
 switch beta
@@ -40,7 +35,10 @@ switch beta
                 H = H.*(W'*V-sparsity)./(W'*Vap);
                 H(H<0)=0;
             else
-                H = updateSmoothEssidEXP(H,W,V,Vap,beta,smoothnessTot,sparsity,lambda);
+                H_1 = [zeros(K,1) H(:,1:end-1)];
+                H_2 = [H(:,2:end) zeros(K,1)];
+                H_12 = [zeros(K,1) H(:,2:end-1) zeros(K,1)];
+                H = H.*(W'*V-sparsity+2*smoothness.*(H_1+H_2))./(W'*Vap+2*smoothness.*(H+H_12));
             end
             H(isnan(H)) = 0;
             Vap = W*H;
@@ -51,7 +49,10 @@ switch beta
             if smoothness==0
                 H = H.*((W'*(V.*Vap.^(-1)))./(sparsity+W'*Vap.^(0)));
             else
-                H = updateSmoothEssidEXP(H,W,V,Vap,beta,smoothnessTot,sparsity,lambda);
+                H_1 = [zeros(K,1) H(:,1:end-1)];
[…]

### 2018-03-13 experiment/nmfthgr2estimation.m

@@ -55,7 +55,7 @@ else
     sequentialData = config.sequentialData;
 end
 
-for ii = 1:numberScene
+parfor ii = 1:numberScene
     
     fileTraffic = audioread(strcat(creationSceneDir,globalName{ii},'_traffic.wav'));
     fileRest = audioread(strcat(creationSceneDir,globalName{ii},'_interfering.wav'));

### 2018-04-06 paper/articleNmfTrafficSimScene2018.tex
jfp corrections

@@
 
-In this paper, a Non-negative Matrix Factorization (NMF) framework is developed to estimate road traffic noise levels within urban sound scenes.
-To evaluate the performance of the proposed approach, a synthetic corpus of sound scenes is designed to cover most common soundscape settings.
-The realism of the produced scenes is validated through a perceptual test, forming a protocol that both reproduces the sensor network outputs, and in which the actual occurrence and sound level of each source is known.
+In this paper, the Non-negative Matrix Factorization (NMF) framework is developed to estimate road traffic noise levels within urban sound scenes.
+To evaluate the performances of the proposed approach, a synthetic corpus of sound scenes is designed, to cover most common soundscape settings, and whom realism is validated through a perceptual test.
+The simulated scenes reproduce then the sensor network outputs, in which the actual occurrence and sound level of each source are known.
 
 Several variants of NMF are tested.
-The threshold initialized NMF appears to be the most reliable approach, allowing road traffic noise level estimation with average errors of less than 1.3 dB over the tested corpus of sound scenes.
+The proposed approach, named threshold initialized NMF, appears to be the most reliable approach, allowing road traffic noise level estimation with average errors of less than 1.3 dB over the tested corpus of sound scenes.
 
@@
 \begin{keyword}
-non-negative matrix factorization \sep urban sound environment \sep road traffic sound level estimation
+Non-negative Matrix Factorization \sep urban sound environment \sep road traffic sound level estimation
 \end{keyword}
@@
 Road traffic concentrates particular attention as it is the main urban source of noise annoyance.
-Road traffic noise maps are typically built from data collection that consist of traffic data collected on the main roads (flow rates, mean speeds and heavy vehicle ratio) and urban geographic data (building heights and location, topology, ground surfaces, etc.).
+Road traffic noise maps are typically built from data collection that consist of traffic data collected on the main roads (flow rates, mean speeds and heavy vehicle ratio) and urban geographic data (building heights and location, topology, ground surfaces \dots).
 Follows sound emission and sound propagation computational techniques, resulting in the production of the two indicators equivalent A-weighted sound levels, $L_{DEN}$ (\textit{Day-Evening-Night}) and $L_N$ (\textit{Night}) \cite{kephalopoulos2012common}.
 This procedure also enables drawing up action plans to reduce the noise exposure.
-Despite their unanimously recognized interest, noise maps suffer some limitations.
+Despite their unanimously recognized interest, noise maps suffer from some limitations.
 The computing cost required to produce noise maps at the city scale calls simplifications of the numerical tools and the simulation models that both generate uncertainties \cite{van_leeuwen_noise_2015}.
@@
 
-All these measurement protocols allows the combination of measures and predictions to improve the accuracy of the produced noise maps.
+All these measurement protocols allow the combination of measures and predictions to improve the accuracy of the produced noise maps.
 Traffic noise maps and measurements were compared on restrictive areas in \cite{lefebvre2017traffic} and \cite{mioduszewski2011noise}.
@@
 Many works have dealt with the detection \cite{luitel2016sound} or the recognition \cite{defreville_automatic_2006} of sound events in environmental sound scenes.
-In these cases, a two-step scheme is followed where audio samples are described with a set of features (Mel Frequency Cepstral Coefficient, MPEG-7 descriptors \dots) and classified them with the help of a classifiers (Gaussian Mixtures Models, Artificial Neural Network \dots) \cite{chu2008environmental, cowling_comparison_2003}.
-The classifiers is learnt from a learning database and are next applied on a test database to validate the algorithms.
+In these cases, a two-step scheme is followed where audio samples are described with a set of features (Mel Frequency Cepstral Coefficient, MPEG-7 descriptors \dots) and classified with the help of a classifier (Gaussian Mixtures Models, Artificial Neural Network \dots) \cite{chu2008environmental, cowling_comparison_2003}.
+The classifier is learnt from a learning database and is next applied on a test database to validate the algorithms.
 Dedicated to the traffic, in \cite{socoro_anomalous_2017}, an Anomalous Event Detection, based on MFCC features, is proposed with the specific aim to improve the traffic sound estimation.
@@
 Many applications can be found for musical \cite{smaragdis_non-negative_2003, benetos2006musical} and speech \cite{wilson_speech_2008, mysore2011non} contents.
-Dedicated to sound separation with environmental sound, Immani and Kasa\"i \cite{satoshi_innami_nmf-based_2012} used NMF in a two steps sound separation with the help of time variant gain features.
+Dedicated to sound separation with environmental sounds, Immani and Kasa\"i \cite{satoshi_innami_nmf-based_2012} used NMF in a two steps sound separation with the help of time variant gain features.
 Dedicated to the traffic sound separation, a first study \cite{gloaguen2018Estimation} has been conducted, in which diverse NMF estimation rules are compared, namely the supervised, the semi-supervised, and the threshold initialized NMF, have been applied on a large set of simulated sound scenes.
-This corpus mixes 6 sound categories (\textit{alert, animals, climate, humans, mechanics, transportation}) with a traffic component calibrated to different sound levels, according to the other sound classes (in the rest of the document, these sound classes are resumed as the \textit{interfering} sound class), to obtain variable traffic predominance.
+This corpus mixes 6 sound categories (\textit{alert, animals, climate, humans, mechanics, transportation}) with a traffic component calibrated to different sound levels, according to the other sound classes (in the rest of the document, these sound classes, not related to the traffic component, are resumed as the \textit{interfering} sound class), to obtain variable traffic predominance.
 The diversity of this corpus was made to assess the performances and the limits of each NMF formula. 
 However, if this study reveals the interest of NMF for urban sound environments, the assessment of its performance on a corpus of realistic sound scenes must be carried out in order to implement it on a sensor network.
-Design urban sound mixtures makes it possible to access to many acoustic proprieties as the onset and offset time and the sound level of each sound class and especially the traffic component.
-The realistic aspect of such a corpus is essential to obtain sound scenes similar to recordings and then validate NMF performances.
-However, like all simulated process, the issue of its realism has to be resolved.
+Design urban sound mixtures makes it possible to access to many acoustic properties as the onset and offset time and the sound level of each sound class and especially the traffic component.
+The realistic aspect of such a corpus is essential to obtain sound scenes similar to recordings and to validate NMF performances.
+However, like all simulated process, the realism of the scenes must be perceptually verified.
 
-In this paper, a realistic urban sound corpus based on annotated urban recordings, and whose degree of realism is validated through a perceptual test, is designed in order to estimate the traffic sound level with the help of the NMF framework.
+In this paper, an urban sound corpus based on annotated urban recordings, and whose degree of realism is assessed through a perceptual test, is designed in order to estimate the traffic sound level with the help of the NMF framework.
 The different NMF approaches are described in section \ref{part:nmf}.
@@
 This condition however is not mandatory.
-When  applying NMF to audio data, $\mathbf{V}$ is usually considered as the magnitude spectrogram obtained by a Short-Time Frequency Transform, $\mathbf{W}$ includes audio spectra and $\mathbf{H}$ is equivalent to the temporal activation of each spectrum, see Figure \ref{fig:exampleNMF}.
+When  applying NMF to audio data, $\mathbf{V}$ is usually considered as the magnitude spectrogram obtained by a Short-Time Fourier Transform, $\mathbf{W}$ includes audio spectra and $\mathbf{H}$ is equivalent to the temporal activation of each spectrum, see Figure \ref{fig:exampleNMF}.
 Because of the non-negativity constraint, only additive combinations between the elements of $\mathbf{W}$ are considered.
@@
 \includegraphics[width=.9\linewidth]{./figures/schema_introduction_nmf.pdf}
-\caption{NMF decomposition of an audio spectrogram composed of 3 elements ($K$ = 3): passing car (a), car horn (b) and whistling bird (c).}
+\caption{NMF decomposition of an audio spectrogram $V$ composed of 3 elements ($K$ = 3): passing car (a), car horn (b) and whistling bird (c).}
 \label{fig:exampleNMF}
@@
 $\mathbf{H}$ is then the only matrix to estimate and is updated at every iteration (eq.
-\ref{eq:updateH}) \cite{fevotte_algorithms_2011}.
+\ref{eq:updateH}) \cite{fevotte_algorithms_2011}:
 
@@
 with $\gamma(\beta) = \frac{1}{2-\beta},$ for $\beta < 1$, $ \gamma(\beta) = 1$, for $\beta \in \left[1,2\right]$ and $\gamma(\beta) = \frac{1}{\beta-1}$ for $\beta > 2$.
-The product $A\otimes B$ and $A/B$ symbolized the Hadamard product and ratio.
+The $A\otimes B$ and $A/B$ operators represent the Hadamard product and ratio.
 
@@
 The main issue with the supervised approach is the representational limit imposed by a fixed $\mathbf{W}$.
-To be completely successful, all the acoustical sources must be considered in it which is not possible in an urban environment.
+To be completely successful, all the acoustical sources must be considered in the basis $\mathbf{W}$ which is not always possible in a complex urban environment.
 To overcome this issue, semi-supervised NMF (SEM-NMF) \cite{lee_semi-supervised_2010} has been proposed.
-It consists in decomposing, $\mathbf{W}_{F \times (K+J)}$ into two distinctive matrices: $\mathbf{W} = \left[ \mathbf{W_s}~\mathbf{W_r} \right]$ where $\mathbf{W_s}_{F \times K}$ is a fixed part of $\mathbf{W}$ composed of audio spectra and $ \mathbf{W_r}_{F \times J}$, a mobile part which is updated, see eq.
+It consists in decomposing, $\mathbf{W}_{F \times (K+J)}$ into two distinctive matrices: $\mathbf{W} = \left[ \mathbf{W_s}~\mathbf{W_r} \right]$ where $\mathbf{W_s}_{F \times K}$ is a fixed part of $\mathbf{W}$ composed of known audio spectra and $ \mathbf{W_r}_{F \times J}$, a mobile part which is updated, see eq.
 \ref{eq:W_r_SS}.
-Thus it is possible to include in it elements not present in $\mathbf{W_s}$.
-The dimension of $\mathbf{W_r}$ is set up as $J << K$ in order to best considered the sound source present in $\mathbf{W_s}$.
+Thus it is possible to let the method define the best elements to include in $\mathbf{W_r}$.
+Its dimension is set up as $J << K$ in order to consider, as a priority, the sound sources present in $\mathbf{W_s}$.
 $\mathbf{H}$ is then also decomposed in two matrices, $\mathbf{H}_{(K+J) \times N} = \genfrac[]{0pt}{0}{\mathbf{H_s}}{\mathbf{H_r}}$.
-The eq.
+Eq.
 \ref{eq:nmf} becomes
@@
 
-$\mathbf{H_r}$ and $\mathbf{H_s}$ are updated separately, see eq. \ref{eq:H_r_SS} \ref{eq:H_s_SS}.
+$\mathbf{H_r}$ and $\mathbf{H_s}$ are updated separately, see eq. \ref{eq:H_r_SS} to eq. \ref{eq:H_s_SS}.
 
@@
 
-In this study, $\mathbf{W_s}$ is composed of traffic audio spectra, as it is the sound source of interest, to include in $\mathbf{W_r}$ the interfering class that can be present in the urban sound scenes.
+In this study, $\mathbf{W_s}$ is composed of traffic audio spectra, as it is the sound source of interest.
+Sources included in $\mathbf{W_r}$ are other sound sources (corresponding to the interfering class) that can be present in the urban sound scenes.
 The traffic signal estimation is next defined by the fixed part,
@@
 
-The addition of the mobile part $\mathbf{W_r}$ gives more flexibility.
+The addition of $\mathbf{W_r}$ gives more flexibility to the method to represent correctly the spectrogram $\mathbf{V}$.
 The representational capability is increased, thus the approach is more adaptive to the different urban sound environments.
-Applications of SEM-NMF can be found for musical \cite{weninger2012supervised, kitamura_music_2014} and speech content \cite{joder2012real, mysore2011non}.
+Applications of SEM-NMF can be found for musical \cite{weninger2012supervised, kitamura_music_2014} and speech contents \cite{joder2012real, mysore2011non}.
 
@@
 With this operation, $\mathbf{W_0}$ is oriented to the focused sound source (the road traffic) but also can be adapted to the content of the scene thanks to the updates.
-After $N$ iterations, each element $k$ of the final dictionary, $\mathbf{W'}$, is compared with its initial value in $\mathbf{W_0}$, in order to identify which element is stayed closed to the traffic component.
-A cosine similarity $D_{\theta}\left(\mathbf{W_0} \Vert \mathbf{W'} \right)$ is computed for each element $k$ as it is an invariant scale and a bounded method,
+After $N$ iterations, each element $k$ of the final dictionary, $\mathbf{W'}$, is compared with its initial value in $\mathbf{W_0}$, in order to identify which element is stayed close to the traffic component.
+A cosine similarity $D_{\theta}\left(\mathbf{W_0} \Vert \mathbf{W'} \right)$ is computed for each element $k$ as it is scale-invariant and bounded,
 
@@
 where $\mathbf{w}$ is a $k$ element of $\mathbf{W}$ of $F \times 1$ dimensions.
-When $D_{\theta}\left(\mathbf{w_0} \Vert \mathbf{w'} \right)$=1, both element are identicals.
-If $D_{\theta}\left(\mathbf{w_0} \Vert \mathbf{w'} \right)$=0, the element $\mathbf{w'}$ is fully different from $\mathbf{w_0}$.
+When $D_{\theta}\left(\mathbf{w_0} \Vert \mathbf{w'} \right)$=1, the element $\mathbf{w'}$ is identical to $\mathbf{w_0}$.
+If $D_{\theta}\left(\mathbf{w_0} \Vert \mathbf{w'} \right)$=1, both elements are fully different.
 The extraction of traffic elements in $\mathbf{W'}$ is carried out by a hard thresholding method \cite{donoho1994threshold}.
-It consists in weighting in a binary way $\mathbf{W'}$ according to $D_{\theta}\left(\mathbf{w_0} \Vert \mathbf{w'} \right)$ and a threshold value such as:
+It consists in weighting in a binary way $\mathbf{W'}$ according to $D_{\theta}\left(\mathbf{w_0} \Vert \mathbf{w'} \right)$ and a threshold value $\alpha_k$ such as:
 
@@
 
-To summarize, to approximate the audio spectrogram, $\mathbf{V}$ and estimate the traffic component, 3 methods are used which deal differently the \textit{a priori} knowledge on the traffic.
+To summarize, to approximate the audio spectrogram $\mathbf{V}$ and estimate the traffic component, 3 methods are used which deal differently with the \textit{a priori} knowledge on the traffic.
 SUP-NMF is only based on fixed \textit{traffic} elements and is constrained to used it to obtain $\mathbf{\tilde{V}}_{traffic}$.
-SEM-NMF completes this dictionary by the add of a mobile dictionary to consider the interfering sound events and then offer more flexibility.
-Finally, TI-NMF considers first a dictionary composed of \textit{traffic} spectra, as SUP-NMF, but to then update them.
-This method therefore makes it the most suitable for solving the generalization issues as it is the entire dictionary that can be fully adapted to the sound scenes.
+SEM-NMF completes this dictionary by the add of a mobile dictionary to better consider the interfering sound events and then offer more flexibility.
+Finally, TI-NMF considers first a dictionary composed of \textit{traffic} spectra, as SUP-NMF, but allows an update of them.
 The interest of an initiated dictionary is then to focus on the source of interest during the updates.
+This method therefore makes it more suitable for solving the generalization issues as it is the entire dictionary that can be fully adapted to the sound scenes.
+The elements that deviates too much from the originals spectra are then discarded by the thresholding step.
 
-These above described methods are applied on an evaluation corpus,  composed of simulated sound scenes, in order to compare the estimated sound levels with a non ambiguous reference.
-This new and more realistic sound corpus is designed by considering realistic urban sound environments with many kind of sound sources mixed.
+These above described methods are applied on an evaluation corpus, composed of simulated sound scenes, in order to compare the estimated traffic sound levels with an exact reference.
+This new sound corpus is designed by considering realistic urban sound environments with many kinds of mixed sound sources.
 
@@
 \centering
-\includegraphics[width=0.7\linewidth]{figures/realistic_urban_sound_scene_design.pdf}
-\caption{Bloc diagram of the design of evaluation corpus}
+\includegraphics[width=0.8\linewidth]{figures/realistic_urban_sound_scene_design.pdf}
+\caption{Bloc diagram of the design of the evaluation corpus}
 \label{fig:transcription}
@@
 
-The evaluation corpus is both required to be realistic and non ambiguous in terms of traffic sound level.
-The former calls for the use of recordings of urban sound environments, and the latter imposes the use of controlled sound sources with known properties: onset and offset time location, and sound level.
-Indeed, the precise annotation of the sound level of a given source of interest in a polyphonic mixture is not feasible.
-Two corpora are proposed to design such a corpus of  evaluation.
-The first, called the reference corpus, is a set of recordings of urban sound environments, from which a sort of transcription is made, \textit{i.e.} which sound source is present from this time to this time, at which sound level.
-The second, called the elemental corpus, is a set of monophonic recordings of isolated sound sources that are occurs in the scenes of the reference.
-The elemental corpus is built with still existing sound and with specific recordings, see part \ref{part:simScene}.
+The evaluation corpus must be both realistic and non ambiguous in terms of traffic sound level.
+The former calls for the use of real urban sound environments, and the latter imposes the use of controlled sound sources with known properties: onset and offset time location, and sound level.
+Indeed, the precise annotation of the sound level of a given source of interest in a real recording is a very complex problem.
+Two corpora are proposed to design a valid corpus of  evaluation.
+The first, called the reference corpus, is a set of recordings of urban sound environments, from which, what we call, a transcription is made, \textit{i.e.} which sound source is present from this time to this time.
+The second, called the elemental corpus, is a set of monophonic recordings of isolated sound sources that represent categories of events that are present in the reference scenes.
+The elemental corpus is built with still existing sounds and with specific recordings, see part \ref{part:simScene}.
 The evaluation corpus is then designed by replicating the reference corpus according to the transcription with the help of the elemental corpus, see Figure \ref{fig:transcription}.
-More precisely, the urban sound scenes are taken from 76 recordings from 2 to 5 min, achieved in the \nth{13} district of Paris (France) at 19 different locations \footnote{Recordings were made as part of the Grafic project funded by Ademe}, which cover four various sound environments (Figure \ref{fig:map_grafic}), see Figure \ref{fig:map_grafic}.
+More precisely, the reference corpus is composed of 76 recordings from 2 to 5 min, achieved in the \nth{13} district of Paris (France) at 19 different locations \footnote{Recordings were made as part of the Grafic project funded by Ademe}, see Figure \ref{fig:map_grafic}, which cover various sound environments.
 A complete description of the experimental protocol can be found in \cite{aumond_modelling_2017}.
-Two of the 76 recordings are rejected for the analysis because the audio files were corrupted, resulting in 74 valid audio files assumed as representative of the variety of sound environments.
-The recordings are listened and categorized within four different sound environments, as proposed in \cite{can_describing_2015}: \textit{park} (P, 8 audio files with a cumulative duration of 16min01), \textit{quiet street} (Q, 35 audio files with a cumulative duration of 77min27), \textit{noisy street} (N, 23 audio files with a cumulative duration of 56min10) and \textit{very noisy street} (vN, 8 audio files with a cumulative duration of 21min42).
+Two of the 76 recordings are rejected for the analysis because the audio files were corrupted, resulting in 74 valid audio files assumed as representative of the variety of urban sound environments.
+The recordings are listened and categorized within four different sound environments, as proposed in \cite{can_describing_2015}: \textit{park} (8 audio files with a cumulative duration of 16min01), \textit{quiet street} (35 audio files with a cumulative duration of 77min27), \textit{noisy street} (23 audio files with a cumulative duration of 56min10) and \textit{very noisy street} (8 audio files with a cumulative duration of 21min42).
 Then, each audio file is transcripted, noticing the start, end time and level of each sound event along with its sound class.
-This annotation phase makes it possible to obtain simulated sound scenes with the same positions of sound events than the recordings and therefore as close as possible to the realistic scenes.
+This annotation phase makes it possible to produce simulated sound scenes with the same positions of sound events than the recordings and therefore as close as possible to the real scenes.
 
@@
 \subsection{Generation of the evaluation corpus}\label{part:simScene}
-The sound scenes are generated with the \textit{SimScene} software\footnote{Open-source project available at: \url{https://bitbucket.org/mlagrange/simscene}}, \cite{rossignol_simscene:_2015}, which is a simulation software generating monaural sound mixtures in wav format with a 44.1 kHz sampling rate from an isolated sound database.
-This software have already been used in a wide range of experiments for sound detection algorithm assessment \cite{lafay_new_2014, benetos2016detection}.
-The \textit{SimScene} software allows for several type of sequencing, from abstract ones (time indexes and amplitudes are drawn from random distributions) to  precise ones, where the time indexes and amplitudes for each event is set by the user.
+The sound scenes are generated with the \textit{SimScene} software\footnote{Open-source project available at: \url{https://bitbucket.org/mlagrange/simscene}}, \cite{rossignol_simscene:_2015}, a simulation software generating monaural sound mixtures in wav format from an isolated sounds database.
+This software has already been used in a wide range of experiments for sound detection algorithm assessment \cite{lafay_new_2014, benetos2016detection}.
+The \textit{SimScene} software allows the design of several type of sequencing, from \textit{abstract} ones (time indexes and amplitudes are drawn from random distributions) to  precise ones, where the time indexes and amplitudes for each event are set by the user.
 The latter type is considered in this study.
-As output, \textit{SimScene} generates an audio of the global sound mixture and an audio for each sound class present in the scene, which makes it possible to know their exact contributions in the scene, see Figure \ref{fig:example_simScene}.
+As output, \textit{SimScene} generates an audio file of the global sound mixture as well as for each sound class present in the scene, see Figure \ref{fig:example_simScene}.
+It makes it possible to know their exact contributions in the scene, in particular their sound levels.
 
@@
     \centering
-       \includegraphics[width=.8\linewidth]{./figures/exampleSimScene.pdf}
-    \caption{Spectrogram of a simple scene created with the \textit{SimScene} software with a sound background (road traffic in green) and 3 sound events (car horn in purple, passing car in red and whistling bird in blue).}
+       \includegraphics[width=.85\linewidth]{./figures/exampleSimScene.pdf}
+    \caption{Spectrogram of a simple scene created with the \textit{SimScene} software with one sound background (road traffic in blue) and 3 sound events (car horn in green, passing car in yellow and whistling bird in red).}
     \label{fig:example_simScene}
@@
 
-To transcribe the recordings in simulated scenes, a high quality sound database (wav format, 44.1 kHz sampling rate, high \textit{Signal Noise Ratio}) has been built-up from audio samples found online (\textit{freesound.org}) or with the help of an already existing sound database \cite{salamon2014dataset}.
-The sound database is composed of two categories of sound: the \textit{event} category, which includes 245 brief sound samples considered as salient, with a 1 to 20 seconds duration and classified among 21 sound classes (\textit{ringing bell, whistling bird, car horn, passing car, hammer, barking dog, siren, footstep, metallic noise, voice \dots}) and the \textit{background} (or \textit{texture}) category gathering 154 long duration sounds ($\approx$ 1m30), whose acoustic properties do not vary in time.
-This category includes among others \textit{whistling bird, crowd noise, rain, children playing in schoolyard, constant traffic noise} \dots~sound classes.
+To replicate the 74 recordings in simulated scenes, a high quality (wav format, 44.1 kHz sampling rate, high \textit{Signal Noise Ratio}) elemental corpus has been built-up from audio samples found online (\textit{freesound.org}) or with the help of an already existing sound database \cite{salamon2014dataset}.
+The elemental corpus is composed of two categories of sound: the \textit{event} category, which includes 245 brief sound samples considered as salient, with a 1 to 20 seconds duration and classified among 21 sound classes (\textit{ringing bell, whistling bird, car horn, passing car, hammer, barking dog, siren, footstep, metallic noise, voice \dots}) and the \textit{background} (or \textit{texture}) category gathering 154 long duration sounds ($\approx$ 1mn30), whose acoustic properties do not vary in time.
+This category includes among others \textit{whistling bird, crowd noise, rain, children playing in schoolyard, constant traffic noise} sound classes.
 Each sound class is composed of multiple samples (\textit{carHorn01.wav, carHorn02.wav \dots}) that are randomly chosen by the \textit{simScene} software to bring diversity.
-As the road traffic is the main component in urban environment and is the sound source of interest, recordings of car passages has been made on the Ifsttar's runway.
-The recordings has been made for 4 cars (Renault Scenic, Renault M\'egane, Renault Clio and Dacia Sandero),  at different speeds and gear ratios.
+As the road traffic is the main component in urban environment and is the sound source of interest, recordings of car passages have been made on the Ifsttar's runway.
+The recordings have been made for 4 different cars (Renault Scenic, Renault M\'egane, Renault Clio and Dacia Sandero), at different speeds and gear ratios.
 Overall, 103 car passages have been recorded.
-In order to avoid overfitting issues, the audio samples of the first two cars (Renault Scenic and Renault M\'egane) are included in the \textit{SimScene}'s sound database (50 audio files in total).
-The last 53 audio samples are dedicated to the dictionary design as part of NMF, see section \ref{part:dictionary_building}.
+In order to avoid overfitting issues, the audio samples of the first two cars (Renault Scenic and Renault M\'egane) are included in the \textit{SimScene}'s elemental corpus (50 audio files in total).
+The last 53 audio samples are dedicated to the dictionary design, $\mathbf{W}$ as part of NMF, see section \ref{part:dictionary_building}.
 A full description of the recordings can be found in \cite{gloaguen_creation_2017}.
-
-% \ml{ajouter une description sur les problématiques d'overfitting et de generalisation.
-Cela doit etre un rappel d'un texte ecrit dans la partie nmf.
-A ce titre, cela pourrait etre une extension de l'etude, on en reparle}.
 
@@
 \centering
-\caption{Mean \textit{Traffic Interfering Ratio} for each sound environment.}
+\caption{Mean \textit{Traffic Interfering Ratio} and its standard deviation for each sound environment.}
 \begin{tabular}{lc}
@@
 
-With this built-up database, the \textit{SimScene} software and the transcriptions of the recordings, 74 simulated sound scenes are generated, which have the temporal structure of the reference recordings.
+With this built-up corpus, the \textit{SimScene} software and the transcriptions of the recordings, 74 simulated sound scenes are generated, which have the same temporal structure of the reference recordings.
 The sound level of each sound class is adjusted manually on each sound scene to be faithful compared to the recorded scenes.
-To validate this adjustment, the mean \textit{Traffic Interfering Ratio} ($mTIR$) is calculated, see Table \ref{tab:mTIR}.
-It expresses, on all the scenes of an sound environment, the mean difference between the equivalent traffic sound levels of each scene, $L_{p,traffic}$, with the sound level of the \textit{interfering} sound class, $L_{p, interfering}$, which gathered all the other sound sources not related to the traffic, see Eq \ref{eq:mTIR}.
-It quantifies the predominance of the traffic component for the 4 types of sound environments.
+To check this adjustment, the mean \textit{Traffic Interfering Ratio} ($mTIR$) is calculated, see Table \ref{tab:mTIR}.
+It expresses, on all the scenes of a sound environment, the mean difference between the equivalent traffic sound levels of each scene, $L_{p,traffic}$, with the sound level of the \textit{interfering} sound class, $L_{p, interfering}$, which gathered all the other sound sources not related to the traffic, see Eq.
+\ref{eq:mTIR}.
+It quantifies the predominance of the traffic component for the 4 types of sound environments,
 
@@
 
-As it is the sound environment where the traffic is less present, the $mTIR$ is always negative for the \textit{park}.
+where $M$ is the number of available scenes for each sound environment ($M$ = 8 for \textit{park} environment, $M$ = 35 for \textit{quiet street} environment, $M$ = 23 for \textit{noisy street} environment and $M$ = 8 for \textit{very noisy street} environment).
+The $mTIR$ is always negative for the \textit{park} as it is the sound environment where the traffic is less present.
 The \textit{interfering} sound class is therefore the main sound source which is coherent with this kind of environment.
-In the case of the \textit{quiet street}, the $TIR$ can be positive or negative depending on the traffic presence on the scene.
-For the 2 others sound environments, when the traffic become the main sound source, $mTIR$ is always positive.
+The $mTIR$ can be positive or negative depending on the traffic presence on the scene in the case of the \textit{quiet street}.
+For the 2 others sound environments, when the traffic becomes the main sound source, $mTIR$ is always positive.
 
@@
 The total number of sound scenes tested is set at 40.
-The first half includes 20 30-seconds audio files, including 5 scenes that belong to the sound environment \textit{park}, 6 from \textit{quiet street}, 4 from \textit{noisy street} and 5 from \textit{very noisy street} chosen randomly among the recorded scenes.
-The second half is composed of the transcribed scenes corresponding to the same 20 reference scenes.
+This is less than the number of scenes available to ensure that each audio sample is sufficiently assess.
+The first half includes 20 30-seconds audio files corresponding to the real scenes, including 5 scenes that belong to the sound environment \textit{park}, 6 from \textit{quiet street}, 4 from \textit{noisy street} and 5 from \textit{very noisy street} chosen randomly among the recorded scenes.
+The second half is composed of the replicated scenes from the evaluation corpus corresponding to the same 20 reference scenes.
 In order to limit the duration of the test to preserve the concentration of the subjects, each subject listens to only a subset of 20 sound scenes.
@@
 
-The experimental design is elaborated following a partially Balanced Incomplete Block Design (PBIBD) \cite{john1977optimal} that determines the listening order of each participant based on fixed parameters (number of listeners, total number of audio, number of audio assessed by each listeners).
-The listening order is then built in a such way than each listener listens almost the same amount of recorded scenes and transcribed scenes to avoid statistical biais.
+The experimental design is elaborated following a partially Balanced Incomplete Block Design (PBIBD) \cite{john1977optimal} the audio allocated to each participant and the listening order.
+This process allows to each sound sample to be assess almost the same number of time and to avoid statistical biais.
 The experimental design and the listening order per participant are performed with the package \textit{sensoMineR} on the \textit{R} software \cite{le_sensominer:_2008}.\\
 
-The test was available online on the 8 February 2017 and the needed number of participant has been reached 12 days later.
+The test was available online from February 8th, 2017 and the needed number of participant (50) has been reached 12 days later.
 During the test, the participants had the possibility to listen to each scene as many times as wanted before assessing, without being able to change their judgment afterwards.
@@
 $62\%$ of the participants declared having no experience in the listening of urban sound mixtures.
-The average rate of all the transcribed scenes ($m_{transcribed} = 5.1 (\pm 1.6)$) is then similar to the recorded ones ($m_{recorded} = 4.9 (\pm 1.6)$).
-In complement, a Student's t-test is performed with the $H_0$ hypothesis which considers the similarity between the distribution of the average rates for the recorded and the transcribed scenes of each participant with a \textit{p-value} threshold of 5$\%$.
-Results are displayed in Table \ref{tab:ttest_type_of_scene}.
+The results show that the average score (with its standard deviation) of all the replicated scenes ($m_{replicated}$ = 5.1 ($\pm$ 1.6)) is close to the recorded ones ($m_{recorded}$ = 4.9 ($\pm$ 1.6)).
 
-\begin{table}[t]
-\centering
-\begin{tabular}{cccc}
- 	& \textbf{DOF} & \textbf{\textit{$\mid$ t $\mid$} }   & \textbf{\textit{p-value}} \\ 
-\hline
- type & 49 & 1.37 &  0.17   \\
-\hline
-\end{tabular}
-\caption{Student's t-test performed on the distribution of the average ratings for the recorded and replicated scenes}
-\label{tab:ttest_type_of_scene}
-\end{table}
+To determine whether a replicated scene is perceived in a similar way to that recorded, a Student's t-test is performed, for each scene, between the scores from the recorded sample and those from the replicated sample with the $H_0$ hypothesis which considers the similarity between the distribution of the rates for the recorded and the replicated scenes of each participant with a \textit{p-value} threshold of 5$\%$.
+The 20 performed t-tests show that the differences in the assessment of the replicated and the recorded scene, according to their degree or realism, is not significant (p-value$>5\%$).
 
-This result confirms that the recorded and the transcribed scenes are perceived in a similar way by the panel (\textit{p-value} $>$ $\alpha$).
 More details on the results can be found in  \cite{gloaguen_creation_2017}.
-
-As the transcribed sound corpus is sufficiently close to the recordings, these sound mixtures can be used to assess the performances of NMF to estimate the traffic sound level.
-
-\section{Experimental protocol}\label{part:expProtocol}
-
-The design of the experiment aims at evaluating in a meaningful manner the performance of the different systems.
-To do so, the 74 sound scenes, organized on 4 different types of sound environments are iteratively fed to the processing system which output an estimate of the equivalent sound level of the traffic within each scene $\tilde{L}_{p, traffic}$ (dB).
-This result is then compared to the reference sound level value given by the simulation process, $L_{p,traffic}$, see Figure \ref{fig:bloc_diagram_estimator}.
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=.9\linewidth]{figures/bloc_diagram_estimator.pdf}
-\caption{Bloc diagram of the experimental protocol.}
-\label{fig:bloc_diagram_estimator}
-\end{figure}
-
-\subsection{Metrics}
-
-The traffic sound levels, $\tilde{L}_{p,traffic}$, are compared to the exact values, $L_{p,traffic}$, through the Mean Absolute Error ($MAE$) \cite{willmott2005advantages}.
-The $MAE$ consists in the average of the absolute difference between the exact and the estimated sound levels,
-
-\begin{equation}
-MAE = \frac{\sum_{i = 1}^{M} \vert L_{p,traffic}^i - \tilde{L}_{p,traffic}^i \vert}{M}.
-\end{equation}
-
-and expresses the quality of the long-term reconstruction of the signal.
-It is then possible to express the $MAE$ error for each unique setting but, also, to average this metric according to the 4 sound environments to be able to estimate the optimal setting that offers the lowest error for all the sound environments:
-
-\begin{equation}
-mMAE = \frac{\sum_{i = 1}^4 MAE_i}{4},
-\end{equation}
-
-where the other experimental factors (method, $f_c$, $\mathbf{K}$, $\mathbf{w_t}$, $\beta$, threshold value $\mathbf{t_h}$) are fixed.
+As the perceived realism of the replicated and the recorded scenes are not significantly different, we consider that  these sound mixtures are relevant to assess the performances of NMF according to the traffic sound level estimate.
 
@@
 \centering
-\includegraphics[width=0.8\linewidth]{figures/extractionDictionary2.pdf}
-\caption{Dictionary building on a 3 second example of a passing car with $w_t = 1$ second.
-In this case, the dictionary $\mathbf{W}$ is made of 3 spectra, each representative of a texture frame of $w_t$ duration.}
+\includegraphics[width=0.7\linewidth]{figures/bloc_diagram_estimator.pdf}
+\caption{Bloc diagram of the different stages for the estimation of the traffic sound level.}
+\label{fig:bloc_diagram_estimator}
+\end{figure*}
+
+\section{Methods for the performance evaluation}\label{part:expProtocol}
+
+The experiment aims at evaluating in a meaningful manner the performance of the NMF approach.
+To do so, the 74 replicated sound scenes, organized on 4 different types of sound environments are iteratively fed to the estimator which output an estimate of the equivalent sound level of the traffic within each scene $\tilde{L}_{p, traffic}$ (dB).
+This result is then compared to the reference sound level value given by the simulation process, $L_{p,traffic}$, see Figure \ref{fig:bloc_diagram_estimator}.
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=\linewidth]{figures/extractionDictionary3.pdf}
+\caption{Dictionary building on a 3 second example of a passing car with $\mathbf{w_t} = 1$ second, in dashed line a temporal frame of 1 s.
+In this case, the dictionary $\mathbf{W}$ is made of 3 spectra, each representative of a texture frame of $\mathbf{w_t}$ duration.}
 \label{fig:example_dictionary}
@@
 
-\subsection{Methods}
+\subsection{Reference estimator}
 
+A reference estimator is necessary to be able to compare the performances of the different NMF-methods.
 As the road traffic is mainly composed of a low frequency content, a frequency low-pass filter (LP filter) is considered as baseline.
-It consists in filtering the sound scenes at different cut-off frequencies $f_c \in \lbrace$500, 1, 2k, 5k, 10k, 20k$\rbrace$ Hz.
-The remaining energy located in the pass-band is assimilated to road traffic,
+The estimation of the traffic sound level with this reference estimator is simply the sound level after low-pass filtering,
 
@@
 
+Different cut-off frequencies have been chosen such as $f_c \in \lbrace$500, 1k, 2k, 5k, 10k, 20k$\rbrace$ Hz.
 The experimental factors related to this estimator are summarized in Table \ref{tab:experimental_factorsFilter}.
@@
 The second estimator is based on the three NMF formula presented in part \ref{part:nmf} (see Figure \ref{fig:bloc_diagram_nmf}).
-Multiple experimental factors take part in these cases where each of them having different modalities.
+Multiple experimental factors are involved in this second estimator where each of them having different modalities.
 
@@
 
-\subsection{Dictionary building for NMF methods}\label{part:dictionary_building}
+\subsection{Dictionary building for NMF-methods}\label{part:dictionary_building}
 
 The dictionary is designed from a sound database specially dedicated to this task.
-To prevent any overfitting issues, it contains 53 audio files of the 2 other cars (Dacia Sandero and Renault Clio) recorded on the runway, see part \ref{part:simScene}.
-Those recordings have not been considered for the creation of the evaluation dataset.
+To prevent any overfitting issues, it contains the 53 audio files of the 2 cars not included  in the creation of the evaluation corpus, see part \ref{part:simScene}.
 
-First the spectrogram of each audio file is computed (window $w = 2^{12}$ with 50 $\%$ overlap).
-The spectrogram is then cut in multiple temporal frames of $w_t  = \lbrace$0.5, 1$\rbrace$ second duration.
-In each cut spectrogram, the root mean square on each frequency bin is calculated to obtained a spectrum of $F \times 1$ dimension.
-This method allows the description of the audio sample with finer spectra and then having the different characteristic pitches of the traffic spectra.
+First the spectrogram of each audio file is computed ($w = 2^{12}$ sample points with 50 $\%$ overlap).
+The spectrogram is then cut in multiple temporal frames of $\mathbf{w_t}  \in \lbrace$0.5, 1$\rbrace$ second duration.
+In each frame, the root mean square on each frequency bin is calculated to obtained a spectrum of $F \times 1$ dimension.
+This method allows the description of the audio sample with a finer spectra and then having the different characteristic pitches of the traffic spectra.
 An illustrative example on a 3 seconds sample is displayed in Figure \ref{fig:example_dictionary}.
-From the 53 audio files, we obtain respectively for $\mathbf{w_t}\in$ $\lbrace$0.5, 1$\rbrace$ second, 2218 and 1109 elements.
+From the 53 audio files, we obtain respectively 2218 elements for $\mathbf{w_t} = 0.5$ second and 1109 elements for $\mathbf{w_t} = 1$ second.
 
-A $\mathbf{K}$-means clustering algorithm is applied to reduce these dimensions to $\mathbf{K} \in \lbrace$25, 50, 100, 200$\rbrace$ in order to avoid redundant information and decrease the computation time.
-The resulting $\mathbf{K}$ clusters are taken as the elements of $\mathbf{W}$.
-The obtained dictionary is expressed in third octave bands to decrease the dimensionality.
-Each basis of $\mathbf{W}$ is then normalized as $ \Vert \mathbf{w}_k \Vert = 1$ where $\Vert \bullet \Vert$ is the $\ell_1$ norm.
+A $\mathbf{K}$-means clustering algorithm is applied to those elements to reduce these dimensions to $\mathbf{K} \in \lbrace$25, 50, 100, 200$\rbrace$ in order to avoid redundant information and decrease the computation time.
+The resulting $\mathbf{K}$ cluster centroids are taken as the elements of $\mathbf{W}$.
+Each basis of $\mathbf{W}$ is then normalized as $ \Vert \mathbf{w} \Vert = 1$ where $\Vert \bullet \Vert$ is the $\ell_1$ norm.
 Table \ref{tab:experimental_factorsNMF} summarizes the different modalities of the two experimental factors ($\mathbf{K}$ and $\mathbf{w_t}$).
@@
 
-NMF is performed for 3 $\beta$-divergences: $\beta$ = 2 (euclidean distance), $\beta$ = 1 (Kullback-Leibler divergence) and $\beta = 0$ (Itakura-Sa\"ito divergence).
-The spectrogram $\mathbf{V}$ and the dictionary $\mathbf{W}$ are represented in a logarithmic scale through a third band octave representation that reduces the high frequency predominance where the traffic component is absent.
+NMF is performed for 3 $\beta$-divergences: $\beta$ = 2 (Euclidean distance), $\beta$ = 1 (Kullback-Leibler divergence) and $\beta = 0$ (Itakura-Sa\"ito divergence).
+The spectrogram $\mathbf{V}$ and the dictionary $\mathbf{W}$ are expressed in a logarithmic scale through a third octave band representation that reduces the high frequency predominance where the traffic component is absent.
 In addition, as the number of frequency bins is reduced ($F$ = 29), the computation time is reduced too.
@@
 Finally, for TI-NMF, the number of settings is much higher (2976) due to the higher cardinality of the set of threshold values (4 $\times$ 1 $\times$ 2 $\times$ 4  $\times$ 3 $\times$ 31).
-The experimental factors and their different modalities are displayed on Tables \ref{tab:experimental_factorsFilter} and \ref{tab:experimental_factorsNMF}.
+The experimental factors and their different modalities are displayed in Tables \ref{tab:experimental_factorsFilter} and \ref{tab:experimental_factorsNMF}.
 
@@
 \begin{equation}
-\tilde{L}_{p, traffic} = 20\log_{10}\frac{p_{rms}}{p_0},
+\tilde{L}_{p, traffic} = 20\log\left(\frac{p_{rms}}{p_0}\right),
 \end{equation}
@@
 
+\subsection{Metrics}
+
+The traffic sound levels, $\tilde{L}_{p,traffic}$, are compared to the exact values, $L_{p,traffic}$, through the Mean Absolute Error ($MAE$) \cite{willmott2005advantages}.
+The $MAE$ consists in the average across of the absolute difference between the exact and the estimated sound levels,
+
+\begin{equation}
+MAE_j = \left[\frac{\sum_{i = 1}^{M} \vert L_{p,traffic}^i - \tilde{L}_{p,traffic}^i \vert}{M}\right]_j,
+\end{equation}
+
+for each setting $j$.
+But it is also possible to average this metric according to the 4 different sound environments, through the mean $MAE$ error $mMAE$, to estimate the optimal setting that offers the lowest error for all the sound environments:
+
+\begin{equation}
+mMAE = \frac{\sum_{i = 1}^4 MAE_i}{4},
+\end{equation}
+
+where the other experimental factors (estimator, $f_c$, $\mathbf{K}$, $\mathbf{w_t}$, $\beta$, threshold value $\mathbf{t_h}$) are fixed.
+
 \section{Results and discussion}\label{part:results}
 
-Table \ref{tab:results} summarizes the lowest $mMAE$ errors according to the method (LP filter, SUP-NMF, SEM-NMF and TI-NMF) and $\beta$ with the others corresponding modalities.
+Table \ref{tab:results} summarizes the lowest $mMAE$ errors according to the estimator (LP filter, NMF) and $\beta$ with the best setting of the experimental factors.
 
@@
 \centering
-\caption{Best $mMAE$ errors according to the experimental factors $\beta$ and \textit{method} (in bold letter, the lowest error).}
+\caption{Best $mMAE$ errors according to the experimental factor $\beta$ and the traffic sound level assessment method (in bold letter, the lowest error).}
 \begin{tabular}{@{}ccccccc@{}}
@@
 
-The LP filter with $f_c = 20$ kHz cut-off frequency is equivalent to considerate the sound level of the entire scene without specific distinction between the sound sources.
+The LP filter with $f_c = 20$ kHz cut-off frequency is equivalent to consider the sound level of the entire scene without specific distinction between the sound sources.
 The error is then important with a high standard deviation ($mMAE =$ 3.76 ($\pm$ 4.35) dB).
@@
 The best result is obtained for TI-NMF ($MAE$ = 1.24 ($\pm$ 1.24) dB) with $\beta = 2$, $\mathbf{K}$ = 200, $\mathbf{w_t}$ = 0.5 s and as threshold value $\mathbf{t_h}$ = 0.32.
-This combination of settings offers the most fitted method to be adapted to all the sound environments.
+This combination of settings offers the most efficient method adapted to the different sound environments.
 Furthermore, on the dictionary creation, only SEM-NMF proposes the same dictionary design for all the best methods according to $\beta$.
@@
 \includegraphics[width=\linewidth]{figures/mea_grafic_bar.pdf}
-\caption{$MAE$ errors according to each sound environment for the best combination of the LP filter ($f_c$ = 500 Hz), SUP-NMF ($\beta$ = 2, $\mathbf{K}$ = 25, $\mathbf{w_t}$ = 1 second), SEM-NMF ($\beta$ = 1, $\mathbf{K}$ = 200, $\mathbf{w_t}$ = 1 second) and TI-NMF ($\beta$ = 2, $\mathbf{K}$ = 200, $\mathbf{w_t}$ = 0.5 second, $\mathbf{t_{h}} = 0.32$).}
+\caption{$MAE$ errors with the standard deviations according to each sound environment for the best combination of the LP filter ($f_c$ = 500 Hz), SUP-NMF ($\beta$ = 2, $\mathbf{K}$ = 25, $\mathbf{w_t}$ = 1 second), SEM-NMF ($\beta$ = 1, $\mathbf{K}$ = 200, $\mathbf{w_t}$ = 1 second) and TI-NMF ($\beta$ = 2, $\mathbf{K}$ = 200, $\mathbf{w_t}$ = 0.5 second, $\mathbf{t_{h}} = 0.32$).}
 \label{fig:mae_env}
@@
 
-Aside SEM-NMF, all the methods show the same error evolution the same error evolution: a decrease of the error with the increase of the traffic predominance.
-On contrary, SEM-NMF show an almost constant error for all 4 sound environments.
+Aside SEM-NMF, all the methods show the same error evolution: a decrease of the error with the increase of the traffic predominance.
+On contrary, SEM-NMF shows an almost constant error for all 4 sound environments.
 The LP filter error is mainly important for environments where the traffic is less present.
@@
 The interfering sound class is then wrongly considered as traffic component.
-In the opposite, for noisy and very noisy environments, the performances of the LP filter are good ($MAE$ $<$ 1 dB).
+On the opposite, for noisy and very noisy environments, the performances of the LP filter are good ($MAE$ $<$ 1 dB).
 The errors are here due to a high deletion of the traffic energy by the filter while it becomes the main sound source.
@@
 \ref{eq:min-D-WH}, the dictionary's elements are used to model the other sound sources which can not allow a rightful approximation of the traffic component.
-In the opposite, for \textit{noisy} and \textit{very noisy} environments, SUP-NMF identifies correctly the traffic components ($MAE$ $<$ 0.6 dB) as it is the main source.
+On the opposite, for \textit{noisy} and \textit{very noisy} environments, SUP-NMF identifies correctly the traffic components ($MAE$ $<$ 0.6 dB) as it is the main source.
 In the case of SEM-NMF, adding the mobile dictionary, $\mathbf{W_r}$, makes it possible to include the other sound sources not present in the dictionary.
@@
 For very noisy environment the error is even very low ($MAE=$ 0.28 dB).
-With, on average 198 elements in $\mathbf{W}_{traffic}$, almost all the elements present in $\mathbf{W'}$ can be considered as traffic elements.
+With on average 198 elements in $\mathbf{W}_{traffic}$, almost all the elements present in $\mathbf{W'}$ can be considered as traffic elements.
 
@@
 The proposed approach, named Thresholded Initialized NMF, achieved the lowest error in the evaluation corpus.
-Consequently, in the case where the location or the type of sound environments the sensors are monitoring cannot be identified (for instance within a mobile measurement framework), TI-NMF appears to be the method of choice.
+Consequently, in the case where the location or the type of sound environments the sensors are monitoring cannot be identified (for instance within a mobile measurement framework), TI-NMF appears to be the most appropriate method.
 If the sound environment can be identified through a prior analysis, or based on positioning data \cite{can2015noise,lavandier2016urban}, it should be possible to adapt the estimation procedure by selecting the most efficient approach in order to further reduce the error in the estimated road traffic sound levels.
@@
 Further analyses are required to extend the proposed method to other sound sources, such as birds or voices sounds, which can conveniently be done by replacing or adding elements in the dictionary.
-This would prove useful in the context of muti-source noise mapping that is  gaining interest \cite{aumond2017Probabilistic, aletta2015soundscape}.
-Finally, the selected parameters stand for the corpus of sound mixtures of this study.
-Further analyses on various corpus of sound scenes are needed to evaluate the robustness of the method, and select the most relevant approaches for specific sound environments (predominance of water or industrial sounds, rural environments \dots).
+This utilization would be useful in the context of muti-source noise mapping that is gaining interest \cite{aumond2017Probabilistic, aletta2015soundscape}.
+Finally, the parameters selected in this study are valid for this evaluation corpus.
+Further analyses on various corpus of sound scenes are needed to evaluate the robustness of the method and select the most relevant approaches for specific sound environments (predominance of water or industrial sounds, rural environments \dots).
 
-For reproducibility purposes, the evaluation corpus, the experimental protocol and the programs developed under the Matlab software are available online \footnote{\url{https://github.com/jean-remyGloaguen/articleNmfTrafficSimScene2018}} and as the sound dataset \footnote{\url{https://zenodo.org/record/1184443#.WqfVb3wiGpo}}.
+For reproducibility purposes, the evaluation corpus, the experimental protocol and the programs developed under the Matlab software are available online \footnote{\url{https://github.com/jean-remyGloaguen/articleNmfTrafficSimScene2018}} as the sound dataset \footnote{\url{https://zenodo.org/record/1184443#.WqfVb3wiGpo}}.
 

### 2018-04-06 experiment/nmfthgr2estimation.m

@@ -11,7 +11,8 @@ function [config, store, obs] = nmfthgr2estimation(config, setting, data)
 % Date: 01-Dec-2017                                                                
                                                                                    
 % Set behavior for debug mode                                                      
-if nargin==0, NMFThresholdGrafic('do', 2, 'mask', {}); return; else store=[]; obs=[]; end
+if nargin==0, NMFThresholdGrafic('do', 2, 'mask', {2 0 4 3 1, ...
+	3 0 6 0 3, 0 3 0 1 0, 0 1:3 1:2 4}); return; else store=[]; obs=[]; end
 
 dataset = setting.dataset;
 type = setting.type;

### 2018-04-06 paper/articleNmfTrafficSimScene2018.tex
minor improvement

@@
 The evaluation corpus is then designed by replicating the reference corpus according to the transcription with the help of the elemental corpus, see Figure \ref{fig:transcription}.
-More precisely, the reference corpus is composed of 76 recordings from 2 to 5 min, achieved in the \nth{13} district of Paris (France) at 19 different locations \footnote{Recordings were made as part of the Grafic project funded by Ademe}, see Figure \ref{fig:map_grafic}, which cover various sound environments.
+
+The reference corpus is composed of 76 recordings from 2 to 5 min, achieved in the \nth{13} district of Paris (France) at 19 different locations \footnote{Recordings were made as part of the Grafic project funded by Ademe}, see Figure \ref{fig:map_grafic}, which cover various sound environments.
 A complete description of the experimental protocol can be found in \cite{aumond_modelling_2017}.
@@
 
-To evaluate the level of realism of the corpus of transcribed urban sound scenes, a perceptual test is considered.
+To evaluate the level of realism of the evaluation corpus composed of replicated urban sound scenes, a perceptual test is considered.
 
-The perceptual test is conducted with a panel of 50 subjects that are asked to assess the level of realism on a 7-point scale (1 is \textit{not realistic at all}, 7 is \textit{very realistic}) of a set of transcribed and recorded scenes.
+The perceptual test is conducted with a panel of 50 subjects that are asked to assess the level of realism on a 7-point scale (1 is \textit{not realistic at all}, 7 is \textit{very realistic}) of a set of replicated and recorded scenes.
 The total number of sound scenes tested is set at 40.
@@
 The results show that the average score (with its standard deviation) of all the replicated scenes ($m_{replicated}$ = 5.1 ($\pm$ 1.6)) is close to the recorded ones ($m_{recorded}$ = 4.9 ($\pm$ 1.6)).
-
 To determine whether a replicated scene is perceived in a similar way to that recorded, a Student's t-test is performed, for each scene, between the scores from the recorded sample and those from the replicated sample with the $H_0$ hypothesis which considers the similarity between the distribution of the rates for the recorded and the replicated scenes of each participant with a \textit{p-value} threshold of 5$\%$.

### 2018-04-09 paper/articleNmfTrafficSimScene2018.tex
aknowledgment + funding add

@@
 
-The perceptual test is conducted with a panel of 50 subjects that are asked to assess the level of realism on a 7-point scale (1 is \textit{not realistic at all}, 7 is \textit{very realistic}) of a set of replicated and recorded scenes.
+The perceptual test is conducted with a panel of 50 listeners that are asked to assess the level of realism on a 7-point scale (1 is \textit{not realistic at all}, 7 is \textit{very realistic}) of a set of replicated and recorded scenes.
 The total number of sound scenes tested is set at 40.
@@
 The experimental design is elaborated following a partially Balanced Incomplete Block Design (PBIBD) \cite{john1977optimal} the audio allocated to each participant and the listening order.
-This process allows to each sound sample to be assess almost the same number of time and to avoid statistical biais.
+This process allows to each sound sample to be assessed almost the same number of time and to avoid statistical biais.
 The experimental design and the listening order per participant are performed with the package \textit{sensoMineR} on the \textit{R} software \cite{le_sensominer:_2008}.\\
@@
 
-For reproducibility purposes, the evaluation corpus, the experimental protocol and the programs developed under the Matlab software are available online \footnote{\url{https://github.com/jean-remyGloaguen/articleNmfTrafficSimScene2018}} as the sound dataset \footnote{\url{https://zenodo.org/record/1184443#.WqfVb3wiGpo}}.
+For reproducibility purposes, the evaluation corpus, the experimental protocol and the programs developed under the Matlab software are available online \footnote{\url{https://github.com/jean-remyGloaguen/articleNmfTrafficSimScene2018}} as the sound dataset \footnote{\url{https://zenodo.org/record/1184443}}.
+
+\section*{Acknowledgements}
+The authors would like to thank Pierre Aumond and Catherine Lavandier from the University of Cergy-Pontoise for transmitting us the data of the \textit{Grafic} project.
+
+\section*{Funding}
+This study is co-funded by Ifsttar and Pays de la Loire region.
+% with a partial funding from the ANR under project reference ANR-16-CE22-0012.
 

### 2018-07-31 revision/responseToreviewers.tex
Responses to reviewers

\documentclass[10pt]{article}

\title{Reply to reviewer concerning submission APAC$\_$2018$\_$311 'Road traffic sound level estimation from realistic urban sound mixtures by Non-negative Matrix Factorization' in Applied Acoustics.}

\begin{document}

\maketitle

As a preamble, we would like to thank the editor and the reviewer for the comments and suggestions. Following these comments, we made several changes to the article, which are summarized here. The next sections list our answers to the reviewer’s comments, with references to the revised manuscript (page, column, and paragraph) where appropriate.

\section{Answers to Reviewer 1}

An interesting approach to estimate the sound level of road traffic in an urban sound environment composed of a mixture of sound sources is presented in the reviewed paper.

The work is well planned and structured. It has an adequate length and the results are interesting.

I only have some minor indications that I think they can improve a little bit this paper.

\begin{enumerate}
\item The use of the letter beta should be clarified because its use in expressions (5, 8a, 8b and 8c) and in lines 104-105, seems not to be the same that its use in expressions (4a, 4b y 4c) or, for example, in tables 3 and 4.\\
$\rightarrow$ Done, The $\beta$ value is the same, the choice of the $\beta$-divergence in the equation 4 influes the way the expressions 5 and 8 are then updated. This has been precised . 'Thus, the choice of the $\beta$-divergence in the equation affects how the matrix $\mathbf{H}$ is updated.' (section 2.1).

\item In relation with table 4, it may be appropriate that it includes some more complete representation or table of results, for the reader interested in analyzing in more detail the effects that the different variables analyzed may have on errors. For example, one in which, varying one of the variables analyzed, the others remain constant.

$\rightarrow$ We add a part dedicated to this where the influence of the number of element on $\mathbf{W}$ and the temporal window $w_t$ influence the NMF behavior.

\item References: The bibliographic revision is good, but maybe, some references can be improved.

For example, reference (3), because it is a congress, not a peer reviewed paper. Only as an example, you can considerer the references: https://doi.org/10.1016/j.apacoust.2010.06.010 and https://doi.org/10.1016/j.scitotenv.2011.10.018 .

In line 67, it is indicated that “Many works…” but only two references are included. Author may considerer others, for example: 
http://dx.doi.org/10.1016/j.apacoust.2016.05.020 and https://doi.org/10.2478/v10168-012-0054-z

At the end of the lines 30-31, it is possible that some general references are needed. For example, 
https://doi.org/10.3813/AAA.918456 and http://dx.doi.org/10.1016/j.apacoust.2016.04.018.\\
$\rightarrow$ Done, some references have been add to the different locations indicated.

\item Moreover, there are references that just have been submitted to publication. \\
$\rightarrow$ Done for one reference, the other is still in revision.

\end{enumerate}

\end{document}

### 2018-07-31 revision/articleNmfTrafficSimScene2018_v2.tex
Responses to reviewers

--- 2018-04-09 paper/articleNmfTrafficSimScene2018.tex
@@
 Despite their unanimously recognized interest, noise maps suffer from some limitations.
-The computing cost required to produce noise maps at the city scale calls simplifications of the numerical tools and the simulation models that both generate uncertainties \cite{van_leeuwen_noise_2015}.
+The computing cost required to produce noise maps at the city scale calls simplifications of the numerical tools and the simulation models that both generate uncertainties \cite{arana2011precision,van_leeuwen_noise_2015}.
 Data collection is itself also a vector of uncertainty.
@@
 
-Noise measurements are thus increasingly used in addition to simulation to describe urban noise environments.
+Noise measurements are thus increasingly used in addition to simulation to describe urban noise environments \cite{gozalo2016study,zannin2013characterization,can_exploring_2012}.
 Several measurement set-ups have been proposed in the last years, including  mobile measurements with high quality microphones \cite{manvell2004sadmam, can2014measurement}, participative sensing through dedicated smartphone applications \cite{picaut:halshs-01565214, ventura2017evaluation}, or the development of fixed-sensor networks.
@@
 
-Many works have dealt with the detection \cite{luitel2016sound} or the recognition \cite{defreville_automatic_2006} of sound events in environmental sound scenes.
+Many works have dealt with the classification \cite{valero2012hierarchical,shen_environmental_2012}, the detection \cite{luitel2016sound,mesaros2015sound} or the recognition \cite{defreville_automatic_2006,parascandolo_recurrent_2016} of urban sound events.
 In these cases, a two-step scheme is followed where audio samples are described with a set of features (Mel Frequency Cepstral Coefficient, MPEG-7 descriptors \dots) and classified with the help of a classifier (Gaussian Mixtures Models, Artificial Neural Network \dots) \cite{chu2008environmental, cowling_comparison_2003}.
@@
 with $\gamma(\beta) = \frac{1}{2-\beta},$ for $\beta < 1$, $ \gamma(\beta) = 1$, for $\beta \in \left[1,2\right]$ and $\gamma(\beta) = \frac{1}{\beta-1}$ for $\beta > 2$.
+Thus, the choice of the $\beta$-divergence in the equation \ref{eq:divBetaGenerale} affects how the matrix $\mathbf{H}$ is updated.
 The $A\otimes B$ and $A/B$ operators represent the Hadamard product and ratio.
@@
 \centering
-\includegraphics[width=0.8\linewidth]{figures/realistic_urban_sound_scene_design.pdf}
+\includegraphics[width=0.8\linewidth]{./figures/realistic_urban_sound_scene_design.pdf}
 \caption{Bloc diagram of the design of the evaluation corpus}

### 2018-08-02 paper/revision/responseToreviewers.tex
review done, ready to proofread

--- 2018-07-31 revision/responseToreviewers.tex
@@
 
-$\rightarrow$ We add a part dedicated to this where the influence of the number of element on $\mathbf{W}$ and the temporal window $w_t$ influence the NMF behavior.
+$\rightarrow$ We added a part dedicated to this where the evolution of the cost functions and the $MAE$ errors are displayed according to the iteration (section 5).
 

### 2018-08-02 paper/revision/articleNmfTrafficSimScene2018_v2.tex
review done, ready to proofread

--- 2018-07-31 revision/articleNmfTrafficSimScene2018_v2.tex
@@
 
+\begin{figure*}[t]
+\centering
+\begin{subfigure}{0.45\textwidth}
+\includegraphics[width=\linewidth]{./figures/evolutionCost_AA.pdf}
+\caption{Evolution of the cost function.} \label{fig:grafic_cost}
+\end{subfigure}
+\begin{subfigure}{0.45\textwidth}
+\includegraphics[width=\linewidth]{./figures/evolutionMAE_AA.pdf}
+\caption{Evolution of the $MAE$ errors.} \label{fig:grafic_mae}
+\end{subfigure}
+\caption{Normalized reconstruction error of the global sound mixtures \subref{fig:grafic_cost} and the traffic component \subref{fig:grafic_mae} for each sound environment.} \label{fig:grafic_cost_mae}
+\end{figure*}
+
 Despite a fixed dictionary composed of traffic spectra, SUP-NMF fails to identify correctly the traffic component particularly for \textit{park} ($MAE$ = 6.42 dB) environments.
@@
 With on average 198 elements in $\mathbf{W}_{traffic}$, almost all the elements present in $\mathbf{W'}$ can be considered as traffic elements.
+Considering a unique dictionary fitted to the sound scene under evaluation thus makes TI-NMF very effective when traffic is predominant, while the thresholding step make it possible to discards the elements of the dictionary that deviate too much when the traffic is less present.
 
-Considering a unique dictionary fitted to the sound scene under evaluation thus makes TI-NMF very effective when traffic is predominant, while the thresholding step make it possible to discards the elements of the dictionary that deviate too much when the traffic is less present.
+\begin{figure*}[t!]
+\centering
+\begin{subfigure}{0.45\textwidth}
+\includegraphics[width=\linewidth]{./figures/evolutionK_AA.pdf}
+\caption{Influence of the experimental factor $K$ (TI-NMF, $w_t$ = 0.5 s, $\beta$ = 2, $t_h$ = 0.32.} \label{fig:grafic_K}
+\end{subfigure}
+\begin{subfigure}{0.45\textwidth}
+\includegraphics[width=\linewidth]{./figures/evolutionWt_AA.pdf}
+\caption{Influence of the experimental factor $K$ (TI-NMF, $K$ = 200 s, $\beta$ = 2, $t_h$ = 0.32.} \label{fig:grafic_wt}
+\end{subfigure}
+\caption{Influence of the number of elements $K$ \subref{fig:grafic_K} and the temporal frame $w_t$ \subref{fig:grafic_wt} on the $mMAE$ error with the best TI-NMF obtained in Table \ref{tab:results}.} \label{fig:grafic_K_wt}
+\end{figure*}
+
+To visualized the behavior of this approach, the evolution of the normalized cost function ($D(\mathbf{V}\Vert \mathbf{WH})/(F\times N)$ as the duration of each audio sample is different) and the error $MAE$ along the iteration are displayed for each sound environment with the TI-NMF in the Figure \ref{fig:grafic_cost_mae}.
+After the 400 iterations, the cost functions for each sound environment, is very low meaning that the approximations of $\mathbf{V}$ by $\mathbf{WH}$ are very similar.
+However, they did not reach the convergence as the cost functions continue to decrease.
+Nevertheless, these behaviors have to be seen with the evolution of the $MAE$ error in Figure \ref{fig:grafic_mae}.
+For each sound environment, the traffic error reconstruction is constant after the 400 iterations which mean that the updates of $\mathbf{W}$ and $\mathbf{H}$ are dedicated to the interfering sound sources.
+Also, even if the cost function is higher for the \textit{very noisy street} than the \textit{park} sound environment, the $MAE$ error is lower due to the high presence of the \textit{traffic} component in it.
+
+Finally, for TI-NMF, the influence of the number of elements $K$ in $\mathbf{W}$ and the size of the temporal frame $w_t$ on the $mMAE$ error are summarized in Figure \ref{fig:grafic_K_wt}. 
+Naturally, with the increase of the number of basis in $\mathbf{W}$, the traffic error reconstruction decrease as it allows more basis to describe the traffic sound source.
+If this improvement is significant between 25 and 100 elements, its influence decrease between 100 and 200 elements suggesting that the increase of the size of $K$ will not improve a lot the quality of the estimation.
+In opposite, the temporal frame is not an influential  parameter for TI-NMF.
+As the matrices are updates, it is logical that the variations between the initial dictionary $\mathbf{W_0}$ due to this experimental factor disappear.
 

### 2018-08-03 revision/responseToreviewers.tex
Re-submission before acceptation

@@
 Following these comments, we made several changes to the article, which are summarized here.
-The next sections list our answers to the reviewer’s comments, with references to the revised manuscript (page, column, and paragraph) where appropriate.
+The next sections list our answers to the reviewer's comments, with references to the revised manuscript where appropriate.
 
@@
 \item The use of the letter beta should be clarified because its use in expressions (5, 8a, 8b and 8c) and in lines 104-105, seems not to be the same that its use in expressions (4a, 4b y 4c) or, for example, in tables 3 and 4.\\
-$\rightarrow$ Done, The $\beta$ value is the same, the choice of the $\beta$-divergence in the equation 4 influes the way the expressions 5 and 8 are then updated.
-This has been precised .
-'Thus, the choice of the $\beta$-divergence in the equation affects how the matrix $\mathbf{H}$ is updated.' (section 2.1).
+$\rightarrow$ Done: the $\beta$ value in 4 is the same that in the expressions 5 and 8.
+This has been precised: 'Thus, the choice of the $\beta$-divergence in the equation affects how the matrix $\mathbf{H}$ is updated.' (section 2.1).
 
@@
 
-$\rightarrow$ We add a part dedicated to this where the influence of the number of element on $\mathbf{W}$ and the temporal window $w_t$ influence the NMF behavior.
+$\rightarrow$ We added a part, between the lines 610 and 642, dedicated to this where the evolution of the cost functions and the $MAE$ errors are displayed according to the iteration (section 5, Figure 10) and the evolution of the error according to $K$ and $w_t$ are exposed (Figure 11).
 
@@
 https://doi.org/10.3813/AAA.918456 and http://dx.doi.org/10.1016/j.apacoust.2016.04.018.\\
-$\rightarrow$ Done, some references have been add to the different locations indicated.
+$\rightarrow$ Done, some references, suggested by the reviewers, have been added to the different locations indicated.
 
 \item Moreover, there are references that just have been submitted to publication. \\
-$\rightarrow$ Done for one reference, the other is still in revision.
+$\rightarrow$ Done for one reference, the other is in revision.
 

### 2018-08-03 revision/articleNmfTrafficSimScene2018_v2.tex
Re-submission before acceptation

@@
 Different cut-off frequencies have been chosen such as $f_c \in \lbrace$500, 1k, 2k, 5k, 10k, 20k$\rbrace$ Hz.
-The experimental factors related to this estimator are summarized in Table \ref{tab:experimental_factorsFilter}.
+The experimental factors related to this estimator are summarized in Table \ref{tab:experimental_factors}.
 
@@
 Each basis of $\mathbf{W}$ is then normalized as $ \Vert \mathbf{w} \Vert = 1$ where $\Vert \bullet \Vert$ is the $\ell_1$ norm.
-Table \ref{tab:experimental_factorsNMF} summarizes the different modalities of the two experimental factors ($\mathbf{K}$ and $\mathbf{w_t}$).
+Table \ref{tab:experimental_factors} summarizes the different modalities of the two experimental factors ($\mathbf{K}$ and $\mathbf{w_t}$).
 
@@
 Finally, for TI-NMF, the number of settings is much higher (2976) due to the higher cardinality of the set of threshold values (4 $\times$ 1 $\times$ 2 $\times$ 4  $\times$ 3 $\times$ 31).
-The experimental factors and their different modalities are displayed in Tables \ref{tab:experimental_factorsFilter} and \ref{tab:experimental_factorsNMF}.
+The experimental factors and their different modalities are displayed in Table \ref{tab:experimental_factors}.
 
@@
 \begin{table*}[t]
-\caption{Experimental factors and their modalities for the frequency low-pass filter baseline estimator.}
-\centering
-\begin{tabularx}{17.5cm}{L{3cm}@{}C{12cm}@{}C{2cm}@{}}
-	\hline
-    \textbf{\begin{tabular}[c]{@{}l@{}}experimental \\ factors\end{tabular}} & \textbf{modalities} & \begin{tabular}[c]{@{}C{2cm}@{}}\textbf{number of}\\ \textbf{modalities}\end{tabular}\\ \toprule
-\end{tabularx}
-
-\begin{tabularx}{17.5cm}{L{3cm}@{}C{3cm}@{}@{}C{3cm}@{}@{}C{3cm}@{}@{}C{3cm}@{}C{2cm}@{}}
-    \textbf{sound environment} & \begin{tabular}[c]{@{}c@{}}park\\ 'P'\end{tabular} & \begin{tabular}[c]{@{}c@{}}quiet street\\ 'Q'\end{tabular} & \begin{tabular}[c]{@{}c@{}}noisy street\\ 'N' \end{tabular}& \begin{tabular}[c]{@{}c@{}}very noisy street\\ 'vN'\end{tabular} & 4\\
-\end{tabularx}
-
-\begin{tabularx}{17.5cm}{L{3cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}C{2cm}@{}}
-\rowcolor[HTML]{C0C0C0}
-   $\mathbf{f_c}$ (kHz) & 0.5 & 1 & 2 &  5 & 10 & 20 & 6\\
-   \bottomrule
-\end{tabularx}
-
-\label{tab:experimental_factorsFilter}
-\end{table*}
-
-\begin{table*}[t]
 \centering
@@
 \begin{tabularx}{17.5cm}{L{3cm}@{}C{3cm}@{}@{}C{3cm}@{}@{}C{3cm}@{}@{}C{3cm}@{}C{2cm}@{}}
+	\rowcolor[HTML]{C0C0C0}
     \textbf{sound environment} & \begin{tabular}[c]{@{}c@{}}park\\ 'P'\end{tabular} & \begin{tabular}[c]{@{}c@{}}quiet street\\ 'Q'\end{tabular} & \begin{tabular}[c]{@{}c@{}}noisy street\\ 'N' \end{tabular}& \begin{tabular}[c]{@{}c@{}}very noisy street\\ 'vN'\end{tabular} & 4\\
@@
 
-\begin{tabularx}{17.5cm}{L{3cm}@{}C{4cm}@{}@{}@{}C{4cm}@{}@{}C{4cm}@{}C{2cm}@{}}
+\begin{tabularx}{17.5cm}{L{3cm}@{}C{3cm}@{}@{}C{3cm}@{}@{}C{3cm}@{}@{}C{3cm}@{}C{2cm}@{}}
+  \textbf{method} &  LP Filter & SUP NMF & SEM NMF & TI NMF & 4\\
+\end{tabularx}
+
+\begin{tabularx}{17.5cm}{L{3cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}C{2cm}@{}}
 \rowcolor[HTML]{C0C0C0}
-  \textbf{method} & SUP NMF & SEM NMF & TI NMF & 3\\
+   $\mathbf{f_c}$ (kHz) & 0.5 & 1 & 2 &  5 & 10 & 20 & 6\\
 \end{tabularx}
@@
 \end{tabularx}
-\label{tab:experimental_factorsNMF}
+\label{tab:experimental_factors}
 \end{table*}
@@
 With on average 198 elements in $\mathbf{W}_{traffic}$, almost all the elements present in $\mathbf{W'}$ can be considered as traffic elements.
+Considering a unique dictionary fitted to the sound scene under evaluation thus makes TI-NMF very effective when traffic is predominant, while the thresholding step makes it possible to discard the elements of the dictionary that deviate too much when the traffic is less present.
 
-Considering a unique dictionary fitted to the sound scene under evaluation thus makes TI-NMF very effective when traffic is predominant, while the thresholding step make it possible to discards the elements of the dictionary that deviate too much when the traffic is less present.
+\begin{figure*}[t!]
+\centering
+\begin{subfigure}{0.45\textwidth}
+\includegraphics[width=\linewidth]{./figures/evolutionK_AA.pdf}
+\caption{Influence of the experimental factor $\mathbf{K}$ (TI-NMF, $w_t$ = 0.5 s, $\beta$ = 2, $t_h$ = 0.32, 400 iterations).} \label{fig:grafic_K}
+\end{subfigure}
+\begin{subfigure}{0.45\textwidth}
+\includegraphics[width=\linewidth]{./figures/evolutionWt_AA.pdf}
+\caption{Influence of the experimental factor $\mathbf{w_t}$ (TI-NMF, $K$ = 200 s, $\beta$ = 2, $t_h$ = 0.32, 400 iterations).} \label{fig:grafic_wt}
+\end{subfigure}
+\caption{Influence of the number of elements $\mathbf{K}$ (\subref{fig:grafic_K}) and the temporal frame $\mathbf{w_t}$ (\subref{fig:grafic_wt}) on the $mMAE$ error with the best TI-NMF obtained in Table \ref{tab:results}.} \label{fig:grafic_K_wt}
+\end{figure*}
+
+To better understand the behavior of TI-NMF, the evolution of the normalized cost function ($D(\mathbf{V}\Vert \mathbf{WH})/(F\times N)$ as the duration of each audio sample is different) and the error $MAE$ as the function of the number of iterations are displayed for each sound environment in Figure \ref{fig:grafic_cost_mae}.
+For each sound environment, the cost function is very low after 400 iterations, meaning that $\mathbf{WH}$ approximates correctly $\mathbf{V}$.
+However, even if the convergence is not reached (Figure \ref{fig:grafic_cost}), the traffic error reconstruction is constant after the \nth{200} iteration (Figure \ref{fig:grafic_mae}), which means that the updates of $\mathbf{W}$ and $\mathbf{H}$ are mostly dedicated to the interfering sound sources.
+Also, even if the cost function is higher for the \textit{very noisy street} than the \textit{park}, the traffic sound level estimation is better in the \textit{very noisy street} sound environment due to the higher presence of this component in it.
+
+Finally, the influence of the number of elements $\mathbf{K}$ in $\mathbf{W}$ and the size of the temporal frame $\mathbf{w_t}$ on the $mMAE$ error made with TI-NMF, with the other experimental factors keep constant, are summarized in Figure \ref{fig:grafic_K_wt}.
+Naturally, the traffic error reconstruction decreases, with the increase of the number of basis in $\mathbf{W}$, as it allows more basis to describe the traffic sound source.
+If this improvement is major between 25 and 100 elements, its influence decreases between 100 and 200 elements suggesting that the increase of the size of $\mathbf{K}$ will not improve significantly the quality of the estimation.
+The temporal frame $\mathbf{w_t}$ is a less influential parameter for TI-NMF.
+As the matrices are updates at each iteration, it makes sense that the variations in the spectra shape in $\mathbf{W_0}$, due to this experimental factor, will disappear after a set of iterations.
 
@@
 Further analyses are required to extend the proposed method to other sound sources, such as birds or voices sounds, which can conveniently be done by replacing or adding elements in the dictionary.
-This utilization would be useful in the context of muti-source noise mapping that is gaining interest \cite{aumond2017Probabilistic, aletta2015soundscape}.
+This utilization would be useful in the context of muti-source noise mapping that is gaining interest \cite{aumond2018probabilistic, aletta2015soundscape}.
 Finally, the parameters selected in this study are valid for this evaluation corpus.
-Further analyses on various corpus of sound scenes are needed to evaluate the robustness of the method and select the most relevant approaches for specific sound environments (predominance of water or industrial sounds, rural environments \dots).
+Further analyses on various corpus of sound scenes are needed to evaluate the robustness of the method and select the most relevant approaches for specific sound environments (predominance of water or industrial sounds, rural environments\dots).
 
